---
title: "Compare Timeseries of Network Stats"
author: "M. Fisher"
date: "Written Aug 17, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

This document graphs network metrics over time for the US West Coast, and compares metrics across different forms of fisheries participation networks.


The three core network metrics are: 

1. Edge Density (alt: Mean Degree)

2. Weighted Network Centralization (alt: unweighted)

3. Weighted Modularity (alt:unweighted)


**Note that many network metrics are sensitive to network size, which will vary over the time series.** Network size has been incorporated into each graph as ___. Use the timeseries graphs for visual comparison of different network forms, but not for drawing major conclusions around network change / sensitivity. 



# Setup

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("foreign")) {install.packages("foreign")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("facetscales")) {devtools::install_github("zeehio/facetscales"); library(facetscales)}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
if(!require("ggpubr")) {install.packages("ggpubr")}
if(!require("cowplot")) {install.packages("cowplot")}
if(!require("PNWColors")) {devtools::install_github("jakelawler/PNWColors"); library(PNWColors)}
```
<br>

Select your directories.
```{r}
## directory with network statistics 
indir <- 'results/statistics'

## directory to save figures
pngdir <- 'results/figures'
```
<br>

What are the names of the `.csv` files with the network statistics that you want to compare? Would not suggest listing more than 4. Files should all be in the input directory (`indir`) specified above.
```{r}
filenames <- c("NetworkStats_2004_2019_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv", "NetworkStats_2004_2019_21pcgroups_totalrev5000_indivrev500_vessels_rmCrab_10pContribution.csv")
#c("NetworkStats_2004_2019_21pcgroups_rmCrab_5pContribution_bulk.csv",              "NetworkStats_2004_2019_21pcgroups_rmCrab_10pContribution_bulk.csv")
```
<br>

Provide descriptive names to differentiate the datasets drawn from the files listed above. 
```{r}
dataset_names <- c("Edges_Connectivity",
                   "Edges_Number of Vessels")
```
<br>

Specify which network metrics to retain from the full data set.
```{r}
metrics <- c("ed","nc_weighted","m_weighted")
```
<br>

Specify which port groups to retain from the full data set.
```{r}
myports <- c("North WA Coast", "Astoria", "Fort Bragg", "Morro Bay")
```
<br>



# Read in Data


Network metrics from different datasets. 
```{r}
for(i in seq(1,length(filenames))){
  tmpdat <- read.csv(here::here(indir, filenames[i])) %>%
    dplyr::select(pcgroup, y, N, all_of(metrics)) %>%
    filter(pcgroup %in% myports) %>%
    mutate(dataset=dataset_names[i])
  if(i==1){
    dat <- tmpdat
  } else{ dat <- bind_rows(dat,tmpdat)}
}

head(dat)
```
<br>

Check to make sure that all of the datasets were read in. 
```{r echo=FALSE}
with(dat,table(dataset,pcgroup))
```
<br>

Pivot longer to graph multiple metrics at once.
```{r}
dat_long <- dat %>%
  pivot_longer(cols=c(all_of(metrics)), names_to = "metric", values_to = "value")
```
<br>

Clean up metric names.
```{r}
dat_long$metric <- recode(dat_long$metric, 
                          N = "Size",
                          ed = "Edge Density",
                          mean_deg = "Mean Degree",
                          nc_weighted = "Central.",
                          nc = "Central. (UnW)",
                          m_weighted = "Mod.",
                          m = "Mod. (UnW)")
dat_long$metric <- factor(dat_long$metric, levels=c("Size", "Edge Density","Mean Degree", "Central.", "Central. (UnW)", "Mod.", "Mod. (UnW)"))
```
<br>



# 2. Graph Timeseries

## Figure 5.Time series of centralization 

... at four example ports (North WA Coast, Astoria, Fort Bragg, Morro Bay) from 2004-2020. Centralization was derived from vessel-level networks summarized at the port level, using edge definitions that follow equation 1 versus equation 2. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

```{r}
plot_metric="Central."
metric_label="Centralization"
```
<br>

```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
plotdat <- dat_long %>% filter(metric==plot_metric)
plotdat$pcgroup <- factor(plotdat$pcgroup,levels=myports)
plotdat.s <- dat_long %>% dplyr::select(pcgroup,y,N) %>% unique()
plotdat.s$pcgroup <- factor(plotdat.s$pcgroup,levels=myports)

# color palette
pal=pnw_palette("Sailboat",length(unique(plotdat$dataset)), type = "discrete") # or choose a different PNWColors palette

# main metric plot
fig5 <- ggplot(plotdat, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid(cols=vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylim(c(0,1)) +
  ylab(metric_label) + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank())
fig5
```
<br>

```{r echo=FALSE, fig.width=10}
# main metric plot
fig5a <- ggplot(plotdat, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid(cols=vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylim(c(0,1)) +
  ylab(metric_label) + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_blank(),
                     axis.text.y=element_text(size=12),
                     axis.title.x=element_blank(),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     plot.margin = margin(t=0, r=0, b=0, l=0, unit="cm"))
# network size plot 
fig5b <- ggplot(plotdat.s, (aes(x=y,y=N))) + 
  geom_point(color="grey52") + geom_line(color="grey52") +
  facet_grid(cols=vars(pcgroup)) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylab("Size") + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5, size=12),
                     axis.text.y=element_text(color="grey52"),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=11, color="grey52"),
                     strip.text=element_blank(),
                     legend.title=element_blank(),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     panel.border=element_rect(color="grey52"),
                     plot.margin = margin(t = 0, r = 4.8, b = 0, l = 0.63, unit = "cm"))

# gridded layout
#   if the left side of the plots don't line up, adjust plot.margin "l" argument in fig5b. 
#   if the right sides of the plots don't line up, adjust plot.margin "r" argument in fig5b.
lay <- rbind(c(1), c(2))
fig5ab <- grid.arrange(grobs=list(fig5a,fig5b), layout_matrix=lay, heights=c(1,0.6))

```
<br>
<br>
<br>

Both plots save as a png, with filename: `r paste0("Central_",min(plotdat$y), "_", max(plotdat$y), "_",length(myports),"pcgroups_", paste(dataset_names,collapse="-"))`
```{r echo=FALSE}
png(here::here(pngdir, paste0("Central_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), ".png")),
    res=200,height=800,width=1700)
fig5
dev.off()
```
```{r echo=FALSE}
png(here::here(pngdir, paste0("Central_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), "_wSize.png")),
    res=200,height=1000,width=1700)
grid.arrange(grobs=list(fig5a,fig5b), layout_matrix=lay, heights=c(1,0.6))
dev.off()
```
<br>

## Figure 6.Time series of edge density

...  at four example ports (North WA Coast, Astoria, Fort Bragg, Morro Bay) from 2004-2020. Edge density was derived from vessel-level networks summarized at the port level, using edge definitions that follow equation 1 versus equation 2. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

```{r}
plot_metric="Edge Density"
metric_label="Edge Density"
```
<br>

```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
plotdat <- dat_long %>% filter(metric==plot_metric)
plotdat$pcgroup <- factor(plotdat$pcgroup,levels=myports)
plotdat.s <- dat_long %>% dplyr::select(pcgroup,y,N) %>% unique()
plotdat.s$pcgroup <- factor(plotdat.s$pcgroup,levels=myports)

# color palette
pal=pnw_palette("Sailboat",length(unique(plotdat$dataset)), type = "discrete") # or choose a different PNWColors palette

# main metric plot
fig6 <- ggplot(plotdat, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid(cols=vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylim(c(0,1)) +
  ylab(metric_label) + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank())
fig6
```
<br>

```{r echo=FALSE, fig.width=10}
# main metric plot
fig6a <- ggplot(plotdat, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid(cols=vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylim(c(0,1)) +
  ylab(metric_label) + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_blank(),
                     axis.text.y=element_text(size=12),
                     axis.title.x=element_blank(),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     plot.margin = margin(t=0, r=0, b=0, l=0, unit="cm"))
# network size plot 
fig6b <- ggplot(plotdat.s, (aes(x=y,y=N))) + 
  geom_point(color="grey52") + geom_line(color="grey52") +
  facet_grid(cols=vars(pcgroup)) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylab("Size") + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5, size=12),
                     axis.text.y=element_text(color="grey52"),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=11, color="grey52"),
                     strip.text=element_blank(),
                     legend.title=element_blank(),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     panel.border=element_rect(color="grey52"),
                     plot.margin = margin(t = 0, r = 4.8, b = 0, l = 0.63, unit = "cm"))

# gridded layout
#   if the left side of the plots don't line up, adjust plot.margin "l" argument in fig5b. 
#   if the right sides of the plots don't line up, adjust plot.margin "r" argument in fig5b.
lay <- rbind(c(1), c(2))
fig6ab <- grid.arrange(grobs=list(fig6a,fig6b), layout_matrix=lay, heights=c(1,0.6))

```
<br>
<br>
<br>

Both plots save as a png, with filename: `r paste0("ED_",min(plotdat$y), "_", max(plotdat$y), "_",length(myports),"pcgroups_", paste(dataset_names,collapse="-"))`
```{r echo=FALSE}
png(here::here(pngdir, paste0("ED_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), ".png")),
    res=200,height=800,width=1700)
fig6
dev.off()
```
```{r echo=FALSE}
png(here::here(pngdir, paste0("ED_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), "_wSize.png")),
    res=200,height=1000,width=1700)
grid.arrange(grobs=list(fig6a,fig6b), layout_matrix=lay, heights=c(1,0.6))
dev.off()
```
<br>



## Figure 7.Time series of modularity

...  at four example ports (North WA Coast, Astoria, Fort Bragg, Morro Bay) from 2004-2020. Modularity was derived from vessel-level networks summarized at the port level, using edge definitions that follow equation 1 versus equation 2. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.


```{r}
plot_metric="Mod."
metric_label="Modularity"
```
<br>

```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
plotdat <- dat_long %>% filter(metric==plot_metric)
plotdat$pcgroup <- factor(plotdat$pcgroup,levels=myports)
plotdat.s <- dat_long %>% dplyr::select(pcgroup,y,N) %>% unique()
plotdat.s$pcgroup <- factor(plotdat.s$pcgroup,levels=myports)

# color palette
pal=pnw_palette("Sailboat",length(unique(plotdat$dataset)), type = "discrete") # or choose a different PNWColors palette

# main metric plot
fig7 <- ggplot(plotdat, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid(cols=vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylim(c(-0.5,0.5)) +
  ylab(metric_label) + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank())
fig7
```
<br>

```{r echo=FALSE, fig.width=10}
# main metric plot
fig7a <- ggplot(plotdat, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid(cols=vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylim(c(-0.5, 0.5)) +
  ylab(metric_label) + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_blank(),
                     axis.text.y=element_text(size=12),
                     axis.title.x=element_blank(),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     plot.margin = margin(t=0, r=0, b=0, l=0, unit="cm"))
# network size plot 
fig7b <- ggplot(plotdat.s, (aes(x=y,y=N))) + 
  geom_point(color="grey52") + geom_line(color="grey52") +
  facet_grid(cols=vars(pcgroup)) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylab("Size") + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5, size=12),
                     axis.text.y=element_text(color="grey52"),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=11, color="grey52"),
                     strip.text=element_blank(),
                     legend.title=element_blank(),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     panel.border=element_rect(color="grey52"),
                     plot.margin = margin(t = 0, r = 4.8, b = 0, l = 0.84, unit = "cm"))

# gridded layout
#   if the left side of the plots don't line up, adjust plot.margin "l" argument in fig5b. 
#   if the right sides of the plots don't line up, adjust plot.margin "r" argument in fig5b.
lay <- rbind(c(1), c(2))
fig7ab <- grid.arrange(grobs=list(fig7a,fig7b), layout_matrix=lay, heights=c(1,0.6))

```
<br>
<br>
<br>

Both plots save as a png, with filename: `r paste0("Modularity_",min(plotdat$y), "_", max(plotdat$y), "_",length(myports),"pcgroups_", paste(dataset_names,collapse="-"))`
```{r echo=FALSE}
png(here::here(pngdir, paste0("Modularity_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), ".png")),
    res=200,height=800,width=1700)
fig7
dev.off()
```
```{r echo=FALSE}
png(here::here(pngdir, paste0("Modularity_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), "_wSize.png")),
    res=200,height=1000,width=1700)
grid.arrange(grobs=list(fig7a,fig7b), layout_matrix=lay, heights=c(1,0.6))
dev.off()
```
<br>



## Extra Code: Multiple Metrics

Graph all metrics in a single plot area. 

```{r echo=FALSE}
# Set y axis limits and breaks by variable
scales_y <- list(
  `Edge Density` = scale_y_continuous(limits = c(0,1)),
  # `Mean Degree` = scale_y_continuous(limits = c(0,max(dat$mean_deg) + 0.5)),
  `Central.` = scale_y_continuous(limits = c(0,1)),
  `Central. (UnW)` = scale_y_continuous(limits=c(0,1)),
  `Mod.` = scale_y_continuous(limits=c(-0.5,0.5)),
  `Mod. (UnW)` = scale_y_continuous(limits=c(-0.5,0.5))
)

# edit data set for plotting
dat_long$pcgroup <- factor(dat_long$pcgroup, levels=myports)

datN <- dplyr::select(dat_long, c(pcgroup,y,N)) %>% unique()
datN$pcgroup <- factor(datN$pcgroup, levels=myports)
```


```{r echo=FALSE, fig.width=9}
figS <- ggplot(dat_long, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid_sc(rows = vars(metric), cols = vars(pcgroup), scales = list(y = scales_y),as.table=FALSE) +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  xlab("Year") + ylab("") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5, size=12),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_blank(),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank())

figS
```
<br>

```{r echo=FALSE, fig.width=10}
plot.all <- ggplot(dat_long, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid_sc(rows = vars(metric), cols = vars(pcgroup), scales = list(y = scales_y),as.table=FALSE) +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  xlab("Year") + ylab("") +
  theme_bw() + theme(axis.text.x=element_blank(),
                     axis.text.y=element_text(size=12),
                     axis.title.x=element_blank(),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     plot.margin = margin(t=0, r=0, b=0, l=0, unit="cm"))
  
  
plot.N <- ggplot(datN, (aes(x=y,y=N))) + 
  geom_point(color="grey52") + geom_line(color="grey52") +
  facet_grid(cols=vars(pcgroup)) +
  scale_x_continuous(breaks=seq(min(plotdat$y), max(plotdat$y), by=2), labels=seq(min(plotdat$y), max(plotdat$y), by=2)) +
  ylab("Size") + xlab("Year") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5, size=12),
                     axis.text.y=element_text(color="grey52"),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=11, color="grey52"),
                     strip.text=element_blank(),
                     legend.title=element_blank(),
                     panel.grid.major.y=element_blank(),
                     panel.grid.major.=element_line(color="grey82", linetype=2),
                     panel.grid.minor=element_blank(),
                     panel.border=element_rect(color="grey52"),
                     plot.margin = margin(t = 0, r = 5.5, b = 0, l = 0.83, unit = "cm"))

# gridded layout
#   if the left side of the plots don't line up, adjust plot.margin "l" argument in fig5b. 
#   if the right sides of the plots don't line up, adjust plot.margin "r" argument in fig5b.
lay <- rbind(c(1), c(2))
grid.arrange(grobs=list(plot.all,plot.N), layout_matrix=lay, heights=c(1,0.4))
```
<br>


Both plots save as a png, with filename: `r paste0("Central-ED-Mod_",min(plotdat$y), "_", max(plotdat$y), "_",length(myports),"pcgroups_", paste(dataset_names,collapse="-"))`
```{r echo=FALSE}
png(here::here(pngdir, paste0("Central-ED-Mod_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), ".png")),
    res=200,height=800,width=1700)
figS
dev.off()
```
```{r echo=FALSE}
png(here::here(pngdir, paste0("Central-ED-Mod_",min(plotdat$y), "_", max(plotdat$y), "_",
                                       length(myports),"pcgroups_", paste(dataset_names,collapse="-"), "_wSize.png")),
    res=200,height=1200,width=1700)
grid.arrange(grobs=list(plot.all,plot.N), layout_matrix=lay, heights=c(1,0.4))
dev.off()
```
<br>













