---
title: "Plot Annual Participation Networks for SSC-ES Review September 2021"
author: "M. Fisher, J.F. Samhouri"
date: "Written Jan. 6, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Summarize networks for SSC-ES presentation

# Description

This document generates fisheries participation networks for the US West Coast using a variety of approaches. 

It visually summarizes a subset of networks for presentation to the SSC-ES in September 2021.

Plot a batch of annual fisheries participation networks (from a list of igraph objects) by port group or state, and year. This script uses a circular layout, with network nodes colored according to the gear group used in each fishery. **The nodes are consistently located across plots,** so that plots can be more easily compared through time or across port groups.

It depends on the code "06_create_networks_4ways.Rmd", which makes network graphs and calculates network statistics.
 

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("igraph")) {install.packages("igraph")}
if(!require("RColorBrewer")) {install.packages("RColorBrewer")}
if(!require("stringr")) {install.packages("stringr")}
```
<br>

And calls the following functions:
```{r}
source(here::here('R/color_network_nodes.R'))
source(here::here('R/plot_network_bulk_comparable.R')) # note this script is flexible whether we want to plot vessel-level or aggregate (bulk) networks. simply change the file_suffix. similarly, it is flexible whether we want to plot port group or state networks, simply change how 'port_group' is defined based on unique(V(g)$p)
```
<br>


# User Inputs 

Select your directories.
```{r}
## location of igraph objects
indir_ports = 'data/networks/participation_vessel_ports'
indir_states = 'data/networks/participation_vessel_states'
indir_ports_agg = 'data/networks/participation_agg_ports'
indir_states_agg = 'data/networks/participation_agg_states'

## output directory for network viz
pngdir_ports = 'data/networks/participation_vessel_ports/plots/comparable'
pngdir_states = 'data/networks/participation_vessel_states/plots/comparable'
pngdir_ports_agg = 'data/networks/participation_agg_ports/plots/comparable'
pngdir_states_agg = 'data/networks/participation_agg_states/plots/comparable'
pngdir_compare = 'data/networks/participation_compare'

```
<br>

Identify the crab years, port groups, and states you would like to produce networks for. 
```{r}
## (crab) years
years <- seq(2004,2019)
# years <- 2008 # testing script, 1/6
# years <- 2018 # testing script, 1/8
# years <- 2014 # testing script, 1/27

## port groups
myports <- c("Puget Sound","North WA Coast","WA Coast","Other Coastal Washington","Astoria","Tillamook","Columbia River","Newport","Coos Bay","Brookings","Crescent City","Eureka","Morro Bay","Fort Bragg","Bodega Bay","San Francisco","San Diego","Monterey","Santa Barbara","Los Angeles","Unknown Ports" )
#myports <- c("BDG") # testing script, 1/6
#myports <- c("Columbia River") # testing script for single node networks, 1/7
# myports <- c("Monterey") # testing script, 1/8
# myports <- c("Bodega Bay") # testing script, 1/27/21

## west coast states
mystates <- c("C", "O","W")

```
<br>


# 1: Read in data

## Figure 1

Comparison of 2019-2020 vessel-level networks for four example ports (North WA Coast, Astoria, Fort Bragg, Morro Bay). Nodes (fisheries) are included in each network either with a minimum total fisheries and individual fishery revenue cutoff for vessels of `$1`, or with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`. Edges based on equation 1.

Files of interest
```{r}

y <- 2019
contr_cutoff <- 0.1

nwc11 <- 'North WA Coast_2019_totalrev1_indivrev1_connectivity'
nwc55 <- 'North WA Coast_2019_totalrev5000_indivrev500_connectivity'
fig1_nwc <- c(nwc11,nwc55)

ast11 <- 'Astoria_2019_totalrev1_indivrev1_connectivity'
ast55 <- 'Astoria_2019_totalrev5000_indivrev500_connectivity'
fig1_ast <- c(ast11,ast55)

ftb11 <- 'Fort Bragg_2019_totalrev1_indivrev1_connectivity'
ftb55 <- 'Fort Bragg_2019_totalrev5000_indivrev500_connectivity'
fig1_ftb <- c(ftb11,ftb55)

mor11 <- 'Morro Bay_2019_totalrev1_indivrev1_connectivity'
mor55 <- 'Morro Bay_2019_totalrev5000_indivrev500_connectivity'
fig1_mor <- c(mor11,mor55)

```
<br>

Read the igraph objects that you want plotted in a comparable way into a list. This for loop will also apply the functions to recolor and rename vertices

### North WA Coast list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig1_nwc_list <- list()
fig1_nwc_names <- c()
i=1
for(g in fig1_nwc){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig1_nwc_list[[i]] <- tmpgraph
      fig1_nwc_names[i] <- g
      i = i + 1
      } 
}
names(fig1_nwc_list) <- fig1_nwc_names
```
<br>

### Astoria list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig1_ast_list <- list()
fig1_ast_names <- c()
i=1
for(g in fig1_ast){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig1_ast_list[[i]] <- tmpgraph
      fig1_ast_names[i] <- g
      i = i + 1
      } 
}
names(fig1_ast_list) <- fig1_ast_names
```
<br>

### Ft Bragg list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig1_ftb_list <- list()
fig1_ftb_names <- c()
i=1
for(g in fig1_ftb){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig1_ftb_list[[i]] <- tmpgraph
      fig1_ftb_names[i] <- g
      i = i + 1
      } 
}
names(fig1_ftb_list) <- fig1_ftb_names
```
<br>

### Morro Bay list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig1_mor_list <- list()
fig1_mor_names <- c()
i=1
for(g in fig1_mor){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig1_mor_list[[i]] <- tmpgraph
      fig1_mor_names[i] <- g
      i = i + 1
      } 
}
names(fig1_mor_list) <- fig1_mor_names
```
<br>

## Figure 2

Comparison of 2019-2020 vessel-level networks for four example ports (North WA Coast, Astoria, Fort Bragg, Morro Bay), using edge definitions that follow equation 1 versus equation 2. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

Files of interest
```{r}

y <- 2019
contr_cutoff <- 0.1

nwc55_v <- 'North WA Coast_2019_totalrev5000_indivrev500_vessels'
fig2_nwc <- c(nwc55, nwc55_v)

ast55_v <- 'Astoria_2019_totalrev5000_indivrev500_vessels'
fig2_ast <- c(ast55, ast55_v)

ftb55_v <- 'Fort Bragg_2019_totalrev5000_indivrev500_vessels'
fig2_ftb <- c(ftb55, ftb55_v)

mor55_v <- 'Morro Bay_2019_totalrev5000_indivrev500_vessels'
fig2_mor <- c(mor55, mor55_v)

```
<br>

Read the igraph objects that you want plotted in a comparable way into a list. This for loop will also apply the functions to recolor and rename vertices

### North WA Coast list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig2_nwc_list <- list()
fig2_nwc_names <- c()
i=1
for(g in fig2_nwc){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig2_nwc_list[[i]] <- tmpgraph
      fig2_nwc_names[i] <- g
      i = i + 1
      } 
}
names(fig2_nwc_list) <- fig2_nwc_names
```
<br>

### Astoria list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig2_ast_list <- list()
fig2_ast_names <- c()
i=1
for(g in fig2_ast){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig2_ast_list[[i]] <- tmpgraph
      fig2_ast_names[i] <- g
      i = i + 1
      } 
}
names(fig2_ast_list) <- fig2_ast_names
```
<br>

### Ft Bragg list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig2_ftb_list <- list()
fig2_ftb_names <- c()
i=1
for(g in fig2_ftb){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig2_ftb_list[[i]] <- tmpgraph
      fig2_ftb_names[i] <- g
      i = i + 1
      } 
}
names(fig2_ftb_list) <- fig2_ftb_names
```
<br>

### Morro Bay list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig2_mor_list <- list()
fig2_mor_names <- c()
i=1
for(g in fig2_mor){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
  # p = "Brookings"  # testing script
    if(file.exists(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',g, '_rmCrab.rds')))
      V(tmpgraph)$p <- g # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig2_mor_list[[i]] <- tmpgraph
      fig2_mor_names[i] <- g
      i = i + 1
      } 
}
names(fig2_mor_list) <- fig2_mor_names
```
<br>

## Figure 3

Comparison of 2019-2020 vessel-level and aggregate networks for four example ports (North WA Coast, Astoria, Fort Bragg, Morro Bay), using edge definitions that follow equation 1. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

Files of interest
```{r}

y <- 2019
contr_cutoff <- 0.1

nwc55_agg <- 'North WA Coast_2019_totalrev5000_indivrev500_connectivity'
fig3_nwc <- c(nwc55, nwc55_agg)

ast55_agg <- 'Astoria_2019_totalrev5000_indivrev500_connectivity'
fig3_ast <- c(ast55, ast55_agg)

ftb55_agg <- 'Fort Bragg_2019_totalrev5000_indivrev500_connectivity'
fig3_ftb <- c(ftb55, ftb55_agg)

mor55_agg <- 'Morro Bay_2019_totalrev5000_indivrev500_connectivity'
fig3_mor <- c(mor55, mor55_agg)

```
<br>

Read the igraph objects that you want plotted in a comparable way into a list. This for loop will also apply the functions to recolor and rename vertices

### North WA Coast list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig3_nwc_list <- list()
fig3_nwc_names <- c()

# vessel-level
i=1
if(file.exists(here::here(indir_ports, paste0('igraph_',fig3_nwc[1], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',fig3_nwc[1], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_nwc[1],"_","vessel") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_nwc_list[[i]] <- tmpgraph
      fig3_nwc_names[i] <-  paste0(fig3_nwc[1],"_","vessel")
} 

# aggregate
i=2
if(file.exists(here::here(indir_ports_agg, paste0('igraph_',fig3_nwc[2], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports_agg, paste0('igraph_',fig3_nwc[2], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_nwc[2],"_","agg") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_nwc_list[[i]] <- tmpgraph
      fig3_nwc_names[i] <- paste0(fig3_nwc[2],"_","agg")
} 

names(fig3_nwc_list) <- fig3_nwc_names
```
<br>

### Astoria list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig3_ast_list <- list()
fig3_ast_names <- c()

# vessel-level
i=1
if(file.exists(here::here(indir_ports, paste0('igraph_',fig3_ast[1], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',fig3_ast[1], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_ast[1],"_","vessel") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_ast_list[[i]] <- tmpgraph
      fig3_ast_names[i] <-  paste0(fig3_ast[1],"_","vessel")
} 

# aggregate
i=2
if(file.exists(here::here(indir_ports_agg, paste0('igraph_',fig3_ast[2], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports_agg, paste0('igraph_',fig3_ast[2], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_ast[2],"_","agg") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_ast_list[[i]] <- tmpgraph
      fig3_ast_names[i] <- paste0(fig3_ast[2],"_","agg")
} 

names(fig3_ast_list) <- fig3_ast_names
```
<br>

### Ft Bragg list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig3_ftb_list <- list()
fig3_ftb_names <- c()

# vessel-level
i=1
if(file.exists(here::here(indir_ports, paste0('igraph_',fig3_ftb[1], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',fig3_ftb[1], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_ftb[1],"_","vessel") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_ftb_list[[i]] <- tmpgraph
      fig3_ftb_names[i] <-  paste0(fig3_ftb[1],"_","vessel")
} 

# aggregate
i=2
if(file.exists(here::here(indir_ports_agg, paste0('igraph_',fig3_ftb[2], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports_agg, paste0('igraph_',fig3_ftb[2], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_ftb[2],"_","agg") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_ftb_list[[i]] <- tmpgraph
      fig3_ftb_names[i] <- paste0(fig3_ftb[2],"_","agg")
} 

names(fig3_ftb_list) <- fig3_ftb_names
```
<br>

### Morro Bay list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig3_mor_list <- list()
fig3_mor_names <- c()

# vessel-level
i=1
if(file.exists(here::here(indir_ports, paste0('igraph_',fig3_mor[1], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports, paste0('igraph_',fig3_mor[1], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_mor[1],"_","vessel") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_mor_list[[i]] <- tmpgraph
      fig3_mor_names[i] <-  paste0(fig3_mor[1],"_","vessel")
} 

# aggregate
i=2
if(file.exists(here::here(indir_ports_agg, paste0('igraph_',fig3_mor[2], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_ports_agg, paste0('igraph_',fig3_mor[2], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig3_mor[2],"_","agg") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig3_mor_list[[i]] <- tmpgraph
      fig3_mor_names[i] <- paste0(fig3_mor[2],"_","agg")
} 

names(fig3_mor_list) <- fig3_mor_names
```
<br>

## Figure 4

Comparison of 2019-2020 vessel-level and aggregate networks summarized at the scale of each of the 3 U.S. West Coast states (Washington, Oregon, California), using edge definitions that follow equation 1. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

Files of interest
```{r}

y <- 2019
contr_cutoff <- 0.1

wa_55 <- 'W_2019_totalrev5000_indivrev500_connectivity'
fig4_wa <- c(wa_55, wa_55)

or_55 <- 'O_2019_totalrev5000_indivrev500_connectivity'
fig4_or <- c(or_55, or_55)

ca_55 <- 'C_2019_totalrev5000_indivrev500_connectivity'
fig4_ca <- c(ca_55, ca_55)

```
<br>

Read the igraph objects that you want plotted in a comparable way into a list. This for loop will also apply the functions to recolor and rename vertices

### WA list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig4_wa_list <- list()
fig4_wa_names <- c()

# vessel-level
i=1
if(file.exists(here::here(indir_states, paste0('igraph_',fig4_wa[1], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_states, paste0('igraph_',fig4_wa[1], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig4_wa[1],"_","vessel") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig4_wa_list[[i]] <- tmpgraph
      fig4_wa_names[i] <-  paste0(fig4_wa[1],"_","vessel")
} 

# aggregate
i=2
if(file.exists(here::here(indir_states_agg, paste0('igraph_',fig4_wa[2], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_states_agg, paste0('igraph_',fig4_wa[2], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig4_wa[2],"_","agg") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig4_wa_list[[i]] <- tmpgraph
      fig4_wa_names[i] <- paste0(fig4_wa[2],"_","agg")
} 

names(fig4_wa_list) <- fig4_wa_names
```
<br>

### OR list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig4_or_list <- list()
fig4_or_names <- c()

# vessel-level
i=1
if(file.exists(here::here(indir_states, paste0('igraph_',fig4_or[1], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_states, paste0('igraph_',fig4_or[1], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig4_or[1],"_","vessel") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig4_or_list[[i]] <- tmpgraph
      fig4_or_names[i] <-  paste0(fig4_or[1],"_","vessel")
} 

# aggregate
i=2
if(file.exists(here::here(indir_states_agg, paste0('igraph_',fig4_or[2], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_states_agg, paste0('igraph_',fig4_or[2], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig4_or[2],"_","agg") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig4_or_list[[i]] <- tmpgraph
      fig4_or_names[i] <- paste0(fig4_or[2],"_","agg")
} 

names(fig4_or_list) <- fig4_or_names
```

### CA list

```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

fig4_ca_list <- list()
fig4_ca_names <- c()

# vessel-level
i=1
if(file.exists(here::here(indir_states, paste0('igraph_',fig4_ca[1], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_states, paste0('igraph_',fig4_ca[1], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig4_ca[1],"_","vessel") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig4_ca_list[[i]] <- tmpgraph
      fig4_ca_names[i] <-  paste0(fig4_ca[1],"_","vessel")
} 

# aggregate
i=2
if(file.exists(here::here(indir_states_agg, paste0('igraph_',fig4_ca[2], '_rmCrab.rds')))){
      tmpgraph <- readRDS(here::here(indir_states_agg, paste0('igraph_',fig4_ca[2], '_rmCrab.rds')))
      V(tmpgraph)$p <- paste0(fig4_ca[2],"_","agg") # this is important because it dictates the name of the output file
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      fig4_ca_list[[i]] <- tmpgraph
      fig4_ca_names[i] <- paste0(fig4_ca[2],"_","agg")
} 

names(fig4_ca_list) <- fig4_ca_names
```

# 2: Plot Comparable Networks

This function plots networks using the circular layout, with consistent node placement across plots to allow for visual comparison. 

Plots will be written out to .png files in the function. This is because the igraph `plot` function won't save into an object that can be returned.

## Figure 1

```{r}
plot_comparable_networks(fig1_nwc_list, outdir=pngdir_compare)
plot_comparable_networks(fig1_ast_list, outdir=pngdir_compare)
plot_comparable_networks(fig1_ftb_list, outdir=pngdir_compare)
plot_comparable_networks(fig1_mor_list, outdir=pngdir_compare)
```


The previous section saves each plotted network into its own file - in some cases, it may be useful to have 2-6 network plots saved as a single image. 


The multi-network .png will be written out to a file in this function. Note that the file suffix should be more distinguishable / verbose; the function only adds "MultiNetwork_" in front of it, with no port group / year details like in the individual graphs. 
```{r}
plot_comparable_networks(fig1_nwc_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("North WA Coast_2019_fig1"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig1_ast_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Astoria_2019_fig1"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig1_ftb_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Fort Bragg_2019_fig1"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig1_mor_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Morro Bay_2019_fig1"),
                         grid_layout=c(2,1))
```
<br>

## Figure 2

```{r}
plot_comparable_networks(fig2_nwc_list, outdir=pngdir_compare)
plot_comparable_networks(fig2_ast_list, outdir=pngdir_compare)
plot_comparable_networks(fig2_ftb_list, outdir=pngdir_compare)
plot_comparable_networks(fig2_mor_list, outdir=pngdir_compare)
```


The previous section saves each plotted network into its own file - in some cases, it may be useful to have 2-6 network plots saved as a single image. 


The multi-network .png will be written out to a file in this function. Note that the file suffix should be more distinguishable / verbose; the function only adds "MultiNetwork_" in front of it, with no port group / year details like in the individual graphs. 
```{r}
plot_comparable_networks(fig2_nwc_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("North WA Coast_2019_fig2"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig2_ast_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Astoria_2019_fig2"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig2_ftb_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Fort Bragg_2019_fig2"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig2_mor_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Morro Bay_2019_fig2"),
                         grid_layout=c(2,1))
```
<br>


## Figure 3

```{r}
plot_comparable_networks(fig3_nwc_list, outdir=pngdir_compare)
plot_comparable_networks(fig3_ast_list, outdir=pngdir_compare)
plot_comparable_networks(fig3_ftb_list, outdir=pngdir_compare)
plot_comparable_networks(fig3_mor_list, outdir=pngdir_compare)
```


The previous section saves each plotted network into its own file - in some cases, it may be useful to have 2-6 network plots saved as a single image. 


The multi-network .png will be written out to a file in this function. Note that the file suffix should be more distinguishable / verbose; the function only adds "MultiNetwork_" in front of it, with no port group / year details like in the individual graphs. 
```{r}
plot_comparable_networks(fig3_nwc_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("North WA Coast_2019_fig3"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig3_ast_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Astoria_2019_fig3"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig3_ftb_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Fort Bragg_2019_fig3"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig3_mor_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("Morro Bay_2019_fig3"),
                         grid_layout=c(2,1))
```
<br>

## Figure 4

```{r}
plot_comparable_networks(fig4_wa_list, outdir=pngdir_compare)
plot_comparable_networks(fig4_or_list, outdir=pngdir_compare)
plot_comparable_networks(fig4_ca_list, outdir=pngdir_compare)
```


The previous section saves each plotted network into its own file - in some cases, it may be useful to have 2-6 network plots saved as a single image. 


The multi-network .png will be written out to a file in this function. Note that the file suffix should be more distinguishable / verbose; the function only adds "MultiNetwork_" in front of it, with no port group / year details like in the individual graphs. 
```{r}
plot_comparable_networks(fig4_wa_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("WA_2019_fig4"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig4_or_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("OR_2019_fig4"),
                         grid_layout=c(2,1))

plot_comparable_networks(fig4_ca_list, outdir=pngdir_compare, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("CA_2019_fig4"),
                         grid_layout=c(2,1))

```
<br>



