---
title: "Figure 1: 3-year average Participation Networks"
author: M. Fisher
date: "Written Mar. 29, 2019. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in Fisher et al.
fontsize: 11pt
---


# Description

This code generates the 3-year average "pre-shock" participation networks shown in Figure 1 of Fisher et al. This involves (1) creating a new adjacency matrix with all three crab years of fish ticket data to generate a network object, (2) pulling vertex attributes from the annual networks made for the main analysis, (3) averaging vertex attributes across annual networks, (4) applying averaged vertex attributes to the multi-year network object and filtering out any fisheries for which there are no data in the annual networks.

Section 3 uses network metrics to compare the complexity of the 3-year average network to the annual networks in that crab year range. This section was initially written to help qualitatively compare the different network visualizations; 3-year network metrics were not used in Fisher et al. 
<br>
<br>

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## start time for full script
script_start_time <- Sys.time()

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```


This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("here")) {install.packages("here")}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("igraph")) {install.packages("igraph")}
```
<br>

And calls the following functions. 
```{r}
source(here::here('R','generate_multiyr_participation_network.R'))
source(here::here('R','average_participation_network.R'))
source(here::here('R','calc_network_stats.R'))
source(here::here('R','plot_3yr_networks.R'))
```
<br>

# User Inputs 

Select your directories.
```{r}
## location of fish ticket with metiers assigned, metier key for each port group
indir = 'data/input/metiers'

## location of annual participation networks
netdir = 'data/networks/participation'

## output directory for igraph object
outdir = 'data/networks/participation/3yr'

## output directory where annual network stat summaries are saved 
statdir = 'results/statistics'

## output file for 3-yr stat summaries
statfile = '2013_2015_CA_CrabPorts_noSubgraph_3yr_avg.csv'

## output directory for network plots
pngdir = 'results/figures/Fig1'
```
<br>

Identify the crab years and port groups that you would like to produce networks for. The port groups vector should include only those port groups which are present in the single data file produced with script 05. 
```{r}
## (crab) years
years <- seq(2013,2015)

## port groups
myports <- c("CCA", "ERA", "BGA", "BDA", "SFA", "MRA", "MNA")
```
<br>

Some Dungeness crab landings may be recorded prior to the official opening date of the season as a part of domoic acid testing. We remove these landings because we are interested in flows of fishers between fisheries as a result of within-season activity.
```{r}
rm_crab = TRUE
```
<br>

For confidentiality, three or more vessels much be participating in a given fishery for that fishery to be included in the networks. To avoid inclusion of rare or minimal fishing activity, a given fishery must contribute to at least 10% of a vessel's seasonal revenue to be included in the network data. 
Both cutoff values can be increased if more stringent fishery inclusion criteria are preferred, and the contribution cutoff may be decreased (i.e. for summaries over shorter temporal / smaller spatial scales).
```{r}
vessel_cutoff <- 3
contr_cutoff <- 0.10
```
<br>


Set these objects according to your user inputs from Scripts 1 and 3
```{r}
k <- 1
```
<br>


# 1: Read in fish ticket data

Read in the landings data, from the file containing all fish tickets across port groups, within a single crab year.
```{r}
for(y in years){
  if(y %in% ref_years){
    tmptix <- read.csv(here::here(indir, paste0("FishTix_crab", y, "_metiers.csv")), stringsAsFactors = FALSE) %>%
    filter(pcgroup %in% myports)
  } else{
  tmptix <- read.csv(here::here(indir, paste0("FishTix_crab", y, "_metiers_k",k,"_recoded.csv")), stringsAsFactors = FALSE) %>%
    filter(pcgroup %in% myports)
  }
  if(exists('fishtix')){
    fishtix <- rbind(fishtix, tmptix)
  } else{ fishtix <- tmptix }
}
rm(tmptix)
```
<br>

Get the opening dates associated with each port group, for each crab year. 
```{r}
dates_df <- read.delim(here::here('data/input','crab_open_dates.txt')); head(dates_df)
```
<br>

Make sure that all dates are `Posixt` objects.
```{r}
dates_df$odate <- mdy(dates_df$odate)
fishtix$tdate <- date(parse_date_time(fishtix$tdate, orders=c("ymd", "mdy")))
```
<br>

# 2: 3-year Average Participation Networks

For each season and port group, this code will:

1. Save a list of the annual graph objects

2. Split the fish ticket data into an early and late season

2. Create a `graph` object for the multi-year fisheries participation network, and then save the object as an R data file

3. Average the annual graph vertices, then edit the multi-year network

4. Calculate network-level statistics, and save to an output data frame.

```{r}
for(p in myports){
  for(c in c("early","late")){
    ### get igraph objects for each year ###
    graphs_list = list()
    for(i in seq(1,length(years))){
      tmpgraph <- readRDS(here::here(netdir, paste0("igraph_",p,"_",years[i],"_",c,"_rmCrab.rds")))
      graphs_list[[i]] = tmpgraph
    }
    
    ### get data for multiyr graph ###
    open_info <- filter(dates_df, pcgroup == p)
    odate <- open_info$odate
    ## pull fish ticket data for given year, port group, before Dcrab opened
    sub_fishtix <- fishtix %>%
      filter(pcgroup == p) %>%
      mutate(opening_date = paste0(crab_year+1, "-", month(odate), "-", day(odate))) %>%
      mutate(response = ifelse(ymd(tdate) < ymd(opening_date), "early", "late"))
    # filter data for response period
    mydat <- mydat %>%
      filter(response==period)
  
    ### create avg graph ###
    # multi-year graph
    multiyr_g <- participation_network_multiyr(mydat, filter = TRUE, filter_subgraph =FALSE, min_vessels = 3, min_contribution = 0.10)
    # averaged multi-year graph
    avg_g <- average_participation_networks(multiyr_graph=multiyr_g,annual_graphs=graphs_list,period=c,years=years)[[1]]
    
    # write out graph
    saveRDS(multiyr_g, here::here(outdir, paste0("igraph_",p,"_3yr_",c,".rds")))
    saveRDS(avg_g, here::here(outdir, paste0("igraph_",p,"_3yrAVG_",c,".rds")))
    
    ### get network stats ###
    tmp_stats <- net_stats(tmp_g, y = paste0(years[[1]], "_", years[[length(years)]]), stats="network")
    tmp_stats <- mutate(tmp_stats, period=c, pcgroup=p)
    
    # Save network-level statistics output
    if(p == myports[1] & c=="early"){
      mystats <- tmp_stats
    } else{
      mystats <- rbind(mystats, tmp_stats)
    }

  } # end (for c)
  print(paste0("done with port group ", p))
} # end (for p)
```
<br>

Save the network statistics from the averaged graphs.
```{r}
write.csv(mystats,here::here(statdir, statfile))
```
<br>


# 3. Assess 3-year Networks

Compare the network statistics between the annual networks, and the 3-year average network. 


```{r}
stats1 <- read.csv(here::here(statdir, "2008_2017_CA_CrabPorts_2015HAB_NoSubgraph.csv"))
stats1 <- filter(stats1, y %in% years)
```
<br>

```{r}
myplot.n <- ggplot() +
  geom_boxplot(data=stats1, aes(x=pcgroup,y=N, fill=period)) +
  geom_point(data=mystats, aes(x=pcgroup,y=N, col=period), size=2) +
  geom_point(data=mystats, aes(x=pcgroup,y=N), size=2, pch=1) +
  scale_color_manual(values=c("red","turquoise4")) +
  ggtitle("Node Count") +
  theme(legend.position="none",
        axis.title.y=element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))
myplot.ed <- ggplot() +
  geom_boxplot(data=stats1, aes(x=pcgroup,y=ed, fill=period)) +
  geom_point(data=mystats, aes(x=pcgroup,y=ed, col=period), size=2) +
  geom_point(data=mystats, aes(x=pcgroup,y=ed), size=2, pch=1) +
  scale_color_manual(values=c("red","turquoise4")) +
  ggtitle("Edge Density") +
  theme(legend.position="none",
        axis.title.y=element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))
myplot.m <- ggplot() +
  geom_boxplot(data=stats1, aes(x=pcgroup,y=m, fill=period)) +
  geom_point(data=mystats, aes(x=pcgroup,y=m, col=period), size=2) +
  geom_point(data=mystats, aes(x=pcgroup,y=m), size=2, pch=1) +
  scale_color_manual(values=c("red","turquoise4")) +
  ggtitle("Modularity") +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))
grid.arrange(myplot.n, myplot.ed, myplot.m, ncol=3, widths=c(1,1,1.5))
```
<br>

# 4: Plot 3-year Networks

Read back in the graph objects and plot them.
```{r}
for(p in myports){
  tmpgraph <- readRDS(here::here(outdir, paste0("igraph_",p,"_3yrAVG_",c,".rds")))
  plot_3yr(tmpgraph, outdir=pngdir)
}
```








