---
title: "Directed Participation Networks - Fig. 3"
author: "Mary Fisher"
date: "10/9/2020"
output: html_document
---


# Description

Graph the directed networks which show flow of Dungeness crab vessels between fisheries in the early season of the 2015 --> early season of 2016. 

For more details on how directed vessel networks are built and interpreted, see the [Guide to Directed Networks](https://github.com/mfisher5/ParticipationNetworks/blob/master/doc/Guide_To_Directed_Networks.md) on Github.

There are several different ways to graph the directed networks:



```{r setup, include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("igraph")) {install.packages("igraph")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("ggraph")) {install.packages("ggraph")}
if(!require("grid")) {install.packages("grid")}
if(!require("ggimage")) {install.packages("ggimage")}
if(!require("gridExtra")) {install.packages("gridExtra")}
```
<br>

And calls the following functions:
```{r}
source(here::here("R","graph_directed_network_early.R"))
source(here::here("R","graph_directed_network_early_coords_image.R"))
source(here::here("R","color_network_nodes.R"))
source(here::here("R","rename_network_nodes.R"))
```
<br>

# User Inputs 

Select your directories.
```{r get_dir}
## location of igraph objects
indir <- 'data/networks/vessel_flow'
```
<br>

Identify the port groups that you would like to produce networks for. The port groups vector should include only those port groups which have associated `igraph` objects. Also list out the full names of the ports, so that names are used instead of codes on the figure.
```{r}
myports <- c("CCA", "ERA", "BGA", "BDA", "SFA", "MNA", "MRA")
portnames <- c("Crescent City", "Eureka", "Fort Bragg", "Bodega Bay", "San Francisco", "Monterey", "Morro Bay")
```
<br>

Provide the name of the `.png` for the Dungeness crab network node.
```{r}
img_file <- here::here("input_data","crab_resized.png")
```
```{r include=FALSE, eval=FALSE}
## when writing out the plot, it may distort the image. Use this code to adjust image aspect ##
## the image_aspect argument in the function only works within R, not when saving to a file ##
## this is a weird thing that needs to be solved in the geom_image() source code ##
img <- imager::load.image(here::here("input_data","crab.png"))
img_resize <- resize(img,round(width(img)/2.1),round(height(img)))
save.image(img_resize, here::here("input_data","crab_resized.png"), quality = 0.7)
```




# Plot Directed Networks

A quick for-loop to save all of the graph objects into a list.
```{r}
graphs_list <- list()
list_names <- c()
i=1
for(p in myports){
  for(s in c("large","small")){
    tmpgraph <- readRDS(here::here(indir,paste0(p,"_closure_",s,"_v7_newVL.rds")))
    V(tmpgraph)$port = portnames[which(myports == p)]
    V(tmpgraph)$vsize = s
    tmpgraph <- rename_vertices(tmpgraph)
    graphs_list[[i]] <- tmpgraph
    list_names[i] <- paste0(p,"_",s)
    i = i + 1
  }
}
names(graphs_list) <- list_names
```
<br>

The function `plot_directed_early` can produce the different plot types listed at the beginning of the script. The following sections use different function calls to produce each plot type.

## D. crab node labeled

This version of the plotting function applies the default vertical node locations produced by `ggraph`, and labels the Dungeness crab node on the graph with text.

Apply the plotting function to all of the graphs in the list.
```{r}
myplots <- lapply(graphs_list, plot_directed_early, crab_image=FALSE, x_coords=FALSE)
```
<br>

Create Figure 3, with column and x axis titles.
```{r eval=FALSE}
col.titles=c("Large","Small")
ylabel="2015                                      2016"
png(here::here("results/figures","Figure3_base.png"), width=1800,height=3000,res=200)
grid.arrange(grobs=lapply(c(1,2), function(i) {
  arrangeGrob(grobs=myplots[seq(i,14,by=2)], top=textGrob(col.titles[i],gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
}), ncol=2)
dev.off()
```
<br>

## D. crab node icon

This version of the plotting function applies the default vertical node locations produced by `ggraph`, and labels the Dungeness crab node on the graph with an image of a crab outline. 

Apply the plotting function to all of the graphs in the list.
```{r}
myplots <- lapply(graphs_list, plot_directed_early, crab_image=TRUE, image_object=img_file, image_size=0.1, image_aspect=1.2, image_offset=0.4, image_transparent=FALSE)
```
<br>

Create Figure 3, with column and x axis titles.
```{r eval=FALSE}
col.titles=c("Large","Small")
ylabel="2015                                      2016"
png(here::here("results/figures","Figure3_icon.png"), width=1800,height=3000,res=200)
grid.arrange(grobs=lapply(c(1,2), function(i) {
  arrangeGrob(grobs=myplots[seq(i,14,by=2)], top=textGrob(col.titles[i],gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
}), ncol=2)
dev.off()
```



## Fixed node positions

This version of the plotting function manually assigns vertical node locations rather than using the ones produced by `ggraph`; this helps maintain consistency in node positions across different graphs, and prevents node / edge overlap. 

Apply the plotting function to all of the graphs in the list.
```{r}
myplots <- lapply(graphs_list, plot_directed_early, x_coords=TRUE, crab_image=FALSE, dist_y=1.2)
```
<br>

Create Figure 3, with column and x axis titles.
```{r eval=FALSE}
col.titles=c("Large","Small")
ylabel="2015                                      2016"
png(here::here("results/figures","Figure3_coords.png"), width=1800,height=3000,res=200)
grid.arrange(grobs=lapply(c(1,2), function(i) {
  arrangeGrob(grobs=myplots[seq(i,14,by=2)], top=textGrob(col.titles[i],gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
}), ncol=2)
dev.off()
```



## Fixed node positions, D. crab icon

This is the version of Figure 3 used in Fisher et al. The plotting function manually assigns vertical node locations, and uses the crab outline image to label the Dungeness crab node. While Fisher et al. did not adjust the transparency of the image, it is possible to do so using this function. 

Apply the plotting function to all of the graphs in the list. The plotting function can be called through the generic `plot_directed_early`, or directly. 
```{r}
myplots <- lapply(graphs_list, plot_directed_early_coords_image, dist_y=1.1, image_object=img_file, image_size=0.13, image_aspect=1.2, image_offset=0.3, image_transparent=FALSE)
```
<br>

Create Figure 3, with column and x axis titles.
```{r}
col.titles=c("Large","Small")
ylabel="2015                                      2016"

png(here::here("drafts/figures","Figure3.png"), width=1800,height=2500,res=200)
grid.arrange(grobs=lapply(c(1,2), function(i) {
  arrangeGrob(grobs=myplots[seq(i,14,by=2)], top=textGrob(col.titles[i],gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
}), ncol=2)
dev.off()
```
<br>

### Save Directed Network with Legend

Create a data frame with gear type and corresponding color key. Use simple x/y coordinates to create a legend key with ggplot.
```{r}
gear_legend <- data.frame(gear=c("Pot","Hook & Line","Pot/Hook & Line","Trawl","Shrimp Trawl", "Other"),
           colorkey=c("darkorange2","darkgoldenrod1","tan1","chocolate4","hotpink","plum3"),
           x=rep(1,6),
           y=c(6,5,4,3,2,1))

legend_plot <- ggplot(gear_legend, aes(x=x,y=y)) +
  geom_point(size=8,col=gear_legend$colorkey) +
  geom_point(pch=21, size=8,col='black', fill=NA) +
  geom_text(aes(x=x,y=y,label=gear), hjust=0, vjust=0.5, nudge_x=0.05, size=6) +
  xlim(c(1,2)) +
  theme_void()
legend_plot
```
<br>

```{r include=FALSE, eval=FALSE}
## Save the legend ##
png(here::here('results','figures','Fig3_legend.png'), res=200,width=900,height=700)
legend_plot
dev.off()
```
<br>

Split up the large and small vessel plots (n=7 each), then save them each in a single column with vessel size at the top, and the 2015/2016 labels at the bottom (saved as a "grob") 
```{r}
lg_plots <- list()
sm_plots <- list()
s=1
l=1
for(i in seq(1,14)){
  if(i%%2==0){
    sm_plots[[s]] <- myplots[[i]]
    s=s+1
  } else{
    lg_plots[[l]] <- myplots[[i]]
    l=l+1
  }
}

ylabel="2015                  2016"
grob1 <- arrangeGrob(grobs=lg_plots, top=textGrob("Large",gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
grob2 <- arrangeGrob(grobs=sm_plots, top=textGrob("Small",gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
```
<br>

Save the legend plot into a column as well, with blank plots beneath it (saved as "grob") and a blank title at the top to adjust the start of the legend to the top of the network graphs.
```{r}
blank_plot <- ggplot() + theme_void()
grob3 <- arrangeGrob(grobs=list(legend_plot,blank_plot,blank_plot,blank_plot,blank_plot), top=textGrob("",gp=gpar(fontsize=22)), ncol=1)
```
<br>

Save the 3 columns into a single image, with the legend to the right of the network graphs.
```{r}
png(here::here("drafts/figures","Figure3_legend.png"), width=2200,height=2500,res=200)
grid.arrange(grobs = list(grob1,grob2,grob3), ncol=3, widths=c(1,1,0.5))
dev.off()
```
<br>



---

Script run time: `r Sys.time() - script_start_time`

<br>



