---
title: "Explore Node Statistics"
author: "M. Fisher"
date: "Written Jan 3, 2022. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---


# Description

Check for a correlation between network size (node count) and node statistics. The three core network metrics are: 

1. Degree (normalized)

2. Node strength

3. Eigenvector centrality

4. Node size (importance)


*based on "Explore Network Stats" script written Oct. 25, 2019 for Fisher et al.*


# Set up

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>



This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("gridExtra")) {install.packages("gridExtra")}
if(!require("cowplot")) {install.packages("cowplot")}
if(!require("PNWColors")) {devtools::install_github("jakelawler/PNWColors"); library(PNWColors)}
```
<br>


Select your directories.
```{r}
## directory with network statistics 
indir <- 'results/statistics'

## directory to save figures
pngdir <- 'results/figures'
```
<br>

What are the names of the `.csv` files with the network statistics that you want to compare? Would not suggest listing more than 4. Files should all be in the input directory (`indir`) specified above.
```{r}
filenames <- c("VesselLevelNodeStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv") 

#"VesselLevelNetworkStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv")

```
<br>

Specify which network metrics to retain from the full data set.

```{r}
node_metrics <- c('node_strength','eigen','deg_norm')
```
<br>

Specify which fisheries to retain from the full data set.
```{r}
myfisheries <- c('Shellfish', 'Echinoderms', 'Tuna', 'Non-DTS groundfish')
```
<br>

Specify which port groups to retain from the full data set.
```{r}
myports <- c("Fort Bragg", "Bodega Bay",  "San Francisco", "Morro Bay",
             "North WA Coast", "Puget Sound", "WA Coast","Astoria","Tillamook","Newport","Coos Bay")
```
<br>


# Data

Network metrics from different datasets. 
```{r}
for(i in seq(1,length(filenames))){
  tmpdat <- read.csv(here::here(indir, filenames[i])) %>%
    dplyr::select(pcgroup, year, metier.cm, all_of(node_metrics)) %>%
    filter(pcgroup %in% myports)
  if(i==1){
    dat <- tmpdat
  } else{ dat <- bind_rows(dat,tmpdat)}
}

head(dat)
```
<br>

Get node counts (network size) for each port group / year
```{r}
network_sizes <- dat %>%
  dplyr::select(pcgroup,year,metier.cm) %>%
  distinct() %>%
  group_by(pcgroup,year) %>%
  summarize(N=n())
```
<br>

Select only port groups / metiers of interest, them add on network size
```{r}
 dat <- dat %>%
    filter(metier.cm %in% myfisheries) %>%
  left_join(network_sizes, by=c("pcgroup","year"))
```
<br>

Check out the distributions of the metrics.
```{r echo=FALSE}
dat_long <- dat %>%
  pivot_longer(cols=c(all_of(node_metrics)), names_to = "metric", values_to = "value")

ggplot(dat_long,aes(x=value)) +
  geom_histogram() +
  facet_wrap(~metric, scales="free_y") +
  labs(x="value",y="") +
  theme_bw()
```
<br>


# Spearman Rank Correlation Test

This function runs the correlation test on the data for each metric, by port group / fishery. It takes a data frame with 4 columns: period, region, network size (N) and the statistics of interest.

```{r}
size_spearman <- function(mydata, metric){
  output <- data.frame(statistic = as.character(),
                       fishery = as.character(),
                       sp.rho = as.numeric(),
                       p.value = as.numeric())
    for(m in myfisheries){
      tmpdat <- filter(mydata, metier.cm==m)
      my.cor <- cor.test(tmpdat[,2], tmpdat[,3], method=("spearman"))
      output <- rbind(output,
                      data.frame(statistic=metric,fishery=m,sp.rho=my.cor$estimate,p.value=my.cor$p.value))
    }
  return(output)
}
```
<br>

### By fishery

Save an empty data frame
```{r}
spearman_dat <- data.frame(statistic = as.character(),
                           fishery = as.character(),
                           sp.rho = as.numeric(),
                           p.value = as.numeric())
```
<br>

Run Spearman Rank for each metric and append the output to the empty data frame.
```{r}
for(metric in node_metrics){
  tmpstats <- dat %>%
    dplyr::select(metier.cm, N, all_of(metric))
  tmp_out <- size_spearman(mydata=tmpstats,metric=metric)
  spearman_dat <- rbind(spearman_dat,
                        tmp_out)
}

```
<br>

Apply bonferroni correction (3 tests per metric) -- which are significant ($\alpha$=0.05)
```{r}
spearman_dat <- spearman_dat %>%
  mutate(p.value.bonferroni=p.value*4) %>%
  pivot_longer(cols=c("p.value","p.value.bonferroni"), names_to="test_type",values_to="pval")
spearman_dat[which(spearman_dat$pval<0.05),]
```
<br>

Results:
```{r echo=FALSE}
ggplot(data=spearman_dat, aes(x=statistic,y=pval, fill=fishery)) +
  geom_col(position="dodge") + 
  geom_hline(aes(yintercept=0.05)) +
  facet_grid(cols=vars(test_type)) +
  labs(x="Node Metric",y="Spearman p-value") +
  theme_bw()
```
<br>

### Across fisheries


Save an empty data frame
```{r}
spearman_dat_all <- data.frame(statistic = as.character(),
                           fishery = as.character(),
                           sp.rho = as.numeric(),
                           p.value = as.numeric())
```
<br>

Run Spearman Rank for each metric and apply a Bonferroni correction-- which are significant ($\alpha$=0.05)
```{r}
for(metric in node_metrics){
  tmpstats <- dat %>%
    dplyr::select(metier.cm, N, all_of(metric))
  my.cor <- cor.test(tmpstats[,2], tmpstats[,3], method=("spearman"))
  tmp_out <- data.frame(statistic=metric,sp.rho=my.cor$estimate,p.value=my.cor$p.value)
  spearman_dat_all <- rbind(spearman_dat_all,
                            tmp_out)
}

# apply correction
spearman_dat_all <- spearman_dat_all %>%
  mutate(p.value.bonferroni=p.value*3) %>%
  pivot_longer(cols=c("p.value","p.value.bonferroni"), names_to="test_type",values_to="pval")
spearman_dat_all[which(spearman_dat_all$pval<0.05),]

```
<br>

Results:
```{r echo=FALSE}
ggplot(data=spearman_dat_all, aes(x=statistic,y=pval)) +
  geom_col(position="dodge") + 
  geom_hline(aes(yintercept=0.05)) +
  facet_grid(cols=vars(test_type)) +
  labs(x="Node Metric",y="Spearman p-value") +
  theme_bw()
```






