---
title: "Compare Timeseries of Network Stats Across Port Groups and to Other ESR Indicators"
author: "M. Fisher, J.F. Samhouri"
date: "Written Dec 30, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

This document graphs network metrics over time for the US West Coast, and compares metrics across port groups and to other indicators in the ESR.


The core network metric of interest on its own is Edge Density (alt: Mean Degree).

The two core network metrics of interest for comparison to the Habitat Compression Index are: 

1. Node Strength

2. Eigen Vector Centrality (aka Importance).

Initially, we consider the above 2 metrics for salmon and non-DTS groundfish for all ports.

**Note that many network metrics are sensitive to network size, which will vary over the time series.** Network size has been incorporated into each graph as ___. Use the timeseries graphs for visual comparison of different network forms, but not for drawing major conclusions around network change / sensitivity. 



Make sure to define filenames and networkstats_filename so that the date is associated with the most recent files


# Setup

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("foreign")) {install.packages("foreign")}
if(!require("lubridate")) {install.packages("lubridate")}
# if(!require("facetscales")) {devtools::install_github("zeehio/facetscales"); library(facetscales)} # this package was archived by the developer in 2022
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
if(!require("ggpubr")) {install.packages("ggpubr")}
if(!require("cowplot")) {install.packages("cowplot")}
if(!require("viridis")) {install.packages("viridis")}
if(!require("PNWColors")) {devtools::install_github("jakelawler/PNWColors"); library(PNWColors)}
```
<br>

Select your directories.
```{r}
## directory with network statistics 
indir <- 'results/statistics'

## directory with other input data
other_indir <- 'data/input'

## directory to save figures
pngdir <- 'results/figures'
```
<br>

What are the names of the `.csv` files with the network statistics that you want to compare? Would not suggest listing more than 4. Files should all be in the input directory (`indir`) specified above.


```{r}
filenames <- "VesselLevelNodeStats_2004_2025_21pcgroups_totalrev7873.43337267464_indivrev787.343337267464_connectivity_rmCrab_10pContribution_2026-01-15.csv"
networkstats_filename <- c("VesselLevelNetworkStats_2004_2025_21pcgroups_totalrev7873.43337267464_indivrev787.343337267464_connectivity_rmCrab_10pContribution_2026-01-15.csv")
# hci_filename <- 'ei_hci_rgn1_M.csv'
port_filename <- 'iopac_conversion_table.csv'
```
<br>

Provide descriptive names to differentiate the datasets drawn from the files listed above. 
```{r}
dataset_names <- c("Node_Stats")
dataset_names2 <- c("Network_Stats")
```
<br>

Specify which network metrics to retain from the full data set.
```{r}
node_metrics <- c('node_strength','eigen')
network_metrics <- c('ed', 'N') # ,'ed_weighted'
```
<br>

Specify which fisheries to retain from the full data set.
```{r}
myfisheries <- c('Salmon','Non-DTS groundfish', 'DTS groundfish')
```
<br>

<!-- Specify which port groups to retain from the full data set. -->
<!-- ```{r} -->
<!-- myports <- c("North WA Coast", "Puget Sound", "WA Coast", "Astoria", "Tillamook", "Newport", "Coos Bay") # drop "Brookings" for now to correspond with HCI Region 4 and because 7 colors kind of max things out? -->
<!-- ``` -->
<!-- <br> -->



# Read in Data


Network metrics from different datasets. 
```{r}
for(i in seq(1,length(filenames))){
  tmpdat <- read.csv(here::here(indir, filenames[i])) %>%
    dplyr::select(pcgroup, year, metier.cm, all_of(node_metrics)) %>%
    #filter(pcgroup %in% myports) %>%
    filter(metier.cm %in% myfisheries) %>%
    mutate(dataset=dataset_names[i])
  if(i==1){
    dat <- tmpdat
  } else{ dat <- bind_rows(dat,tmpdat)}
}

glimpse(dat)

for(i in seq(1,length(networkstats_filename))){
  tmpdat2 <- read.csv(here::here(indir, networkstats_filename[i])) %>%
    dplyr::select(pcgroup, y, all_of(network_metrics)) %>%
    #filter(pcgroup %in% myports) %>%
    #filter(metier.cm %in% myfisheries) %>%
    mutate(dataset=dataset_names2[i])
  if(i==1){
    dat2 <- tmpdat2
  } else{ dat2 <- bind_rows(dat2,tmpdat2)}
}

glimpse(dat2)

# 01-06-22: note jameal updated OSD, SD, and SDA ports to have the same lat long as OCN. Amanda P has made the same fix. Also reassigned MCR to the Eureka IOPAC port group,  Amanda P agreed with this change
iopac_lookup <- read_csv(here::here(other_indir, port_filename))

```
<br>

Check to make sure that all of the datasets were read in. 
```{r echo=FALSE}
with(dat,table(dataset,pcgroup))
```
<br>

Pivot longer to graph multiple metrics at once.
```{r}
dat_long <- dat %>%
  pivot_longer(cols=c(all_of(node_metrics)), names_to = "metric", values_to = "value")

dat2_long <- dat2 %>%
  pivot_longer(cols=c(all_of(network_metrics)), names_to = "metric", values_to = "value")
```
<br>

Clean up metric names.
```{r}
dat_long$metric <- recode(dat_long$metric, 
                          node_strength = "Node Strength",
                          eigen = "Importance")
#dat_long$metric <- factor(dat_long$metric, levels=c("Size", "Edge Density","Mean Degree", "Central.", "Central. (UnW)", "Mod.", "Mod. (UnW)"))

#dat_long$pcgroup <- factor(dat_long$pcgroup, levels=myports)

colnames(dat2) <- c(
  'Port Group','Year','Edge Density','N','dataset'
)

dat2_long$metric <- recode(dat2_long$metric, 
                          'ed' = 'edge density',
                          'N' = 'number of fisheries')

```
<br>

convert to ESR format
(Drop Other Coastal Washington & Unknown Ports & Columbia River)
```{r}

# format for ESR with the following columns:  
# (1) year - this can be just year or a date, preferably in yyyy or yyyy-mm-dd format. 
# (2) index - this column contains the data you want to plot
# (3) timeseries - title for the time series plot.  
# (4) metric - how the data were measured; essentially the label for the y-axis
# (5) SE or SD if you have it
# 
# NAs or blanks will not be plotted
# Zeros will be plotted
# You don't have to fill in missing years

# note that for each year, it reflects november of (year) through november of (year + 1). specifically week 46 of year 1 through week 45 of year 2

dat_esr1 <- dat_long %>%
  mutate(metric = paste(metier.cm, metric)) %>%
  dplyr:: select(year, value, metric, pcgroup) %>%
  rename(
    index = value,
    timeseries = pcgroup 
  )

dat_esr2 <- dat2_long %>%
  dplyr:: select(y, value, metric, pcgroup) %>%
  rename(
    year = y,
    index = value,
    timeseries = pcgroup 
  )


# Bind and (Drop Other Coastal Washington & Unknown Ports & Columbia River)
dat_esr <- dat_esr1 %>% bind_rows(dat_esr2) %>% 
  filter( timeseries != "Other Coastal Washington" & timeseries != "Unknown Ports" & timeseries != "Columbia River")

write.csv(dat_esr,here::here(indir, paste0('FPN_ed_n_ns_ec_salmon_nondts_dts_FY2026.csv',"_",Sys.Date())), row.names=FALSE)

```
<br>

<!-- Monthly Habitat Compression Index over 43.5-48N (Region 1). Summarize for spring months only (Apr-Jun). -->
<!-- ```{r} -->
<!-- spring_months <- c(4,5,6) -->

<!-- hci <- read_csv(here::here(other_indir,hci_filename)) %>% -->
<!--   dplyr::select(year, month, data) %>% -->
<!--   filter(year >= 2005) %>% -->
<!--   filter(month %in% spring_months) %>% -->
<!--   group_by(year) %>% -->
<!--   summarise( -->
<!--     annual_mean = mean(data), -->
<!--     .groups = 'drop' -->
<!--   ) -->

<!-- glimpse(hci) -->

<!-- ``` -->
<!-- <br> -->

# 2. Graph Timeseries

## Figure x.Time series of node strength 

... at WA and OR ports from 2004-2025. Node strength was derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` (USD in 2005; following Holland et al. diversification indices) and a minimum revenue cutoff for individual fisheries of `$500` (USD in 2005; following Holland et al. diversification indices).

```{r}
plot_metric="Node Strength"
metric_label="Node Strength"
```
<br>

```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
plotdat_nondtsgfish <- dat_long %>% filter(metric==plot_metric) %>% filter(metier.cm == 'Non-DTS groundfish')
plotdat_dtsgfish <- dat_long %>% filter(metric==plot_metric) %>% filter(metier.cm == 'DTS groundfish')
plotdat_salmon <- dat_long %>% filter(metric==plot_metric) %>% filter(metier.cm == 'Salmon')

# color palette
pal=pnw_palette("Starfish",length(unique(plotdat_gfish$pcgroup))) # or choose a different PNWColors palette

# main metric plot - nondtsgfish
figx_nondtsgfish <- ggplot(plotdat_nondtsgfish, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2), labels=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Node Strength of Non-DTS Groundfish in Fisheries Participation Networks',
       subtitle = '2004-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figx_nondtsgfish

# main metric plot - dtsgfish
figx_dtsgfish <- ggplot(plotdat_dtsgfish, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2), labels=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Node Strength of DTS Groundfish in Fisheries Participation Networks',
       subtitle = '2004-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figx_dtsgfish

# color palette
pal=pnw_palette("Starfish",length(unique(plotdat_salmon$pcgroup))) # or choose a different PNWColors palette

# main metric plot - salmon
figx_salmon <- ggplot(plotdat_salmon, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat_salmon$year), max(plotdat_salmon$year), by=2), labels=seq(min(plotdat_salmon$year), max(plotdat_salmon$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Node Strength of Salmon in Fisheries Participation Networks',
       subtitle = '2004-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figx_salmon

```
<br>
<br>
<br>

Save plot as a png, with filename: Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2004-2025
```{r echo=FALSE}
png(here::here(pngdir, "Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2004-2025.png"),
    res=200,height=800,width=1700)
figx_nondtsgfish
dev.off()
```
<br>

Save plot as a png, with filename: Node Strength of DTS Groundfish in Fisheries Participation Networks 2004-2025
```{r echo=FALSE}
png(here::here(pngdir, "Node Strength of DTS Groundfish in Fisheries Participation Networks 2004-2025.png"),
    res=200,height=800,width=1700)
figx_dtsgfish
dev.off()
```


Save plot as a png, with filename: Node Strength of Salmon in Fisheries Participation Networks 2004-2024
```{r echo=FALSE}
png(here::here(pngdir, "Node Strength of Salmon in Fisheries Participation Networks 2004-2025.png"),
    res=200,height=800,width=1700)
figx_salmon
dev.off()
```
<br>

## Figure y.Time series of importance

... at WA and OR ports from 2004-2025. Importance (eigenvalue centrality) was derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` (USD in 2005; following Holland et al. diversification indices) and a minimum revenue cutoff for individual fisheries of `$500` (USD in 2005; following Holland et al. diversification indices).

```{r}
plot_metric="Importance"
metric_label="Importance"
```
<br>

```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
plotdat_nondtsgfish <- dat_long %>% filter(metric==plot_metric) %>% filter(metier.cm == 'Non-DTS groundfish')

plotdat_dtsgfish <- dat_long %>% filter(metric==plot_metric) %>% filter(metier.cm == 'DTS groundfish')


plotdat_salmon <- dat_long %>% filter(metric==plot_metric) %>% filter(metier.cm == 'Salmon')

# color palette
pal=pnw_palette("Starfish",length(unique(plotdat_gfish$pcgroup))) # or choose a different PNWColors palette

# main metric plot - non-DTS gfish
figy_nondtsgfish <- ggplot(plotdat_nondtsgfish, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2), labels=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Importance of Non-DTS Groundfish in Fisheries Participation Networks',
       subtitle = '2004-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figy_nondtsgfish

# main metric plot - DTS gfish
figy_dtsgfish <- ggplot(plotdat_dtsgfish, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2), labels=seq(min(plotdat_gfish$year), max(plotdat_gfish$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Importance of Non-DTS Groundfish in Fisheries Participation Networks',
       subtitle = '2004-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figy_dtsgfish

# color palette
pal=pnw_palette("Starfish",length(unique(plotdat_salmon$pcgroup))) # or choose a different PNWColors palette

# main metric plot - salmon
figy_salmon <- ggplot(plotdat_salmon, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat_salmon$year), max(plotdat_salmon$year), by=2), labels=seq(min(plotdat_salmon$year), max(plotdat_salmon$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Importance of Salmon in Fisheries Participation Networks',
       subtitle = '2004-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figy_salmon
```
<br>
<br>
<br>

Save plot as a png, with filename: Importance of Non-DTS Groundfish in Fisheries Participation Networks 2004-2025
```{r echo=FALSE}
png(here::here(pngdir, "Importance of Non-DTS Groundfish in Fisheries Participation Networks 2004-2025.png"),
    res=200,height=800,width=1700)
figy_nondtsgfish
dev.off()
```
<br>

Save plot as a png, with filename: Importance of DTS Groundfish in Fisheries Participation Networks 2004-2025
```{r echo=FALSE}
png(here::here(pngdir, "Importance of DTS Groundfish in Fisheries Participation Networks 2004-2025.png"),
    res=200,height=800,width=1700)
figy_dtsgfish
dev.off()
```

Save plot as a png, with filename: Importance of Salmon in Fisheries Participation Networks 2004-2025
```{r echo=FALSE}
png(here::here(pngdir, "Importance of Salmon in Fisheries Participation Networks 2004-2025.png"),
    res=200,height=800,width=1700)
figy_salmon
dev.off()
```
<br>



<!-- ## Figure z.Time series of HCI for Region 1  -->

<!-- ```{r echo=FALSE, fig.width=10} -->

<!-- # main metric plot -->
<!-- figz <- ggplot(hci, aes(x=year, y=annual_mean)) +  -->
<!--   geom_point(size=2) + geom_line() + -->
<!--   scale_x_continuous(breaks=seq(min(plotdat$year), max(plotdat$year), by=2), labels=seq(min(plotdat$year), max(plotdat$year), by=2)) + -->
<!--   ylim(c(0,1)) + -->
<!--   labs(title='Habitat Compression Index for Region 1, 43.5-48N', -->
<!--        subtitle = 'Apr-Jun 2005-2021', -->
<!--        x = "Year", -->
<!--        y = 'Habitat Area\n(fraction below monthly threshold)' -->
<!--        ) + -->
<!--   theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5), -->
<!--                      axis.title.x=element_text(size=13), -->
<!--                      axis.title.y=element_text(size=13), -->
<!--                      strip.text=element_text(size=13), -->
<!--                      legend.title=element_text(size=13),legend.text=element_text(size=13), -->
<!--                      panel.grid.major=element_blank(), -->
<!--                      panel.grid.minor=element_blank()) +  -->
<!--   theme(legend.position = 'none') -->
<!-- figz -->
<!-- ``` -->
<!-- <br> -->
<!-- <br> -->
<!-- <br> -->

<!-- Save plot as a png, with filename: Spring Habitat Compression Index for Region 1, 43.5-48N, 2005-2021 -->
<!-- ```{r echo=FALSE} -->
<!-- png(here::here(pngdir, "Spring Habitat Compression Index for Region 1, 43.5-48N, 2005-2021.png"), -->
<!--     res=200,height=800,width=1700) -->
<!-- figz -->
<!-- dev.off() -->
<!-- ``` -->
<!-- <br> -->



## Figure aa. Time series of edge density

... at all ports from 2005-2025. Edge density was derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` (USD in 2005; following Holland et al. diversification indices) and a minimum revenue cutoff for individual fisheries of `$500` (USD in 2005; following Holland et al. diversification indices).

(Drop Other Coastal Washington & Unknown Ports & Columbia River)

```{r}
plot_metric="Edge Density"
metric_label="Edge Density"
```
<br>

```{r echo=FALSE, fig.width=10}

iopac_mean_lat <- iopac_lookup %>% dplyr::select(IOPAC, PACFIN_LAT) %>% group_by(IOPAC) %>% summarise(mean_lat = mean(PACFIN_LAT, na.rm=TRUE)) %>% ungroup() %>% arrange(-mean_lat) %>% dplyr::select(IOPAC) %>% pull() 

# allports_sorted <- dat2 %>% left_join(iopac_mean_lat, by = c('Port Group' = 'IOPAC')) %>% dplyr::select(`Port Group`, mean_lat) %>% arrange(-mean_lat) %>% select(`Port Group`) %>% pull() # use arrange(-lat) if coord_flip() is NOT used

dat2$`Port Group` <- factor(dat2$`Port Group`, levels=iopac_mean_lat)

# main metric plot
fig_ed <- ggplot(dat2 %>% filter( `Port Group` != "Other Coastal Washington" & `Port Group` != "Unknown Ports" & `Port Group` != "Columbia River"), 
                 aes(x=Year+1, y=`Edge Density`, colour=`Port Group`)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(vars(`Port Group`),labeller = label_wrap_gen(16)) +
  scale_colour_viridis(discrete = TRUE) +
  scale_x_continuous(breaks=seq(min(dat2$Year)+1, max(dat2$Year)+1, by=2), labels=seq(min(dat2$Year)+1, max(dat2$Year)+1, by=2)) +
  ylim(c(0,1)) +
  labs(title='Edge Density in Fisheries Participation Networks',
       subtitle = '2005-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
fig_ed
```
<br>
<br>
<br>

Save plot as a png, with filename: Edge Density in Fisheries Participation Networks 2005-2024
```{r echo=FALSE}
png(here::here(pngdir, "Edge Density in Fisheries Participation Networks 2005-2025","_",Sys.Date(),".png"),
    res=200,height=1500,width=1700)
fig_ed
dev.off()
```
<br>

## Figure bb. Time series of number of nodes

... at all ports from 2004-2025. The number of nodes was derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` (USD in 2005; following Holland et al. diversification indices) and a minimum revenue cutoff for individual fisheries of `$500` (USD in 2005; following Holland et al. diversification indices). To avoid inclusion of rare or minimal fishing activity, a given fishery must contribute to at least 10% of a vessel's annual revenue to be included in the network data.

(Drop Other Coastal Washington & Unknown Ports & Columbia River)

```{r}
plot_metric="N"
metric_label="Number of fisheries"
```
<br>

```{r echo=FALSE, fig.width=10}

# iopac_mean_lat <- iopac_lookup %>% dplyr::select(IOPAC, PACFIN_LAT) %>% group_by(IOPAC) %>% summarise(mean_lat = mean(PACFIN_LAT, na.rm=TRUE)) %>% ungroup() %>% arrange(-mean_lat) %>% dplyr::select(IOPAC) %>% pull() 

# allports_sorted <- dat2 %>% left_join(iopac_mean_lat, by = c('Port Group' = 'IOPAC')) %>% dplyr::select(`Port Group`, mean_lat) %>% arrange(-mean_lat) %>% select(`Port Group`) %>% pull() # use arrange(-lat) if coord_flip() is NOT used

# dat2$`Port Group` <- factor(dat2$`Port Group`, levels=iopac_mean_lat)

# main metric plot
fig_n <- ggplot(dat2 %>% filter( `Port Group` != "Other Coastal Washington" & `Port Group` != "Unknown Ports" & `Port Group` != "Columbia River"), 
                 aes(x=Year+1, y=N, colour=`Port Group`)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(vars(`Port Group`),labeller = label_wrap_gen(16)) +
  scale_colour_viridis(discrete = TRUE) +
  scale_x_continuous(breaks=seq(min(dat2$Year), max(dat2$Year), by=2), labels=seq(min(dat2$Year), max(dat2$Year), by=2)) +
  #ylim(c(0,1)) +
  labs(title='Number of Fisheries in Fisheries Participation Networks',
       subtitle = '2004-2025',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
fig_n
```
<br>
<br>
<br>

Save plot as a png, with filename: Nodes in Fisheries Participation Networks 2004-2024
```{r echo=FALSE}
png(here::here(pngdir, "Number of Nodes in Fisheries Participation Networks 2004-2025","_",Sys.Date(),".png"),
    res=200,height=1500,width=1700)
fig_n
dev.off()
```
<br>







