---
title: "Summarize WDFW fish tickets isportderived"
author: "Jameal Samhouri"
date: "Written 2025-02-04 Last Run `r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
fontsize: 11pt
---

# PacFIN query

Make sure to update the definition of `latestpull` with the date on which PacFIN was queried, and `min_crab_year` and `max_crab_year` to define the years reported for the ESR.

<!-- SELECT * -->

<!--     FROM PACFIN_MARTS.COMPREHENSIVE_FT cft -->

<!--          LEFT JOIN PACFIN_MARTS.WDFW_FT_PORT_DERIVED_V ftpd -->

<!--              ON     cft.pacfin_year = ftpd.landing_year -->

<!--                 AND cft.ftid = ftpd.ftid -->

<!--                 AND cft.agency_code = ftpd.agency_code -->

<!--    WHERE cft.pacfin_year BETWEEN 2003 AND 2025 AND cft.agency_code = 'W' -->

<!-- ORDER BY cft.landing_year DESC; -->

# Load libraries and set directories
```{r setup}

if(!require("here")) {install.packages("here")}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("formattable")) {install.packages("formattable")}
if(!require("kableExtra")) {install.packages("kableExtra")}
if(!require("scales")) {install.packages("scales")}
if(!require("dplyr")) {install.packages("dplyr")}
if(!require("tictoc")) {install.packages("tictoc")}

library(here)
library(tidyverse)
library(lubridate)
library(formattable)
library(kableExtra)
library(dplyr)
library(scales)
library(tictoc)

# raw PacFIN fish ticket data file location
indir2 <- "/Users/jameal.samhouri/Documents/CONFIDENTIAL DATA/PacFIN Tickets/"

# output directory for confidential data
out_conf_dir <- "/Users/jameal.samhouri/Documents/CONFIDENTIAL DATA/PacFIN Tickets/"

statdir = 'results/statistics'

```

# Read in csv or rds

If you have a new download of fish ticket data and want to clean it up a bit and convert it to RDS, use this chunk.


<!-- ```{r csv2rds} -->

<!-- # NOTE: reading in PacFIN csv's for multiple years is SLOW, and took >1.5hrs on JS macbook -->
<!-- tic() -->

<!-- tmpfile_name <- 'fish tickets 2003-2025 WDFW  port derived 012725.csv' # latest pull from PacFIN 01-27-2026 -->
<!-- tmpfile2 <- read.csv(paste0(indir2,tmpfile_name)) -->
<!-- #glimpse(tmpfile2) -->

<!-- #as.Date(tmpfile2[1,]$LANDING_DATE, "%m/%d/%Y") -->

<!-- # # LANDING_YEAR in more recent data pulls has a time associated with it. convert myfile2$LANDING_DATE to date formats that exclude time. added 12-29-2021 -->
<!-- # update 01-28-2026. LANDING_DATE format now is 21-SEP-25, so updated conversion below -->
<!-- tmpfile2$LANDING_DATE <- as.Date(tmpfile2$LANDING_DATE, "%d-%b-%y") -->
<!-- write_rds(tmpfile2, paste0(out_conf_dir,'WDFW fish tickets 2003-2025_',Sys.Date(),'.rds')) -->
<!-- rm(tmpfile2) -->

<!-- toc() -->
<!-- # 86.833 sec elapsed -->
<!-- ``` -->


If you already have an rds, start here.
```{r readinrds}
tic()
latestpull <- "2026-01-28"
wdfwtix_filename <- paste0('WDFW fish tickets 2003-2025_',latestpull)
wdfw_tix <- read_rds(paste0(out_conf_dir, wdfwtix_filename, '.rds'))
toc()

# 23.801 sec elapsed

#glimpse(wdfw_tix)

```
# Convert to crab years

A crab year is defined as week 46 of year 1 through week 45 of year 2

```{r crab_year}

wdfw_tix1 <- wdfw_tix %>%
  mutate(tdate = ymd(LANDING_DATE)) %>%
  ## create 'calendar week of landing' variable
  mutate(tweek=week(tdate)) %>%
  ## create crab year variable 
  mutate(crab_year = as.numeric(ifelse(tweek > 45, LANDING_YEAR + 1, LANDING_YEAR)))
rm(wdfw_tix)

```

Set the crab years of interest for this ESR.

```{r crab_year_ESR}

min_crab_year <- 2004
max_crab_year <- 2025
year_range <- as.numeric(min_crab_year:max_crab_year)

```

# Summarize

For the ISDERIVED column: 0 means it is not derived, 
1 means it is derived, Null means we do not know

Theresa Tsou (WDFW) 01-27-2026. "I would like to point out again that if a landing port is derived from the license database, it doesn’t always mean it is an error. Most of the time, the address on license reflects their home port, the port they use most often."

```{r sum_isportderived}

wdfw_tix1$IS_PORT_DERIVED <- as.factor(wdfw_tix1$IS_PORT_DERIVED)

portderivedsum <- wdfw_tix1 %>%
  filter(crab_year >= min_crab_year) %>%
  filter(crab_year <= max_crab_year) %>%
  group_by(crab_year, IS_PORT_DERIVED) %>%
  summarise(
    num_tix = n(),
    total_landings = sum(LANDED_WEIGHT_MTONS, na.rm=TRUE), # 01-28-2026. warning: added na.rm = TRUE; implications unclear but it prevents in the total_landings column
    total_afi_revenue = sum(AFI_EXVESSEL_REVENUE, na.rm=TRUE), # 01-28-2026. warning: added na.rm = TRUE; implications unclear but it prevents in the total_afi_revenue column
    .groups = 'drop' #'drop_last'. changed 01-28-2026
  ) %>%
  
  # make sure every crab_year is represented for every level of IS_PORT_DERIVED
  complete(IS_PORT_DERIVED, crab_year = year_range, fill = list(num_tix = 0, total_landings = 0, total_afi_revenue = 0), explicit = FALSE) %>% 
  
  group_by(crab_year) %>%
  mutate(
    percent_tix = scales::percent(num_tix/sum(num_tix), accuracy = 0.1),
    percent_landings = scales::percent(total_landings/sum(total_landings), accuracy = 0.1),
    percent_revenue = scales::percent(total_afi_revenue/sum(total_afi_revenue), accuracy = 0.1)
  )

portderivedsumpercent <- portderivedsum %>%
  dplyr::select(crab_year, IS_PORT_DERIVED, percent_tix, percent_landings, percent_revenue) %>%
  pivot_wider(names_from = IS_PORT_DERIVED, values_from = c(percent_tix, percent_landings, percent_revenue)) %>%
  dplyr::select(crab_year,
          percent_tix_0,
          percent_landings_0,
          percent_revenue_0,
          percent_tix_1,
          percent_landings_1,
          percent_revenue_1,
          percent_tix_NA,
          percent_landings_NA,
          percent_revenue_NA
          )


```

# Make 3 pretty tables and combine

```{r pretty_table}

portNOTderived_out <- portderivedsumpercent %>%
  dplyr::select(
    crab_year,
    percent_tix_0,
    percent_landings_0,
    percent_revenue_0) %>%
  rename(
    Year = 'crab_year',
    `% of Landings Receipts` = percent_tix_0,
    `% of Landings` = percent_landings_0,
    `% of Revenue` = percent_revenue_0)

portderived_out <- portderivedsumpercent %>%
  dplyr::select(
    crab_year,
    percent_tix_1,
    percent_landings_1,
    percent_revenue_1) %>%
  rename(
    Year = 'crab_year',
    `% of Landings Receipts` = percent_tix_1,
    `% of Landings` = percent_landings_1,
    `% of Revenue` = percent_revenue_1)

portderivedUNKNOWN_out <- portderivedsumpercent %>%
  dplyr::select(
    crab_year,
    percent_tix_NA,
    percent_landings_NA,
    percent_revenue_NA) %>%
  rename(
    Year = 'crab_year',
    `% of Landings Receipts` = percent_tix_NA,
    `% of Landings` = percent_landings_NA,
    `% of Revenue` = percent_revenue_NA)





# knitr::kable(
#   portderivedsumprecent,
#   col.names = c(
#     'Year',
#     '% of Landings Receipts',
#     '% of Landings',
#     '% of Revenue'
#   )
# ) %>%
#   add_header_above(
#     c(""=""," " = '% of Landings Receipts', "Port Reported" = 2, " " = '% of Revenue')
#     ) #"Group 2" = 2, "Group 3" = 2))

```

## Port not derived by year: percentages of landings receipts, landings, and revenue 

```{r portNOTderived_out}

portNOTderived_out

```

## Port derived by year: percentages of landings receipts, landings, and revenue 

```{r portderived_out}

portderived_out

```

## Port unknown if derived by year: percentages of landings receipts, landings, and revenue 

```{r portderivedUNKNOWN_out}

portderivedUNKNOWN_out

```

## Write to csv

```{r write}

write_csv(portNOTderived_out,
          here::here(statdir, paste0('portNOTderived_out_FY2026_',Sys.Date(),".csv"))
          )

write_csv(portderived_out,
          here::here(statdir, paste0('portderived_out_FY2026_',Sys.Date(),".csv"))
          )

write_csv(portderivedUNKNOWN_out,
          here::here(statdir, paste0('portderivedUNKNOWN_out_FY2026_',Sys.Date(),".csv"))
          )

```


