---
title: "Test KNN Parameters"
author: "M. Fisher"
date: "Written Jan. 26, 2019. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in Fisher et al.
fontsize: 11pt
---

# Description

KNN identifies the "nearest neighbors" of a given fish ticket based on the Hellinger distance. Different values of `k`, and the use of a maximum acceptable distance value, should be explored before using the KNN to assign metiers, 

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("vegan")) {install.packages("vegan")}
if(!require("purrr")) {install.packages("purrr")}
if(!require("parallelDist")) {install.packages("parallelDist")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
```
<br>

And calls the following functions:
```{r}
```
<br>

# User Inputs 

Select your directories.
```{r get_dir}
## input directory containing fish ticket data without assigned metiers
indir_tix <- 'data/input'

## input directory containing fish tickets from reference years, with infomap-assigned metiers
indir_met <- 'data/intermediate_metiers/reference'
```
<br>

Set the range of values of `k` for the k-nearest neighbor function.
```{r}
kvals <- seq(from=1,to=5,by=1)
```
<br>

Filter the new fish tickets without metiers for commercial tickets?
```{r}
filter_type = TRUE
```
<br>

The script will loop over each of the ports and crab years in these vectors.
```{r}
myports <- c("CCA","ERA","BGA","BDA","SFA","MNA","MRA")
years <- c(2016,2017)
```
<br>

Set these objects according to your user inputs from Script 1
```{r}
## trip comparisons for metier identification based on revenue or lbs
metric <- "revenue"

## which year(s) of fish ticket data were used to identify metiers with infomap?
infomap_years <- c(2010,2011)
```
<br>
