---
title: "Organize Network Data"
author: "E. Fuller, M. Fisher"
date: "Written Feb. 4, 2019. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in Fisher et al.
fontsize: 11pt
---


# Description

Metier identification and assignment is conducted by port group due to computational restrictions. This script will recombine fish tickets across port groups for each crab year. It will also produce a text file for each port group listing all metier IDs and names (species / gear group abbreviations) found in the fish tickets. 

After running this script, you can manually name each metier as you wish to see it represented in network graphs in the text file.

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
```
<br>

# User Inputs 

Select your directories. If following the default directory structure, `indir` and `outdir` will be the same.
```{r get_dir}
## input directory containing fish tickets fully assigned metiers (output from script 04)
indir <- 'data/input/metiers'

## output directory for recombined tickets
outdir <- 'data/input/metiers'
```
<br>

Identify the crab years and port groups covered by your landings data.
```{r}
# years <- seq(2008,2017)
# myports <- c("CCA","ERA","BGA","BDA","SFA","MNA","MRA")
years <- c(2010,2011,2016,2017)
myports <- c("CCA")
```
<br>

Set these objects according to your user inputs from Scripts 1 and 3
```{r}
## the value of `k` for the k-nearest neighbor function
k <- 1

## the reference years first run through infomap
ref_years <- c(2010,2011)
```
<br>


# 1: Combine Data

Save all of the individual fish ticket files as a single data frame.
```{r}
for(p in myports){
  for(y in years){
    if(y %in% ref_years){
      tmptix <- read.csv(here::here(indir, paste0("FishTix_",p,"_crab",y,"_metiers.csv")))
    } else{
      tmptix <- read.csv(here::here(indir, paste0("FishTix_",p,"_crab",y,"_metiers_k",k,"_recoded.csv")))
    }
    if(exists('all_data')){
      all_data <- rbind(all_data,tmptix)
    } else{
      all_data <- tmptix
    } # end if(exists)
  } #end for(y)
} #end for(p)
```
<br>


# 2: Write out data

## 2.1. Fish tickets

Keep in mind that metier IDs / names have not been changed, so there will be duplicates between port groups.
```{r}
for(y in years){
  yearly_tix <- filter(all_data, crab_year==y)
  if(y %in% ref_years){
    write.csv(yearly_tix, here::here(outdir, paste0('FishTix_crab', y, '_metiers.csv')),row.names=FALSE)
  } else{
      write.csv(yearly_tix, here::here(outdir, paste0('FishTix_crab', y, '_metiers_k1_recoded.csv')),row.names=FALSE)
    }
  message("crab year ",y, " contains landings data for port groups: ", paste0(unique(yearly_tix$pcgroup), collapse=", "), ".\n")
}
```
<br>

## 2.2. Metier keys

For each port group, get a data frame with all unique metier IDs / names. Then write to file.

*When splitting the metier name (abbreviation) into species and gear group, mixed-species metiers may generate a warning message about 'additional pieces'. Use this to identify metiers which may require customized `metier.spid` column entries or names.*

```{r}
for(i in seq(1,length(myports))){
  pcgroup_metiers <- filter(all_data,pcgroup==myports[i]) %>%
    dplyr::select(metier.num,metier.name) %>%
    distinct() %>%
    separate(col=metier.name, into=c('metier.spid','metier.gear'), sep='_', remove=FALSE) %>%
    rename(metier.abbrev=metier.name) %>%
    mutate(metier.name=NA)
  write_delim(pcgroup_metiers, here::here(outdir,paste0(myports[i], '_metier_names.txt')), delim='\t')
}
```

For the remaining scripts that generate and plot `igraph` network objects, make sure that you have manually replaced 'NAs' in the `metier.name` column with a customized name for each metier. The text files should include the following columns: 

| metier.num | metier.abbrev |	metier.spid |	metier.gear |	metier.name |
| :--------: | :-----------: | :----------: | :---------: | :---------- |
|     1      |    DCRB_POT   |     DCRB     |    POT      |   D. crab   |
|     2      |    BLCK_HKL   |     BLCK     |    HKL      |  RockLing   |




