---
title: "Process Fish Tickets for Participation Networks"
author: M. Fisher, J.F. Samhouri
date: "Written Dec. 7, 2020. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---


# Description

This code processes raw [PacFIN](www.pacfin.pmsfc.org) fish tickets, which will then be used to generate participation networks. It is designed to write out a separate file for each year of input data, by calendar and crab year.

This version of the script will replace PacFIN's nominal species IDs with the equivalent regular species IDs. For more on nominal IDs, see the [PacFIN FAQs page](https://pacfin.psmfc.org/data/faqs/).


Additional Resources: 
* [PMSC Fleet Report](https://www.psmfc.org/efin/docs/fleetreport.pdf), esp Table 8
* [PacFIN Column names key](https://pacfin.psmfc.org//wp-content/uploads/2016/06/PacFIN_Comprehensive_Fish_Tickets.pdf)

<br>
<br>

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!require("here")) {install.packages("here")}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("foreign")) {install.packages("foreign")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("magrittr")) {install.packages("magrittr")}

## start time for full script
script_start_time <- Sys.time()

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```


This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(foreign)
library(lubridate)
```
<br>

And calls the following functions. 
```{r}
# source(here::here('R','get_vessel_length.R'))
# source(here::here('R','lengths_2yr_median.R'))
# source(here::here('R','get_mode.R'))
# source(here::here('R','lengths_5yr_registration.R'))
```
<br>

# User Inputs 

Select your directories. 
```{r get_dir}
## raw PacFIN fish ticket data file location
indir1 <- "/Users/jameal.samhouri/Documents/Koehn et al. cc vulnerability/Data/"
indir2 <- "/Users/jameal.samhouri/Documents/CCIEA Networks/raw/"

## output directory for confidential data
out_conf_dir <- "/Users/jameal.samhouri/Documents/CCIEA Networks/processed/"

## output directory for processed fish tickets. not confidential
processdir <- "data/input"
```
<br>

Specify file name of the raw fish ticket data file and supporting files with species and port grouping keys. This script takes a flexible input file name, but will write out hard-coded file names for consistency through the rest of the analysis.
```{r get_filenames}
# will need to glue 2004-2018 data to 2019-2020 data

myfile1 <- "pacfin_compiled_2004thru2018.rds"
myfile2 <- "fish tickets 2019-2020.csv"

myfile3 <- "iopac_conversion_table.csv"
myfile4 <- "species_groupings_spgrpn2.csv" # originally from first tab of species_groupings.xlsx from Dan Holland

```
<br>


Set a few more user inputs to process the fish tickets
<br>
*If you decide to write out the processed fish ticket data using crab years as time increments (`crab_year=TRUE`), be sure to supply all of the necessary calendar years. For example, if you are processing fish tickets for the 2014 crab year, you will need to read in raw PacFIN fish ticket data for the calendar years 2013 and 2014; if you are processing fish tickets for the 2014 and 2015 crab years, you will need to read in raw PacFIN fish ticket data for calendar years 2014,2015,2016.*
```{r inputs}
## do you want to write out the fish ticket data as crab years? [TRUE/FALSE]
crab_year = TRUE

## if crab_year = TRUE, which crab years of fish tickets are you processing? [numeric vector]
crab_years <- seq(2004,2019)

## which years of fish tickets are you processing? [numeric vector]
years <- seq(2004,2020)

## do you want to write out the average price per pounds calculated from the data? [TRUE/FALSE]
write_ppp = FALSE

## do you want to only get vessel lengths for Dungeness crab vessels (`dcrb_only` set to TRUE)?
dcrb_only <- FALSE
```
<br>
<br>


# 1: Edit PacFIN Data

## 1.1. Read, Glue, Subset data

NOTE: you can skip reading in the individual raw data files if "pacfin_compiled_2004thru2020.rds" already exists.

This should be a `.rds` or a `.csv` file containing raw PacFIN data.
```{r rawdata}
rawdat <- read_rds(paste0(out_conf_dir,"pacfin_compiled_2004thru2020.rds"))
colnames(rawdat)

# rawdat1 <- read_rds(paste0(indir1,myfile1)) %>%
#   filter(LANDING_YEAR %in%  years)
# print("2004-2018 fish ticket data column names")
# colnames(rawdat1)
# 
# rawdat2 <- read.csv(paste0(indir2,myfile2)) %>%
#   filter(LANDING_YEAR %in%  years)
# print("2019-2020 fish ticket data column names")
# colnames(rawdat2)

port_grps <- read_csv(here::here(processdir,myfile3))
sp_grps <- read_csv(here::here(processdir,myfile4))

```
<br>

Glue fish ticket data together into single data frame, write out .rds.
```{r glue}

# join fish ticket data for 2019-20, make new RDS

# which(names(rawdat1) != names(rawdat2)) # last 22 columns of 2019-20 data do not match
# pacfin2019_2020.df <- rawdat2[,1:101]
# pacfin.df <- bind_rows(rawdat1,pacfin2019_2020.df)
# write_rds(pacfin.df, paste0(out_conf_dir,"pacfin_compiled_2004thru2020.rds"))
# rm(pacfin.df, rawdat1, rawdat2, pacfin2019_2020.df)

```
<br>


Subset the raw data to include only the columns that are needed. Then rename the columns that will be retained in the final processed data. The last three columns are used to calculate per-species / total revenue and landed weight, but will not ultimately be retained.
```{r subset}
rawdat.sub <- select(rawdat, c(FISH_TICKET_ID, PACFIN_PORT_CODE, PACFIN_GROUP_PORT_CODE, VESSEL_NUM, AGENCY_CODE, COUNCIL_CODE, 
                               GEAR_CODE, PACFIN_GROUP_GEAR_CODE, REMOVAL_TYPE_NAME, REMOVAL_TYPE_CODE, DEALER_NUM, FLEET_CODE,
                               LANDING_DATE, LANDING_YEAR, 
                               PACFIN_SPECIES_CODE, LANDED_WEIGHT_LBS, PRICE_PER_POUND, EXVESSEL_REVENUE))

colnames(rawdat.sub) <- c("trip_id","pcid", "pcgroup","drvid", "agid", "council",
                          "grid","grgroup","removal_type", "removal_type_code","proc", "fleet",
                          "tdate", "year", 
                          "spid", "pounds", "ppp","revenue")
```
<br>

```{r include=FALSE}
## clean up space
rm(rawdat)
# rm(pacfin.df, rawdat1, rawdat2, pacfin2019_2020.df)
```
<br>

Remove fish tickets with an unknown (`UNKNOWN`) OR MISSING (`""` / `MISSING`) vessel identifier.
```{r filter}
rawdat.sub <- rawdat.sub %>%
  filter(drvid != "UNKNOWN") %>%
  filter(drvid != "") %>%
  filter(drvid != "MISSING")
```
<br>


## 1.2. Edit, Add columns

First, edit and add columns which describe the fish ticket date. 
```{r edit_columns}
rawdat.sub <- rawdat.sub %>%
  ## create combined vessel ID / year variable
  mutate(drvid_year = paste0(drvid,"_", year)) %>%
  ## create 'date' object for ticket landing date
  mutate(tdate = as.Date(as.character(tdate), "%d-%b-%y")) %>%
  mutate(tdate = ymd(tdate)) %>%
  ## create 'calendar week of landing' variable
  mutate(tweek=week(tdate))

## QC: check first and last date of landing for each year, and total tickets per year.
rawdat.sub %>% group_by(year) %>%
  summarise(first_ticket = min(tdate), last_ticket = max(tdate), total_tickets = length(unique(trip_id)))
```
<br>

Next edit and add columns which describe the landings recorded on the first ticket. First, replace nominal species IDs (*nominal species id that were not shared: rougheye + blackspot (RBR1)*).
```{r nom_ID}
rawdat.sub <- rawdat.sub %>%
  mutate(spid_recode = recode(as.character(rawdat.sub$spid), BMO1 = "BMOL", DVR1 = "DOVR", EGL1 = "EGLS", PTR1 = "PTRL", CSL1 = "CSOL",
                                     REX1 = "REX", RSL1 = "RSOL", SFL1 = "STRY", SSO1= "SSOL", LDB1 = "LDAB", PDB1 = "PDAB", SDB1 = "SSDB", 
                                     ART1 = "ARTH", BSK1 = "BSKT", BLK1 = "BLCK", CNR1 = "CNRY", DBR1 = "DBRK", BLU1 = "BLUR",
                                     BRW1 = "BRWN", CHN1 = "CHNA", CLC1 = "CLCO", COP1 = "COPP", OLV1 = "OLVE", QLB1 = "QLBK", TRE1 = "TREE",
                                     BYL1 = "BYEL", GPH1 = "GPHR", GRS1 = "GRAS", KLP1 = "KLPR", BCC1 = "BCAC", CLP1 = "CLPR", CWC1 = "CWCD",
                                     BRZ1 = "BRNZ", CML1 = "CMEL", GBL1 = "GBLC", GSP1 = "GSPT", GSR1 = "GSRK", HNY1 = "HNYC", MXR1 = "MXRF",
                                     PNK1 = "PNKR", PRR1 = "PRRK", ROS1 = "ROSY", RST1 = "RSTN", SPK1 = "SPKL", SQR1 = "SQRS", STL1 = "STRK",
                                     STR1 = "STAR", SWS1 = "SWSP", TGR1 = "TIGR", VRM1= "VRML", SNS1 = "SNOS", SRK1 = "SRKR", ARR1 = "ARRA",
                                     BGL1 = "BLGL", BNK1 = "BANK", RDB1 = "RDBD", SBL1 = "SBLY", SCR1 = "SCOR", FLG1 = "FLAG", YTR1 = "YTRK",
                                     POP2 = "POP", LSP1 = "LSPN", SSP1 = "SSPN", THD1 = "THDS", WDW1="WDOW", YEY1 = "YEYE", CBZ1 = "CBZN",
                                     KGL1 = "KLPG", LCD1 = "LCOD", CHL1="CHLB", RGL1 = "RCKG", SHP1 = "SHPD"))
```
<br>

Next add columns which describe the species group and port group for landings, based on Holland CCIEA ESR species groupings and IOPAC port codes.
```{r nom_ID}
rawdat.sub <- rawdat.sub %>%
  left_join(
    port_grps, by = c("pcid" = "PACFIN_PORT_CODE")
  ) %>%
  left_join(
    sp_grps, by = c("spid" = "SPID")
  )
```
<br>

Next, add in an estimated exvessel revenue for commercial fish tickets which have no revenue recorded. To do so, we calculate an average price per pound for each species, for every year / week / port group. Then we create an adjusted `ppp` column for trips without revenue, and calculate the `adj_revenue` by multiplying `adj_ppp * pounds`. 
```{r adjust_revenue}
## get the average ppp for each species/year/week/port group, write out 
ppp_key <- rawdat.sub %>%
  filter(ppp > 0) %>%
  group_by(year, tweek, spid_recode, pcgroup) %>%
  summarise(avg_ppp = mean(ppp, na.rm=TRUE))
if(write_ppp){write.csv(ppp_key, paste0(processdir, "Price_per_Pound_Key_",paste0(years,collapse="-"), ".csv"), row.names=FALSE)}

## recalculate revenue for each ticket
rawdat.sub <- rawdat.sub %>%
  mutate(adj_ppp = ifelse(removal_type_code %in% c("C","D"),ifelse(ppp != 0, ppp, as.numeric(filter(ppp_key, year == year & tweek == tweek & spid_recode == spid_recode & pcgroup == pcgroup)$avg_ppp)),ppp)) %>%
  mutate(adj_revenue = ifelse(revenue != 0, revenue, adj_ppp*pounds))

cat("Added in ", sum(!is.na(rawdat.sub$ppp)) - sum(rawdat.sub$ppp != 0), " adjusted ppp.")
```
<br>

# 2: Write out fish tickets

## 2.1. By year
```{r write, eval=TRUE}
for(y in years){
  ## grab data for given year, reorder columns
  tmp_out <- rawdat.sub %>%
    filter(year == years[y]) %>%
    dplyr::select(c(trip_id, year, tdate,agid, 
                       pcgroup, pcid, IOPAC,
                       spid, spid_recode, SPGRPN2, council,
                       grgroup, grid, removal_type,removal_type_code,
                       drvid, drvid_year,proc, fleet, pounds, ppp, adj_ppp,revenue,adj_revenue))
  write.csv(tmp_out, paste0(out_conf_dir, paste0("/fish_tickets_", y, "_processed_for_networks.csv")), row.names=FALSE)
}
```
<br>

## 2.2. By crab year

A crab year is defined as week 46 of year 1 through week 45 of year 2; for example, the Use crab year to split data frame, but remove the crab_year column from data set before writing out to file.
```{r write_crab, eval=TRUE}
if(crab_year){
  ## create "crab_year" column
  rawdat.sub <- rawdat.sub %>%
    mutate(crab_year = ifelse(tweek > 45, year, year - 1))
  ## for each crab year...
  for(y in crab_years){
    ## subset the data frame
    tmp_out <- rawdat.sub %>%
      filter(crab_year == crab_years[y]) %>%
      dplyr::select(c(trip_id, year, crab_year, tdate,agid, 
                       pcgroup, pcid, IOPAC,
                       spid, spid_recode, SPGRPN2, council,
                       grgroup, grid, removal_type,removal_type_code,
                       drvid, drvid_year,proc, fleet, pounds, ppp, adj_ppp,revenue,adj_revenue))
    ## write out
    write.csv(tmp_out, paste0(out_conf_dir, paste0("/fish_tickets_crab", y, "_processed_for_networks.csv")), row.names=FALSE)
  }
}
```
<br>

### SKIP THIS SECTION FOR CCIEA ANALYSIS DEC 2020 

# 3: Append Vessel Lengths

Briefly, the method for calculating vessel lengths from vessel registration data are as follows:

* Step 1: Pull the registration data up to two years prior to the fish ticket year (3 years total)

* Step 2: Remove any registration entries with vessels larger than 200 feet and smaller than 10 feet. *To change these cutoffs, go to the code chunk length_filter*

* Step 3: Find the number of unique vessel lengths for all unique Vessel Number / Agency Code combinations.

* Step 4: Calculate final vessel length for all unique Vessel Number / Agency Code combinations. **(a)** If vessel length data was available for 2+ years and MAX_LENGTH < MIN_LENGTH + 10: assume this reflects an actual increase or decrease in vessel size, and the final vessel length is the median of the two most recent vessel lengths. **(b)** If vessel length data was available for only one year OR vessel length data was available for 2+ years, but $max.length >= min.length+10$: Go to Step 5. **(c)** Three or more different vessel lengths were recorded: use the median of these lengths

* Step 5: For **(b)** above, Pull registration data up to four years prior to the fish ticket year (5 years total); then if...
  + One vessel length was recorded in 2+ years: use this value as vessel length.
  + Two different vessel lengths were recorded: **(a)** $max.length >= 2*min.length$ -- this is probably a decimal point error that we would need to check manually to determine true length. Denote NA. **(b)** $max.length < 2*min.length$ -- Use the mode of these lengths.


## 3.1. Vessel Registration Data

Load registration data
```{r load reg}
# permitsfiles <- list.files(here::here('data','raw','vessel registration'),full.names = T)
# permits <- purrr::map_df(permitsfiles,function(fl){
#   read_csv(fl,col_types='ddcddccdcccccdddcccccccddll')
# }) %>% distinct()
```



Choose columns of interest and subset the data frame. Change all entries of `777` for vessel length to `NA` (recommendation from Blake Feist)
```{r pthin}
# pcols <- c("VESSEL_NUM", "AGENCY_CODE","REGISTRATION_YEAR","VESSEL_LENGTH", "LENGTH_TYPE_CODE")
# pthin <- permits %>%  select(pcols) %>%
#   mutate(VESSEL_LENGTH=na_if(VESSEL_LENGTH,777))
```
<br>

Remove any vessel lengths greater than 200 or lower than 10. *Note - depending on the fishery, may want to change these cutoffs.*
```{r length_filter}
# pthin_length_filter <- pthin %>%
#   filter(VESSEL_LENGTH < 200,VESSEL_LENGTH > 10)
```
<br>

If option `dcrb_only` is true, filter the landings data so that it only includes vessels with positive exvessel revenue for Dungeness crab.
```{r}
# if(dcrb_only){
#   vessels <- rawdat.sub %>%
#     filter(drvid != "UNKNOWN", drvid != "MISSING", drvid != "", !is.na(drvid)) %>%
#     filter(DCRB_revenue > 0) %>%
#     select(drvid, agency_code, year) %>%
#     distinct()
# } else{
#   vessels <- rawdat.sub %>%
#   filter(drvid != "UNKNOWN", drvid != "MISSING", drvid != "", !is.na(drvid)) %>%
#   select(drvid, agency_code, year) %>%
#   distinct()
# }
```
<br>


## 3.2. Calculate Vessel Lengths

Initiate empty data frame for all vessel length data across years
```{r}
# vessel_length_key_df <- tibble("drvid" = character(),
#                                "agency_code" = character(),
#                                "year" = numeric(),
#                                "FINAL_LENGTH" = numeric(),
#                                "TYPE_CALC" = character(),
#                                "UNIQUE_LENGTHS" = numeric(),
#                                "N_YEARS_LENGTH_RECORDED" = numeric(),
#                                "HISTORIC_DATA" = character())
```
<br>

Do the calculation of vessel lengths.
```{r}
# for(y in unique(vessels$year)){
#   # take only the vessels fishing in year "y"
#   year_vessels <- filter(vessels, year == y)
#   
#   # identify the years of regulation data to pull
#   target_reg_years <- seq(y-2, y)
#   
#   # subset for target registration years and summarise permits for each vessel
#   cat("Calculating 3yr summary statistics for vessels in ", y, "\n")
#   pthin_sumstats <- pthin_length_filter %>%
#     filter(REGISTRATION_YEAR %in% target_reg_years,!is.na(VESSEL_LENGTH)) %>%
#     group_by(VESSEL_NUM, AGENCY_CODE) %>%
#     arrange(desc(REGISTRATION_YEAR)) %>%
#     summarise(n_lengths = length(VESSEL_LENGTH),
#               n_unique = length(unique(VESSEL_LENGTH)),
#               max_length = max(VESSEL_LENGTH),
#               min_length = min(VESSEL_LENGTH),
#               median2yr = get2yrmedian(x=VESSEL_LENGTH, years=REGISTRATION_YEAR)) %>% 
#     ungroup()
#   
#   # create empty vectors for this year
#   final_vessel_lengths <- c()
#   length_calc_vec <- c()
#   n_unique_vec <- c()
#   n_lengths_vec <- c()
#   historic_vec <- c()
#   processed <- 0
#   
#   cat("Calculating vessel lengths for fishing vessels in ", y, "...\n")
#   # for each vessel fishing in this year #
#   for(i in seq(1:length(year_vessels$drvid))){
#     ## use the calc_length function (loaded from the "functions.R" file) to calculate vessel length
#     tmp_vessel_length_info <- calc_length(permits=permits, vesseldat = year_vessels, lengthdat = pthin_length_filter, summarydat = pthin_sumstats, index = i)
#     ## save the calc_length output to the appropriate position ("i") in the output vectors for this year
#     n_lengths_vec[i] <- tmp_vessel_length_info[1] %>% as.numeric()
#     n_unique_vec[i] <- tmp_vessel_length_info[2] %>% as.numeric()
#     final_vessel_lengths[i] <- tmp_vessel_length_info[3] %>% as.numeric()
#     length_calc_vec[i] <- tmp_vessel_length_info[4]
#     ## if the vessel had to be calculated with historical data from over 5 years ago, a warning message will be saved in the calc_length output
#     if(length(tmp_vessel_length_info) > 4){
#       ### print the warning message
#       print(tmp_vessel_length_info[5])
#       ### save "Y" to the historic_vec for this year
#       historic_vec[i] <- "Y"
#     } else{ historic_vec[i] <- "N" }
#     processed <- processed + 1
#   }
#   cat("Done processing", processed, "vessels for", y, "\n")
#   # save allof the output vectors to a data frame for this year
#   tmp_vessel_length_key_df <- tibble("drvid" = year_vessels$drvid,
#                                      "agency_code" = year_vessels$agency_code,
#                                      "year" = year_vessels$year,
#                                      "FINAL_LENGTH" = final_vessel_lengths,
#                                      "TYPE_CALC" = length_calc_vec,
#                                      "UNIQUE_LENGTHS" = n_unique_vec,
#                                      "N_YEARS_LENGTH_RECORDED" = n_lengths_vec,
#                                      "HISTORIC_DATA" = historic_vec)
#   ## bind this year's data frame to the end of the full data frame
#   vessel_length_key_df <- bind_rows(vessel_length_key_df, tmp_vessel_length_key_df)
#   cat("Save", dim(tmp_vessel_length_key_df)[1], "lengths for", y, " to final data frame\n\n")
# }
```
<br>

## 3.3. Check Vessel Lengths

What is the distribution of vessel lengths?
```{r echo=FALSE}
# vessel_length_key_df %>% 
#     filter(!is.na(FINAL_LENGTH)) %>% 
#     ggplot(aes(as.numeric(FINAL_LENGTH))) +
#       geom_histogram(bins=10) +
#       facet_wrap(~year) +
#       labs(x="Vessel Length",y="Number of Vessels",title="Vessel Length Frequency Distribution")
```
<br>

Which calculations were used to estimate vessel lengths? Ideally, the frequency of *unique_1* and *olddat_unique_1* will be low.
```{r echo=FALSE}
# vessel_length_key_df %>% 
#     ggplot(aes(vessel_length_key_df$TYPE_CALC)) +
#     geom_bar() +
#     facet_wrap(~year) +
#     labs(x="Calculation Type",y="Number of Vessels")+
#     theme(axis.text.x = element_text(angle=90, hjust=1))
```
<br>

How many vessel lengths are missing?
```{r echo=FALSE}
# cat(sum(is.na(vessel_length_key_df$FINAL_LENGTH)), 'vessels with Vessel ID numbers are missing calculated lengths, or', sum(is.na(vessel_length_key_df$FINAL_LENGTH)) / length(vessel_length_key_df$FINAL_LENGTH)*100, 'percent.')
```
<br>

```{r echo=FALSE}
# for(y in unique(vessel_length_key_df$year)){
#   for(a in unique(vessel_length_key_df$agency_code)){
#     tmp_dat <- vessel_length_key_df %>%
#       filter(agency_code == a) %>%
#       filter(year == y)
#     missing <- sum(is.na(tmp_dat$FINAL_LENGTH))
#     cat("Number", y, "Vessels Missing Vessel Lengths for", a, ":", missing, "\n")
#     cat("Proportion:", missing/length(tmp_dat$FINAL_LENGTH), "\n\n")
#   }
# }
```
<br>

Save the vessel lengths key for each year.
```{r eval=TRUE}
# for(y in unique(vessel_length_key_df$year)){
#   key_subset <- vessel_length_key_df %>%
#     filter(year == y)
#   write_rds(key_subset,here::here('data','input','length',paste0("vessel_length_key_", y, ".csv")))
# }
```
<br>
<br>


# 4: Write out fish tickets with length


```{r eval=TRUE}
# for(y in unique(vessel_length_key_df$year)){
#   landings_subset <- rawdat.sub %>%
#     filter(year == y)
#   vessel_length_key_df_subset <- vessel_length_key_df %>%
#     filter(year==y) %>%
#     select(-year)
#   dcrb_landings_length <- left_join(landings_subset,vessel_length_key_df_subset, by=c("drvid", "agency_code"))
#   write.csv(dcrb_landings_length, file= here::here('data','input',paste0('processed_fishtickets_with_vlength_', y, '.csv')))
# }
```
<br>


---

Script runtime: 
```{r echo=FALSE}
Sys.time()-script_start_time
```
