---
title: "Identify Metiers, Infomap"
author: "E. Fuller, M. Fisher"
date: "Written Jan. 26, 2019. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in Fisher et al.
fontsize: 11pt
---

# Description

Using the metiers identified in script 1, assign metiers to the remaining fish ticket data based on the k-nearest neighbor (KNN) algorithm. KNN identifies the "nearest neighbors" of a given fish ticket based on Hellinger dissimilarities. 

Before running this script, you should have used `test_kvals_knn.Rmd` to identify the best `k` value for this port group. The default k value is 1, due to testing with port groups CCA and MNA. 

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("vegan")) {install.packages("vegan")}
if(!require("purrr")) {install.packages("purrr")}
if(!require("parallelDist")) {install.packages("parallelDist")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
```
<br>

And calls the following functions:
```{r}
source('R/knn_matrix.R')
source('R/knn_assign.R')
```
<br>

# User Inputs 

Select your directories.
```{r get_dir}
## input directory containing fish ticket data without assigned metiers
indir_tix <- 'data/input'

## input directory containing fish tickets from reference years, with infomap-assigned metiers
indir_ref <- 'data/intermediate_metiers/reference'

## output directory
outdir <- 'data/intermediate_metiers'

## final output directory for tickets without missing metiers
outdir_final <- 'data/input/metiers'
```
<br>

Set the value of `k` for the k-nearest neighbor function. This tells the function the number of 'nearest neighbors' to look at for assigning metier membership. You should have explored various values of `k` to find the one that is most appropriate for your data.
```{r}
k <- 1
```
<br>

Filter the new fish tickets without metiers for commercial tickets?
```{r}
filter_type = TRUE
```
<br>

The script will loop over each of the ports and crab years in these vectors.
```{r}
myports <- c("CCA","ERA","BGA","BDA","SFA","MNA","MRA")
years <- c(2008,2009,2012,2013,2014,2015,2016,2017)
```
<br>

The default settings will write out a different file for each port group and year; you can also write a combined output file with the landings + metiers from all port groups in a given crab year.
```{r}
combined_output <- TRUE
```
<br>

Set these objects according to your user inputs from Script 1
```{r}
## trip comparisons for metier identification based on revenue or lbs
metric <- "revenue"

## which year(s) of fish ticket data were used to identify metiers with infomap?
ref_years <- c(2010,2011)
```
<br>


# 1: Metier Assignment

By port group.

## 1.1. Prep for writing files

List of output data frames with all port groups in a year, to write combined output file.
```{r}
if(combined_output){
  outlist <- vector("list", length(years))
  names(outlist) <- years
}
```
<br>

This short function will write out the fish tickets from the reference years in the same format as the rest of the years of fish tickets post-metier assignment, for future use.
```{r}
rewrite_ref <- function(ref_data, crab_years, port, mydir){
  for(y in crab_years){
    tmptix <- filter(metiers,crab_year==y)
    write.csv(tmptix, here::here(mydir, paste0( "FishTix_",port,"_crab", y, "_metiers.csv")), row.names=FALSE)
  }
}
```
<br>

## 1.2 KNN

Loop over ports and crab years to assign metiers using the k-nearest neighbor algorithm.
```{r knn}
for(myport in myports){
  
  # READ IN TRAINING DATA
  metiers <- read.csv(here::here(indir_ref, paste0("FishTix_", metric, "_", paste0(ref_years,collapse="_"), "_", myport, "_metiers.csv")))
  head(metiers)
  # write out as individual crab years for future use
  rewrite_ref(ref_data=metiers, crab_years = ref_years, port = myport, mydir=outdir_final)
  
  # ORGANIZE TRAINING DATA
  ## Split data into a data frame with ticket information, and a data frame with a key to metiers
  traindat <- dplyr::select(metiers, -metier.num, -metier)
  metier_key <- metiers %>%
    dplyr::select(trip_id, metier.num, metier) %>%
    distinct()
  met_names_key <- metier_key %>%
    dplyr::select(metier.num, metier) %>%
    distinct()
  head(traindat); head(metier_key)
  ## Make sure there are no duplicates in the key *(sum should equal 0)*
  count_trips <- metier_key %>%
    group_by(trip_id) %>%
    summarise( n_trips = n())
  if(sum(count_trips$n_trips > 1) > 1){
    stop("there are duplicates in your metier key!")
  } else{message("there are no duplicates in your metier key.")}
  
  # ASSIGN METIERS
  for(y in years){
    # read in test data without metiers
    tickets <- read.csv(here::here(indir_tix, paste0("fish_tickets_crab", y, "_processed_for_networks.csv")))
    
    # filters for port group of interest / commercial tickets
    if(filter_type){
    tickets_filtered <- tickets %>%
      filter(pcgroup == myport) %>%
      filter(removal_type == "COMMERCIAL (NON-EFP)" | removal_type == "COMMERCIAL(DIRECT SALES)")
    } else{
      tickets_filtered <- tickets %>%
        filter(pcgroup == myport)
    }
    
    # start a list of trips that have been 
    trips_assigned <- c()
    
    # get a list of the gear groups
    gear_list <- unique(tickets_filtered$grgroup)
    
    # prep a data frame to hold metier info
    test_metiers <- data.frame(trip_id = as.character(),
                               metier.num = as.numeric())
    
    # for each gear type...
    for(g in gear_list){
      cat("processing gear group ", as.character(g), ". \n")
      ## get a list of trips NOT already assigned a metier, and within given gear group
      tmp_trips <- tickets_filtered %>%
        filter(!(trip_id %in% trips_assigned)) %>%
        filter(grgroup == g) %>%
        dplyr::select(trip_id) %>%
        distinct()
      if(length(tmp_trips$trip_id) > 0){
        ## get all of the fish tickets associated with those trips
        tmp_tix <- left_join(tmp_trips, tickets_filtered, by="trip_id")
        ## run knn
        tmp_test_metiers <- knn_assign_metiers(train = traindat, key = metier_key, test = tmp_tix, k = k, adjust.revenue = TRUE, print.plot=TRUE)
        
        ## append new assignments to data frame
        test_metiers <- rbind(test_metiers, tmp_test_metiers)
        ## save the trip ids of the trips assigned metiers
        trips_assigned <- c(trips_assigned, tmp_trips$trip_id)
        
        ## remove objects
        rm(tmp_trips)
        rm(tmp_tix)
        rm(tmp_test_metiers)
      } else{
        cat("all trips from gear group ", g, " already classified. moving to next gear group.")
      }
    }
    
    #join metier IDs to fish tickets
    test_metiers$trip_id <- as.character(test_metiers$trip_id); tickets_filtered$trip_id <- as.character(tickets_filtered$trip_id)
    tickets_w_metiers <- left_join(tickets_filtered, test_metiers, by="trip_id")
    
    #join metier names to fish tickets
    tickets_w_metiers$metier.num <- as.character(tickets_w_metiers$metier.num)
    met_names_key$metier.num <- as.character(met_names_key$metier.num)
    tickets_w_metiers <- left_join(tickets_w_metiers, met_names_key, by="metier.num")
    colnames(tickets_w_metiers)[dim(tickets_w_metiers)[2]] <- "metier.name"
    missing_metiers <- sum(is.na(tickets_w_metiers$metier.num))
    
    # write out
    cat("writing out file for year", y, ". ", missing_metiers, " tickets (", missing_metiers/dim(tickets_filtered)[1]*100, "%) missing metier assignment. \n")
    write.csv(tickets_w_metiers, here::here(outdir, paste0( "FishTix_",myport,"_crab", y, "_metiers_k", k, ".csv")), row.names=FALSE)
    
    # add to combined output data frame
    if(combined_output){
      if(myport == myports[1]){
        outlist[[which(names(outlist)==y)]] <- tickets_w_metiers
      } else{ #this could be rewritten using purrr
        tmpdf_fromList <- outlist[[which(names(outlist) == y)]]
        tmpdf_fromList <- rbind(tmpdf_fromList, tickets_w_metiers)
        outlist[[which(names(outlist)==y)]] <- tickets_w_metiers
      }
    } #end if(combined_output)
    
    # remove test metiers
    rm(tickets_w_metiers, test_metiers, tmpdf_fromList)
    
  } #end for(y in years)
  
} #end for(port in myports)
```
<br>

# 2. Write Combined Output

```{r write_all}
if(combined_output){
  for(y in years){
    tmpdf_fromList <- outlist[[which(names(outlist) == y)]]
    write.csv(tmpdf_fromList, here::here(outdir, paste0("FishTix_crab", y, "_metiers_k", k, ".csv")), row.names=FALSE)
  }
}
```
<br>

---

Script Runtime:
```{r echo=FALSE}
Sys.time()-script_start_time
```


