---
title: "Identify Metiers P. 2, Infomap"
author: "E. Fuller, M. Fisher"
date: "Written Jan. 30, 2019. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in Fisher et al.
fontsize: 11pt
---

# Description

Fish tickets which were not assigned a metier through k-nearest neighbor analysis will be combined across years, within each port group, and run through infomap to identify metiers that may not have been present in the original reference years.

It is important to run this script with as many years of data as possible, to help infomap identify unique metiers without oversplitting the data. Best practice is to wait until all years of interest have been processed with script 03 before running this script.
<br>

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("parallelDist")) {install.packages("parallelDist")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("labdsv")) {install.packages("labdsv")}
if(!require("igraph")) {install.packages("igraph")}
```
<br>

And calls the following functions:
```{r}
source("R/generate_catch_matrix_reruns.R")
source("R/filter_matrix.R")
source("R/split_matrix.R")
source("R/convert_distance_matrix.R")
source("R/distance_to_similarity.R")
source("R/merge_data.R")
source("R/average_similarities.R")
source("R/network_edges.R")
source("R/identify_metiers.R")
source("R/metier_key_reruns.R")
```
<br>

# User Inputs 

Select your directories.
```{r get_dir}
## input directory containing fish ticket data with KNN-assigned metiers
indir_tix <- 'data/intermediate_metiers'

## input directory containing fish tickets from reference years, with infomap-assigned metiers
indir_ref <- 'data/intermediate_metiers/reference'

## output directory for tickets fully assigned metiers
outdir <- 'data/input/metiers'
```
<br>

Identify the crab years and port groups with landings data that need to be checked for fish tickets that were not assigned metiers using the KNN algorithm.
```{r}
years <- c(2008,2009,2012,2013,2014,2015,2016,2017)
myports <- c("CCA","ERA","BGA","BDA","SFA","MNA","MRA")
```
<br>

Set these objects according to your user inputs from Scripts 1 and 3
```{r}
## trip comparisons for metier identification based on revenue or lbs
metric <- "revenue"

## filter to include only commercial tickets?
filter_type = TRUE

## the value of `k` for the k-nearest neighbor function
k <- 1

## the reference years first run through infomap
ref_years <- c(2010,2011)
```
<br>

Do you want to filter the fish tickets to remove catch from rare species? 
```{r}
filter_rare <- 2
```
<br>

# 1: Prepare Data

Read in fish tickets from non-reference years, with KNN-assigned metiers.
```{r rawdata}
for(y in years){
  for(port in myports){
    if(port == myports[1] & y == min(years)){
      rawdat <- read.csv(here::here(indir_tix, paste0("FishTix_",port,"_crab", y, "_metiers_k", k, ".csv")))
    } else{
      tmp_rawdat <- read.csv(here::here(indir_tix, paste0("FishTix_",port,"_crab", y, "_metiers_k", k, ".csv")))
      rawdat <- rbind(rawdat,tmp_rawdat)
    }
  }
}

colnames(rawdat)
```


## 1.1. Select unassigned tickets 

Filter for trip IDs from fish tickets that do not have an assigned metier, but do have non-zero revenue. 
```{r NAtrips}
tripsNA <- rawdat %>%
  dplyr::select(trip_id, crab_year, metier.num) %>%
  filter(is.na(metier.num)) %>%
  distinct()
cat("Total trips in data: ", length(unique(rawdat$trip_id)), "\nTrips missing metier assignments: ", dim(tripsNA)[1])
```
<br>

Grab fish tickets associated with the trip IDs
```{r NAtrips_data}
mydat <- filter(rawdat, trip_id %in% tripsNA$trip_id)
```
<br>

## 1.2. Filter for nonzero revenue

```{r filter_rev}
# get total revenue for each trip
rev_bytrip <- mydat %>%
  group_by(trip_id) %>%
  summarise(total_rev = sum(revenue))
# filter for trips with revenue > 0
trips_w_revenue <- filter(rev_bytrip, total_rev > 0)
# select 'mydat' tickets with revenue > 0
mydat <- filter(mydat, trip_id %in% trips_w_revenue$trip_id)
cat("Out of ", length(rev_bytrip$trip_id), " trips, ", length(trips_w_revenue$trip_id), " have nonzero revenue.\n")
```
<br>


# 2: Run Infomap

The following code loops over the list of port groups, and for each: 

1. Process raw fish tickets into a matrix format, with each row a trip and each column a gear-catch combination

2. Creates the Hellinger distance matrix

3. Converts the distance matrix to a similarity (adjacency) matrix

4. Uses `infomap` to identify metiers in R

5. Names metiers. This can be done manually, or just by taking the highest grossing species / gear combination.

6. Adds metiers to processed fish tickets

```{r infomap}
for(p in unique(mydat$pcgroup)){
  cat("working on port group: ", p, ".\n")
  portdat <- filter(mydat, pcgroup == p)
  cat(length(unique(portdat$trip_id)), " trips at this port.\n")
  
  ##### 1. CREATE CATCH MATRIX #####
  revmat <- gen_catch_matrix_rerun(mydata = portdat, multigear=TRUE, NAto0 = TRUE)
  # Optional filtering
  final.revmat <- filter_matrix(mydata = revmat, commercial = FALSE, rare = filter_rare, type= "number", ids = c("drvid_year", "crab_year", "trip_id"))
  
  
  inmat <- final.revmat
  # Separate IDs from Data
  inmat_ids <- inmat %>%
    dplyr::select(drvid_year, crab_year, trip_id) %>%
    mutate(index = row.names(inmat))
  inmat_ids$index <- as.numeric(inmat_ids$index)
  inmat <- as.matrix(dplyr::select(inmat, -drvid_year, -crab_year, -trip_id))
  
  if(dim(inmat)[2] == 1){
    # there is only one species; create community key for that species
    trip_community_key <- final.revmat[,c("trip_id", "drvid_year", "crab_year")]
    trip_community_key <- trip_community_key %>%
      mutate(comm_num = 1,
             comm_id = colnames(inmat))
    # plot
    myplot <- ggplot(data=melt(final.revmat, id.vars = c("drvid_year", "crab_year", "trip_id")), aes(x=name, y=value, fill=name)) +
      geom_col() +
      ggtitle(paste0(p, " - 1")) +
      guides(fill=FALSE) +
      theme(axis.text.x = element_text(angle=90, hjust=1))
    print(myplot)
    png(here::here('data/intermediate_metiers/plots/', paste0(p, "_rerun_community_1.png")))
    print(myplot)
    dev.off()
    rm(final.revmat, revmat) #clean up
  } else{
    rm(final.revmat, revmat) #clean up
    
    
    ##### 2. HELLINGER DISTANCE #####
    hell.dist <- parDist(inmat, method="hellinger", diag=FALSE, upper=FALSE)
    hell.df <- convertDist(hell.dist)
    rm(hell.dist)
    
    
    ##### 3. SIMILARITY #####
    hell.sim.df <- dist2sim(x=hell.df,col=3, max.type = "individual")
    rm(hell.df)
    similarities <- left_join(hell.sim.df, dplyr::select(inmat_ids, trip_id, index), by=c("row" = "index"))
    colnames(similarities) <- c(colnames(hell.sim.df), "trip1") #rename
    similarities <- left_join(similarities, dplyr::select(inmat_ids, trip_id, index), by=c("col" = "index"))
    colnames(similarities)[6] <- "trip2" #rename
    similarities <- similarities[,c("trip1", "trip2", "value.sim")]
    
    
    ##### 4. CREATING IGRAPH OBJECT; RUN INFOMAP ALGORITHM #####
    # Save unique trip names; associated drvid_year
    all_trips <- c(similarities$trip1, similarities$trip2)
    all_trips <- unique(all_trips)
    trip_info <- dplyr::select(inmat_ids, trip_id, drvid_year, crab_year)
    cat("There are ", length(all_trips), " trips in similarities data set.")
    # Create vertices data frame
    vertices_df <- data.frame(rownum = seq(1, length(all_trips)),
                              node = as.character(all_trips))
    vertices_df$node <- as.character(vertices_df$node); trip_info$trip_id <- as.character(trip_info$trip_id)
    vertices_ext <- left_join(vertices_df, trip_info, by=c("node" = "trip_id"))
    # Create edges data frame
    edges_df <- generate_network_edges(similarities = similarities, vertices = vertices_ext)
    cat("There are ", dim(edges_df)[1], " edges.")
    
    # igraph & infomap
    mycommunities <- identify_metiers(vertices=vertices_ext, edges = edges_df, trials = 100)
    
    
    
   ##### 5. ASSIGN COMMUNITIES BACK TO TRIPS #####
    # match community designations with trip ids
    vertices_membership <- membership(mycommunities)
    vertices_ext[,"community"] <- NA
    for(i in seq(1,length(vertices_ext$node))){
      vertices_ext[i,"community"] <- vertices_membership[[i]]
    }
    # put info back into revenue matrix
    inmat <- cbind(trip_info, inmat)
    inmat_comm <- left_join(inmat, vertices_ext, by=c("trip_id" = "node", "drvid_year"="drvid_year", "crab_year" = "crab_year"))
    inmat_comm <- dplyr::select(inmat_comm, -rownum)
    inmat_comm.melt <- pivot_longer(inmat_comm, cols = colnames(inmat_comm)[which(!(colnames(inmat_comm) %in% c("drvid_year", "crab_year", "trip_id", "community")))])
    
    
    ##### 6. NAME COMMUNITIES #####
    # summarise the revenue from each species in the community
    rev_by_community <- inmat_comm.melt %>%
      group_by(community, name) %>%
      summarise(total_rev = sum(value))
    cat("There are ", length(unique(rev_by_community$community)), " new communities.\n")
    # plot
    for(i in unique(rev_by_community$community)){
      tmp_dat <- rev_by_community %>%
        filter(community == i) %>%
        filter(total_rev > 0)
      myplot <- ggplot(tmp_dat, aes(x=name, y=total_rev, fill=name)) +
        geom_col() +
        ggtitle(paste0(p, "-", i)) +
        guides(fill=FALSE) + theme_bw() +
        theme(axis.text.x = element_text(angle=90, hjust=1))
      print(myplot)
      png(here::here('data/intermediate_metiers/plots/', p, paste0( "rerun_metier_",i,".png")))
      print(myplot)
      dev.off()
    }
    # create trip-community key, with named communities
    trip_community_key <- metier_key_rerun(revenue = rev_by_community, vertices = vertices_ext, name=TRUE)
  } #end if more than one species in final revenue matrix
  # Write out key
  write.csv(trip_community_key, here::here(indir_ref, paste0("FishTix_",metric,"_", years[1],"_",tail(years, n=1),"_",p, "_InfoMap_Metier_Key_ReRun.csv")), row.names=FALSE)
  
  cat("\n\n---\n")
}
```
<br>

```{r include=FALSE}
## Clean up environment ##
rm(edges_df, vertices_df, vertices_ext, hell.sim.df, inmat, inmat_comm, inmat_comm.melt, inmat_ids, 
   mycommunities, mydat, rawdat, myplot, portdat, rev_by_community, rev_bytrip,
   similarities, tmp_dat, tmp_rawdat, trip_community_key, trip_info, trips_w_revenue, tripsNA, p)
```
<br>

# 3: Add Metiers to Fish Tickets

Use trip-community key from Section 2 to replace "NA" metiers in fish ticket data
```{r add_metiers}
for(p in myports){
  cat("Port Group ", p, "\n")
  rerun_key <- read.csv(here::here(indir_ref,paste0("FishTix_",metric,"_", years[1],"_",tail(years, n=1),"_",p, "_InfoMap_Metier_Key_ReRun.csv")))
  years <- sort(unique(rerun_key$crab_year))
  
  # read in fish tickets
  for(y in years){
    if(y == min(years)){
      fishtix <- read.csv(here::here(indir_tix,paste0("FishTix_", p, "_crab", y, "_metiers_k", k, ".csv")))
      fishtix$metier.name <- as.character(fishtix$metier.name)
      fishtix$metier.num <- as.character(fishtix$metier.num)
      cat("read in fish tickets from crab year ", y, "\n")
    } else{
      tmp_tix <- read.csv(here::here(indir_tix, paste0("FishTix_", p, "_crab", y, "_metiers_k", k, ".csv")))
      tmp_tix$metier.name <- as.character(tmp_tix$metier.name)
      tmp_tix$metier.num <- as.character(tmp_tix$metier.num)
      fishtix <- rbind(fishtix, tmp_tix)
      cat("appended fish tickets from crab year ", y, "\n")
    }
  }
  
  # read in original metiers
  orig_key <- read.csv(here::here(indir_ref, paste0("FishTix_", metric, "_", paste0(ref_years,collapse="_"), "_", p, "_metiers.csv"))) %>%
    dplyr::select(metier.num, metier) %>%
    distinct()
  
  # get last metier number and all metier names of existing fish tickets
  met_exist <- as.character(unique(orig_key$metier))
  max_num <- max(orig_key$metier.num, na.rm=TRUE)
  
  # add column to key that re-numbers new metiers, and that adds an "r" if the metier name already exists
  rerun_key$comm_id <- as.character(rerun_key$comm_id)
  rerun_key <- rerun_key %>%
    mutate(comm_num_edit = comm_num + max_num) %>%
    mutate(comm_id_edit = ifelse(comm_id %in% met_exist, paste0(comm_id, "r"), comm_id))
  cat("replicated metier names: ", length(which("r" %in% rerun_key$comm_id_edit)), "\n\n")
  
  # for each trip in the key
  for(trip in rerun_key$trip_id){
    ## save the original metier
    old_met <- unique(fishtix$metier.num[which(fishtix$trip_id == trip)])
    ## if the original metier is NA (it should be!)
    if(is.na(old_met)){
      ### save the new metier info from the key for that trip
      tmp_key <- filter(rerun_key, trip_id == trip)
      ### replace the metier num and name with the new metier info
      fishtix$metier.num[which(fishtix$trip_id == trip)] <- tmp_key$comm_num_edit
      fishtix$metier.name[which(fishtix$trip_id == trip)] <- tmp_key$comm_id_edit
    }
    ## if the original metier is not NA, there was a mistake!
    else{
      message("ERROR: Trying to replace an existing metier for trip ", trip)
    }
  }
  
  # write out new fish tickets
  for(y in years){
    tmp_fishtix <- filter(fishtix, crab_year == y)
    write.csv(tmp_fishtix, here::here(outdir, paste0("FishTix_", p, "_crab", y, "_metiers_k", k, "_recoded.csv")), row.names=FALSE)
  }
}
```
*Note - sometimes R doesn't like reading in the "NAs" from .csv file, so will return the warning "invalid factor level, NA generated." Just ignore - I checked to make sure that all NAs in R were NAs in the original .csv*




---

Script runtime:
```{r echo=FALSE}
Sys.time()-script_start_time
```
