---
title: "Create Participation Networks"
author: "E. Fuller, M. Fisher, J.F. Samhouri"
date: "Written Dec. 14, 2020. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

Creates an annual fisheries particiation network (igraph object) for each port group, for every crab year. Then calculates summary statistics for each network.

Section 3 allows you to calculate network metrics from existing igraph objects. Set Section 2 code chunks to `eval=FALSE` and Section 3 code chunks to `eval=TRUE` to run Section 3. 


<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("igraph")) {install.packages("igraph")}
```
<br>

And calls the following functions:
```{r}
source("R/generate_participation_network.R")
source("R/color_network_nodes.R")
source("R/network_stats_functions.R")
source("R/calc_network_stats.R")
```
<br>

# User Inputs 

Select your directories.
```{r}
# indir = '/Users/jameal.samhouri/Documents/CCIEA Networks/processed'
indir = 'data/input'

## output directory for igraph object
outdir = 'data/networks/participation/edge-type-vessels'

## output directory for network viz
pngdir = 'data/networks/participation/edge-type-vessels/plots'

## output file (including directory) for stat summaries
statdir = 'results/statistics'

## directory for species group key. not confidential
processdir <- "data/input"
```
<br>

Specify file name of the species groupings lookup key that contains labels for the network graphs.
```{r get_filenames}

myfile5 <- "spgrpn2_iea_key.csv"

```
<br>

Identify the crab years and port groups that you would like to produce networks for. The port groups vector should include only those port groups which are present in the single data file produced with script 05. 
```{r}
## (crab) years
years <- c(2019)

## IOPAC port groups
myports <- c("Puget Sound","North WA Coast","WA Coast","Other Coastal Washington","Astoria","Tillamook","Columbia River","Newport","Coos Bay","Brookings","Crescent City","Eureka","Morro Bay","Fort Bragg","Bodega Bay","San Francisco","San Diego","Monterey","Santa Barbara","Los Angeles","Unknown Ports" )
```
<br>

Some Dungeness crab landings may be recorded prior to the official opening date of the season as a part of domoic acid testing. We remove these landings because we are interested in flows of fishers between fisheries as a result of within-season activity.
```{r}
rm_crab = TRUE
```
<br>

For confidentiality, three or more vessels must be participating in a given fishery for that fishery to be included in the networks. To avoid inclusion of rare or minimal fishing activity, a given fishery must contribute to at least 10% of total seasonal revenue at the given port group to be included in the network data. 
Both cutoff values can be increased if more stringent fishery inclusion criteria are preferred, and the contribution cutoff may be decreased (i.e. for summaries over shorter temporal / smaller spatial scales).
```{r}
vessel_cutoff <- 3
contr_cutoff <- 0.10
```
<br>

To avoid inclusion of rare or minimal fishing activity, (1) a given vessel must make at least XX dollars of total commercial revenue, and (2) a given fishery must contribute at least XX dollars of seasonal revenue for a given vessel to be included in the network data. 
Both cutoff values can be increased if more stringent fishery inclusion criteria are preferred, and the contribution cutoff may be decreased (i.e. for summaries over shorter temporal / smaller spatial scales).
```{r}
vessel_rev_cutoff <- 5000
fish_rev_cutoff <- 500
```
<br>


Set these objects according to your user inputs from Scripts 1 and 3
```{r}
## the value of `k` for the k-nearest neighbor function
k <- 1

## the reference years first run through infomap
## for the cciea network analysis this does not matter because we use a prior sp groupings
ref_years <- c(2004,2005)
```
<br>


# 1: Read in data

Read in the landings data, from the file containing all fish tickets across port groups, within a single crab year.

NOTE: this step takes a while.

NOTES FOR JAMEAL: 
1) read in fish ticket files with this naming convention: fish_tickets_crab2004_processed_for_networks.csv
2) if species groupings column is named something other than metier.name, you'll have to go through the generate_participation_network.R function to replace it (done, JS note 010521)
```{r}
for(y in years){
  if(y %in% ref_years){
    tmptix <- read.csv(here::here(indir, paste0("fish_tickets_crab", y, "_processed_for_networks.csv")), stringsAsFactors = FALSE) %>%
    filter(IOPAC %in% myports)
  } else{
    tmptix <- read.csv(here::here(indir, paste0("fish_tickets_crab", y, "_processed_for_networks.csv")), stringsAsFactors = FALSE) %>%
    filter(IOPAC %in% myports)
  }
  if(exists('fishtix')){
    fishtix <- rbind(fishtix, tmptix)
  } else{ fishtix <- tmptix }
}
rm(tmptix)

# some checks on species groups
unique(fishtix$SPGRPN2)
length(which(is.na(fishtix$SPGRPN2)))
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$spid)
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$year)
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$agid)
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$pcid)

```
<br>

Get the opening dates associated with each port group, for each crab year. 
NOTE: Not needed for CCIEA networks analysis
```{r}
#dates_df <- read.delim(here::here('data/input','crab_open_dates.txt')); head(dates_df)
```
<br>

Make sure that all dates are `Posixt` objects and species group labels are characters.
```{r}
#dates_df$odate <- mdy(dates_df$odate)
fishtix$tdate <- date(parse_date_time(fishtix$tdate, orders=c("ymd", "mdy")))
fishtix$SPGRPN2 = as.character(fishtix$SPGRPN2)
```
<br>


# 2: Make Participation Networks

For each crab year and port group, this code will:
1. Create a `graph` object for each fisheries participation network, and then save the object as an R data file

2. Plot the network graphs

3. Calculate network-level statistics, and save to an output data frame.


Legacy from Fisher et al analysis:
1. split the data into an early and late season

NOTE FOR JAMEAL: remove reference to met_names
```{r eval=TRUE}

#### Read in species group labels key ####
node_names <- read_csv(here::here(processdir,myfile5),
                       col_types = list(col_character(), col_character(), col_character()))

# y=2019
# p = "AST"
# iopac = "Astoria"
for(y in years){
  for(iopac in myports){ # updated to iopac 03-10-21, was p
    cat(iopac, "\n")
    
    # #### Subset data ####
    # # subset dates data frame to get opening date for given port group (not specific for year)
    # open_info <- filter(dates_df, pcgroup == p)
    # open_month <- month(filter(dates_df, pcgroup == p)$odate); open_day <- day(filter(dates_df, pcgroup == p)$odate)
    # # create open date object for the current fishtix year, 'y'
    # if(year(open_info$odate) == open_info$crab_year){
    #   open_date <- paste0(y, "-", open_month, "-", open_day)
    # } else{
    #   open_date <- paste0(y+1, "-", open_month, "-", open_day)
    # }
    # open_date <- ymd(open_date)
    
    # pull fish ticket data for given year, port group, before Dcrab opened
    dat <- fishtix %>%
      #filter(crab_year == 2008 & IOPAC == "Bodega Bay") # this line is just for running through for loop manually
      filter(crab_year == y & IOPAC == iopac) #%>% # this is the line we want to active in the for loop
       #filter(tdate < open_date) # updated to iopac 03-10-21, was IOPAC == p
    
    # pull fish ticket data for given year, port group, after Dcrab opened
    # late_dat <- fishtix %>%
    #   filter(crab_year == y & pcgroup == p) %>%
    #   filter(tdate >= open_date)
    # 
    # #### Read in metier names key ####
    # met_names <- read.delim(here::here(indir, paste0(p, "_metier_names.txt")), stringsAsFactors = FALSE)
    
    
    ##### Create igraph object ####
    
    ## using participation_network function *if write_out = true, set pcid_choose/year_choose to p/y for file naming!
    tmp_g <- participation_network_crabyear(dat, filter = TRUE, edge_type="vessels", filter_subgraph =FALSE, 
                                            min_vessels = vessel_cutoff, min_contribution = contr_cutoff, 
                                            min_rev = vessel_rev_cutoff, min_rev_indiv = fish_rev_cutoff,
                                            write_out=TRUE, out_dir=outdir, pcid_choose=iopac, year_choose=y) # updated to pcid_choose=iopac 03-10-21
    if(any(is.na(tmp_g))){
      message("\n\n WARNING: No fishing at ", iopac, " during in crab year ", y, "!\n\n") # updated to iopac 03-10-21, was No fishing at ", p, "
    } else if(vcount(tmp_g) > 0){
      ## add in the full metier names as vertex attribute
      filtered_metier_info <- as.data.frame(x = V(tmp_g)$name)
      colnames(filtered_metier_info) <- "SPGRPN2" #"metier.abbrev"
      filtered_metier_info <- left_join(filtered_metier_info, node_names, by = "SPGRPN2")
      V(tmp_g)$common_name <- as.character(filtered_metier_info$species_group_2_label2)
      ## save vertex colors as attribute
      vertex_cols <- vertex_color(g = tmp_g)
      V(tmp_g)$colors <- vertex_cols
      ## save objects
      rds_name <- ifelse(rm_crab,  paste0("igraph_", iopac, "_", y, "_rmCrab.rds"),  paste0("igraph_", iopac, "_", y, ".rds")) # updated to iopac 03-10-21, was "igraph_", p
      saveRDS(tmp_g, file = here::here(outdir, rds_name))
      
    # Calculate network-level statistics
    ## using the `net_stats` function
    g_stats <- net_stats(tmp_g, y = y, stats="network")
    #g_stats <- net_stats(tmp_g, y = 2008, stats="network")
    ## add on period (closure/open) and port group info
    g_stats <- g_stats %>%
      #mutate(period = rep("early", times=dim(g_stats)[1])) %>%
      mutate(pcgroup = rep(iopac, times=dim(g_stats)[1])) # updated to iopac 03-10-21, was rep(p
    } #end 'else' close-g has > 0 vertex
    
    #### SAVE NETWORK STATS OUTPUT ####
    ## merge network stats from closure / open period
    # if(length(V(tmp_g)$name) == 0){
    #   tmp_stats <- g_stats
    # } else{
    #   tmp_stats <- rbind(g_stats, late_stats)
    # }
    ## append to network stats output data frame
    if(exists('mystats')){
      mystats <- rbind(mystats, g_stats)
    } else{
      mystats <- g_stats
    }
    
  } #end (p in myports)
  cat("\nfinished with all port groups in crab year ", y, "\n-----\n")
} #end (y in years)
```
<br>

Check the network stats data frame
```{r eval=TRUE}
head(mystats)
```
<br>


Write stats out to file. 2008_2015_CA_CrabPorts_NoSubgraph_01contribution.csv
```{r eval=TRUE}
filename <- paste0("NetworkStats_", min(years), "_", max(years), "_", length(myports), "pcgroups_", ifelse(rm_crab, 'rmCrab_','_'), contr_cutoff*100, "pContribution.csv")

write.csv(mystats,here::here(statdir, filename), row.names=FALSE)
```
<br>


# 3: Seasonal Participation Network Statistics

If you have already produced a `graph` object for each fisheries participation network, this section of the script will read in those objects and calculate network-level statistics

```{r eval=FALSE}
for(y in years){
  for(iopac in myports){
    cat(iopac, "\n")
    
    # Read in igraph object
    rds_name <- ifelse(rm_crab,  paste0("igraph_", iopac, "_", y, "_rmCrab.rds"),  paste0("igraph_", iopac, "_", y, ".rds"))
    ## (if the object exists)
    if(file_test("-f", here::here(outdir, rds_name))){
      tmp_g <- readRDS(here::here(outdir, rds_name))
      
      # Calculate network-level statistics
      ## using the `net_stats` function
      g_stats <- net_stats(tmp_g, y = y, stats="network")
      ## add on period (closure/open) and port group info
      g_stats <- g_stats %>%
        #mutate(period = rep("early", times=dim(g_stats)[1])) %>%
        mutate(pcgroup = rep(iopac, times=dim(g_stats)[1]))
      ## save
      tmp_stats <- g_stats
    } else{message("Warning: Could not find early season network for ",iopac,", ", y, "\n")}
    
    
    #### SAVE NETWORK STATS OUTPUT ####
    ## merge network stats from closure / open period
    # if(length(V(tmp_g)$name) == 0){
    #   tmp_stats <- g_stats
    # } else{
    #   tmp_stats <- rbind(g_stats, late_stats)
    # }
    ## append to network stats output data frame
    if(exists('mystats')){
      mystats <- rbind(mystats, g_stats)
    } else{
      mystats <- g_stats
    }
    
    
  } #end (p in myports)
  cat("\nfinished with all port groups in crab year ", y, "\n-----\n")
} #end (y in years)
```
<br>

Check the network stats data frame
```{r eval=FALSE}
head(mystats)
```
<br>


Write stats out to file. 2008_2015_CA_CrabPorts_NoSubgraph_01contribution.csv
```{r eval=FALSE}
filename <- paste0("NetworkStats_", min(years), "_", max(years), "_", length(myports), "pcgroups_", ifelse(rm_crab, 'rmCrab_','_'), contr_cutoff*100, "pContribution.csv")

write.csv(mystats,here::here(statdir, filename), row.names=FALSE)
```
<br>

---

Script Runtime:
```{r}
Sys.time() - script_start_time
```



