---
title: "Compare Timeseries of Network Stats Across Port Groups and to Other ESR Indicators"
author: "M. Fisher, J. Samhouri"
date: "Written Dec 30, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

This document graphs network metrics over time for the US West Coast, and compares metrics across port groups and to other indicators in the ESR.


The core network metric of interest on its own is Edge Density (alt: Mean Degree).

The two core network metrics of interest for comparison to the Habitat Compression Index are: 

1. Node Strength

2. Eigen Vector Centrality.

Initially, we consider the above 2 metrics for tuna in ports in WA and OR.

**Note that many network metrics are sensitive to network size, which will vary over the time series.** Network size has been incorporated into each graph as ___. Use the timeseries graphs for visual comparison of different network forms, but not for drawing major conclusions around network change / sensitivity. 



# Setup

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("foreign")) {install.packages("foreign")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("facetscales")) {devtools::install_github("zeehio/facetscales"); library(facetscales)}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
if(!require("ggpubr")) {install.packages("ggpubr")}
if(!require("cowplot")) {install.packages("cowplot")}
if(!require("viridis")) {install.packages("viridis")}
if(!require("PNWColors")) {devtools::install_github("jakelawler/PNWColors"); library(PNWColors)}
```
<br>

Select your directories.
```{r}
## directory with network statistics 
indir <- 'results/statistics'

## directory with other input data
other_indir <- 'data/input'

## directory to save figures
pngdir <- 'results/figures'
```
<br>

What are the names of the `.csv` files with the network statistics that you want to compare? Would not suggest listing more than 4. Files should all be in the input directory (`indir`) specified above.
```{r}
filenames <- c("VesselLevelNodeStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv")
networkstats_filename <- "VesselLevelNetworkStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv"
hci_filename <- 'ei_hci_rgn1_M.csv'
port_filename <- 'iopac_conversion_table.csv'
```
<br>

Provide descriptive names to differentiate the datasets drawn from the files listed above. 
```{r}
dataset_names <- c("Node_Stats")
dataset_names2 <- c("Network_Stats")
```
<br>

Specify which network metrics to retain from the full data set.
```{r}
node_metrics <- c('node_strength','eigen')
network_metrics <- c('ed') # ,'ed_weighted'
```
<br>

Specify which fisheries to retain from the full data set.
```{r}
myfisheries <- c('Tuna')
```
<br>

Specify which port groups to retain from the full data set.
```{r}
myports <- c("North WA Coast", "Puget Sound", "WA Coast", "Astoria", "Tillamook", "Newport", "Coos Bay") # drop "Brookings" for now to correspond with HCI Region 4 and because 7 colors kind of max things out?
```
<br>



# Read in Data


Network metrics from different datasets. 
```{r}
for(i in seq(1,length(filenames))){
  tmpdat <- read.csv(here::here(indir, filenames[i])) %>%
    dplyr::select(pcgroup, year, metier.cm, all_of(node_metrics)) %>%
    filter(pcgroup %in% myports) %>%
    filter(metier.cm %in% myfisheries) %>%
    mutate(dataset=dataset_names[i])
  if(i==1){
    dat <- tmpdat
  } else{ dat <- bind_rows(dat,tmpdat)}
}

head(dat)

for(i in seq(1,length(networkstats_filename))){
  tmpdat2 <- read.csv(here::here(indir, networkstats_filename[i])) %>%
    dplyr::select(pcgroup, y, all_of(network_metrics)) %>%
    #filter(pcgroup %in% myports) %>%
    #filter(metier.cm %in% myfisheries) %>%
    mutate(dataset=dataset_names2[i])
  if(i==1){
    dat2 <- tmpdat2
  } else{ dat2 <- bind_rows(dat2,tmpdat2)}
}

head(dat2)

# 01-06-22: note jameal updated OSD, SD, and SDA ports to have the same lat long as OCN. Amanda P has made the same fix. Also reassigned MCR to the Eureka IOPAC port group,  Amanda P agreed with this change
iopac_lookup <- read_csv(here::here(other_indir, port_filename))

```
<br>

Check to make sure that all of the datasets were read in. 
```{r echo=FALSE}
with(dat,table(dataset,pcgroup))
```
<br>

Pivot longer to graph multiple metrics at once.
```{r}
dat_long <- dat %>%
  pivot_longer(cols=c(all_of(node_metrics)), names_to = "metric", values_to = "value")
```
<br>

Clean up metric names.
```{r}
dat_long$metric <- recode(dat_long$metric, 
                          node_strength = "Node Strength",
                          eigen = "Importance")
#dat_long$metric <- factor(dat_long$metric, levels=c("Size", "Edge Density","Mean Degree", "Central.", "Central. (UnW)", "Mod.", "Mod. (UnW)"))

dat_long$pcgroup <- factor(dat_long$pcgroup, levels=myports)

colnames(dat2) <- c(
  'Port Group','Year','Edge Density','dataset'
)

```
<br>

Monthly Habitat Compression Index over 43.5-48N (Region 1). Summarize for spring months only (Apr-Jun).
```{r}
spring_months <- c(4,5,6)

hci <- read_csv(here::here(other_indir,hci_filename)) %>%
  dplyr::select(year, month, data) %>%
  filter(year >= 2005) %>%
  filter(month %in% spring_months) %>%
  group_by(year) %>%
  summarise(
    annual_mean = mean(data),
    .groups = 'drop'
  )

glimpse(hci)

```
<br>

# 2. Graph Timeseries

## Figure x.Time series of node strength 

... at WA and OR ports from 2004-2020. Node strength was derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

```{r}
plot_metric="Node Strength"
metric_label="Node Strength"
```
<br>

```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
plotdat <- dat_long %>% filter(metric==plot_metric)

# color palette
pal=pnw_palette("Starfish",length(unique(plotdat$pcgroup)), type = "discrete") # or choose a different PNWColors palette

# main metric plot
figx <- ggplot(plotdat, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$year), max(plotdat$year), by=2), labels=seq(min(plotdat$year), max(plotdat$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Node Strength of Tuna in Fisheries Participation Networks',
       subtitle = '2004-2021',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figx
```
<br>
<br>
<br>

Save plot as a png, with filename: Node Strength of Tuna in Fisheries Participation Networks 2004-2021
```{r echo=FALSE}
png(here::here(pngdir, "Node Strength of Tuna in Fisheries Participation Networks 2004-2021.png"),
    res=200,height=800,width=1700)
figx
dev.off()
```
<br>

## Figure y.Time series of importance

... at WA and OR ports from 2004-2020. Importance (eigenvalue centrality) was derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

```{r}
plot_metric="Importance"
metric_label="Importance"
```
<br>

```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
plotdat <- dat_long %>% filter(metric==plot_metric)

# color palette
pal=pnw_palette("Starfish",length(unique(plotdat$pcgroup)), type = "discrete") # or choose a different PNWColors palette

# main metric plot
figy <- ggplot(plotdat, aes(x=year, y=value, color=pcgroup)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(nrow = 2, vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  scale_x_continuous(breaks=seq(min(plotdat$year), max(plotdat$year), by=2), labels=seq(min(plotdat$year), max(plotdat$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Importance of Tuna in Fisheries Participation Networks',
       subtitle = '2004-2021',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figy
```
<br>
<br>
<br>

Save plot as a png, with filename: Importance of Tuna in Fisheries Participation Networks 2004-2021
```{r echo=FALSE}
png(here::here(pngdir, "Importance of Tuna in Fisheries Participation Networks 2004-2021.png"),
    res=200,height=800,width=1700)
figy
dev.off()
```
<br>



## Figure z.Time series of HCI for Region 1 

```{r echo=FALSE, fig.width=10}

# main metric plot
figz <- ggplot(hci, aes(x=year, y=annual_mean)) + 
  geom_point(size=2) + geom_line() +
  scale_x_continuous(breaks=seq(min(plotdat$year), max(plotdat$year), by=2), labels=seq(min(plotdat$year), max(plotdat$year), by=2)) +
  ylim(c(0,1)) +
  labs(title='Habitat Compression Index for Region 1, 43.5-48N',
       subtitle = 'Apr-Jun 2005-2021',
       x = "Year",
       y = 'Habitat Area\n(fraction below monthly threshold)'
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
figz
```
<br>
<br>
<br>

Save plot as a png, with filename: Spring Habitat Compression Index for Region 1, 43.5-48N, 2005-2021
```{r echo=FALSE}
png(here::here(pngdir, "Spring Habitat Compression Index for Region 1, 43.5-48N, 2005-2021.png"),
    res=200,height=800,width=1700)
figz
dev.off()
```
<br>



## Figure aa. Time series of edge density

... at all ports from 2004-2020. Edge density was derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

(Drop Other Coastal Washington & Unknown Ports & Columbia River)

```{r}
plot_metric="Edge Density"
metric_label="Edge Density"
```
<br>

```{r echo=FALSE, fig.width=10}

allports_sorted <- dat2 %>% left_join(iopac_lookup %>% select(IOPAC, PACFIN_LAT), by = c('Port Group' = 'IOPAC')) %>% select(`Port Group`, PACFIN_LAT) %>% group_by(`Port Group`) %>% summarise(lat = mean(PACFIN_LAT)) %>% arrange(-lat) %>% select(`Port Group`) %>% pull() # use arrange(-lat) if coord_flip() is NOT used

dat2$`Port Group` <- factor(dat2$`Port Group`, levels=allports_sorted)

# main metric plot
fig_ed <- ggplot(dat2 %>% filter( `Port Group` != "Other Coastal Washington" & `Port Group` != "Unknown Ports" & `Port Group` != "Columbia River"), 
                 aes(x=Year+1, y=`Edge Density`, colour=`Port Group`)) + 
  geom_point(size=2) + geom_line() +
  facet_wrap(vars(`Port Group`),labeller = label_wrap_gen(16)) +
  scale_colour_viridis(discrete = TRUE) +
  scale_x_continuous(breaks=seq(min(dat2$Year)+1, max(dat2$Year)+1, by=2), labels=seq(min(dat2$Year)+1, max(dat2$Year)+1, by=2)) +
  ylim(c(0,1)) +
  labs(title='Edge Density in Fisheries Participation Networks',
       subtitle = '2005-2021',
       x = "Year",
       y = metric_label
       ) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_text(size=13),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank()) + 
  theme(legend.position = 'none')
fig_ed
```
<br>
<br>
<br>

Save plot as a png, with filename: Edge Density in Fisheries Participation Networks 2004-2021
```{r echo=FALSE}
png(here::here(pngdir, "Edge Density in Fisheries Participation Networks 2005-2021.png"),
    res=200,height=1500,width=1700)
fig_ed
dev.off()
```
<br>







