---
title: "Compare Network Stats Across Ports"
author: "M. Fisher, J. Samhouri"
date: "Written Dec 30, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

This document graphs network metrics by port group for the US West Coast.


The two core network metrics are: 

1. Betweenness

2. Eigen Vector Centrality.

We focus on these to get a sense of which nodes involve consistently heavy levels of cross-fishery participation across port groups.


**Note that many network metrics are sensitive to network size, which will vary over the time series.** Network size has been incorporated into each graph as ___. Use the graphs for visual comparison of different network forms, but not for drawing major conclusions around network change / sensitivity. 



# Setup

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("foreign")) {install.packages("foreign")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("facetscales")) {devtools::install_github("zeehio/facetscales"); library(facetscales)}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("viridis")) {install.packages("viridis")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
if(!require("ggpubr")) {install.packages("ggpubr")}
if(!require("cowplot")) {install.packages("cowplot")}
if(!require("PNWColors")) {devtools::install_github("jakelawler/PNWColors"); library(PNWColors)}
```
<br>

Select your directories.
```{r}
## directory with network statistics 
indir <- 'results/statistics'

## directory with port group lookup key
indir2 <- 'data/input'

## directory to save figures
pngdir <- 'results/figures'
```
<br>

What are the names of the `.csv` files with the network statistics that you want to compare? Would not suggest listing more than 4. Files should all be in the input directory (`indir`) specified above.
```{r}
filenames <- c("VesselLevelNodeStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv") 

port_filename <- 'iopac_conversion_table.csv'

#"VesselLevelNetworkStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv")

```
<br>

Provide descriptive names to differentiate the datasets drawn from the files listed above. 
```{r}
dataset_names <- c("NodeStats")

#"NetworkStats")
```
<br>

Specify which network metrics to retain from the full data set.
```{r}
node_metrics <- c('basic_btwn','eigen')
```
<br>

<!-- Specify which fisheries to retain from the full data set. -->
<!-- ```{r} -->
<!-- myfisheries <- c('Non-DTS groundfish') -->
<!-- ``` -->
<!-- <br> -->

Specify which port groups to retain from the full data set. Focus on all port groups
```{r}

# 01-06-22: note jameal updated OSD, SD, and SDA ports to have the same lat long as OCN. Amanda P has made the same fix. Also reassigned MCR to the Eureka IOPAC port group,  Amanda P agreed with this change
iopac_lookup <- read_csv(here::here(indir2, port_filename))

all_iopac <- iopac_lookup %>% select(IOPAC) %>% distinct() %>% pull()

# ca_or_ports <- iopac_lookup %>% filter(`Agency Name` != "WDFW") %>% select(IOPAC) %>% distinct() %>% pull() 
# 
wa_ports <- iopac_lookup %>% filter(`Agency Name` == "WDFW") %>% select(IOPAC) %>% distinct() %>% pull()

or_ports <- iopac_lookup %>% filter(`Agency Name` == "ODFW") %>% select(IOPAC) %>% distinct() %>% pull()

morrobaymaxlat <- iopac_lookup %>% filter(IOPAC == "Morro Bay") %>% summarise(maxlat = max(PACFIN_LAT)) %>% pull()

nca_ports <- iopac_lookup %>% filter(`Agency Name` == "CDFW" & PACFIN_LAT > morrobaymaxlat) %>% select(IOPAC) %>% distinct() %>% pull()

sca_ports <- iopac_lookup %>% filter(`Agency Name` == "CDFW" & PACFIN_LAT <= morrobaymaxlat) %>% select(IOPAC) %>% distinct() %>% pull()


```
<br>



# Read in Data


Network metrics from different datasets. 
```{r}
for(i in seq(1,length(filenames))){
  tmpdat <- read.csv(here::here(indir, filenames[i])) %>%
    dplyr::select(pcgroup, year, metier.cm, all_of(node_metrics)) %>%
    filter(pcgroup %in% all_iopac) %>%
    #filter(metier.cm %in% myfisheries) %>%
    mutate(dataset=dataset_names[i])
  if(i==1){
    dat <- tmpdat
  } else{ dat <- bind_rows(dat,tmpdat)}
}

head(dat)
```
<br>

Check to make sure that all of the datasets were read in. 
```{r echo=FALSE}
with(dat,table(dataset,pcgroup))
```
<br>

Pivot longer to graph multiple metrics at once.
```{r}
dat_long2 <- dat %>%
  pivot_longer(cols=c(all_of(node_metrics)), names_to = "metric", values_to = "value")
```
<br>

Clean up metric names.
```{r}
dat_long2$metric <- recode(dat_long2$metric, 
                          basic_btwn = "Betweenness",
                          eigen = "Importance")
#dat_long$metric <- factor(dat_long$metric, levels=c("Size", "Edge Density","Mean Degree", "Central.", "Central. (UnW)", "Mod.", "Mod. (UnW)"))

allports_sorted <- dat_long2 %>% left_join(iopac_lookup %>% select(IOPAC, PACFIN_LAT), by = c('pcgroup' = 'IOPAC')) %>% select(pcgroup, PACFIN_LAT) %>% group_by(pcgroup) %>% summarise(lat = mean(PACFIN_LAT)) %>% arrange(-lat) %>% select(pcgroup) %>% pull() # use arrange(-lat) if coord_flip() is NOT used

dat_long2$pcgroup <- factor(dat_long2$pcgroup, levels=allports_sorted)

```
<br>



# 2. Graph Box Plots

## Figure x. Box plots of Betweenness and Importance

... at all IOPAC ports from 2016-2020. Network metrics were derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

Make WA, OR, NCA, SCA data frames and color palettes.
```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
years <- seq(2016,2020,1)
plotdat_wa <- dat_long2 %>% filter(year %in% years & pcgroup %in% wa_ports)
plotdat_or <- dat_long2 %>% filter(year %in% years & pcgroup %in% or_ports)
plotdat_nca <- dat_long2 %>% filter(year %in% years & pcgroup %in% nca_ports)
plotdat_sca <- dat_long2 %>% filter(year %in% years & pcgroup %in% sca_ports)

```

Make WA plot
```{r echo=FALSE, fig.width=10}
# main metric plot
fig_wa <- ggplot(plotdat_wa, aes(x=metier.cm, y=value, fill=metier.cm)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  facet_grid(
    vars(pcgroup), 
    vars(metric), 
    #scales="free_y",
    labeller = label_wrap_gen(15)
    ) +
  scale_fill_viridis(discrete = TRUE) +
  #scale_fill_manual(values=pal_or) +
  ylim(c(0,1)) +
  labs(title='Betweenness and Importance of Species Groups\nin Fisheries Participation Networks: Washington',
       subtitle = '2017-2021',
       x = "Species Group",
       y = "Value"#,
       #fill = 'Port Group'
       ) +
  #coord_flip() +
  theme_bw() +
  theme(legend.position = 'none') +
  theme(axis.text.x=element_text(angle=45,vjust=1,hjust=1))#,
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
fig_wa
```
<br>
<br>
<br>

Make OR plot
```{r echo=FALSE, fig.width=10}
# main metric plot
fig_or <- ggplot(plotdat_or, aes(x=metier.cm, y=value, fill=metier.cm)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  facet_grid(
    vars(pcgroup), 
    vars(metric), 
    #scales="free_y",
    labeller = label_wrap_gen(12)
    ) +
  scale_fill_viridis(discrete = TRUE) +
  #scale_fill_manual(values=pal_or) +
  ylim(c(0,1)) +
  labs(title='Betweenness and Importance of Species Groups\nin Fisheries Participation Networks: Oregon',
       subtitle = '2017-2021',
       x = "Species Group",
       y = "Value"#,
       #fill = 'Port Group'
       ) +
  #coord_flip() +
  theme_bw() +
  theme(legend.position = 'none') +
  theme(axis.text.x=element_text(angle=45,vjust=1,hjust=1))#,
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
fig_or
```
<br>
<br>
<br>

Make NCA plot
```{r echo=FALSE, fig.width=10}
# main metric plot
fig_nca <- ggplot(plotdat_nca, aes(x=metier.cm, y=value, fill=metier.cm)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  facet_grid(
    vars(pcgroup), 
    vars(metric), 
    #scales="free_y",
    labeller = label_wrap_gen(12)
    ) +
  scale_fill_viridis(discrete = TRUE) +
  #scale_fill_manual(values=pal_or) +
  ylim(c(0,1)) +
  labs(title='Betweenness and Importance of Species Groups\nin Fisheries Participation Networks:\nNorthern and Central California',
       subtitle = '2017-2021',
       x = "Species Group",
       y = "Value"#,
       #fill = 'Port Group'
       ) +
  #coord_flip() +
  theme_bw() +
  theme(legend.position = 'none') +
  theme(axis.text.x=element_text(angle=45,vjust=1,hjust=1))#,
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
fig_nca
```
<br>
<br>
<br>


Make SCA plot
```{r echo=FALSE, fig.width=10}
# main metric plot
fig_sca <- ggplot(plotdat_sca, aes(x=metier.cm, y=value, fill=metier.cm)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  facet_grid(
    vars(pcgroup), 
    vars(metric), 
    #scales="free_y",
    labeller = label_wrap_gen(12)
    ) +
  scale_fill_viridis(discrete = TRUE) +
  #scale_fill_manual(values=pal_or) +
  ylim(c(0,1)) +
  labs(title='Betweenness and Importance of Species Groups\nin Fisheries Participation Networks:\nSouthern California',
       subtitle = '2017-2021',
       x = "Species Group",
       y = "Value"#,
       #fill = 'Port Group'
       ) +
  #coord_flip() +
  theme_bw() +
  theme(legend.position = 'none') +
  theme(axis.text.x=element_text(angle=45,vjust=1,hjust=1))#,
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
fig_sca
```
<br>
<br>
<br>

Save plots as pngs, with filename: State. Betweenness and Importance of Species Groups in Fisheries Participation Networks 2017-2021.png
```{r echo=FALSE}
png(here::here(pngdir, paste0("Washington. Betweenness and Importance of Species Groups in Fisheries Participation Networks 2017-2021.png")),
    res=200,height=1200,width=1400)
fig_wa
dev.off()

png(here::here(pngdir, paste0("Oregon. Betweenness and Importance of Species Groups in Fisheries Participation Networks 2017-2021.png")),
    res=200,height=1200,width=1400)
fig_or
dev.off()

png(here::here(pngdir, paste0("Northern and Central CA. Betweenness and Importance of Species Groups in Fisheries Participation Networks 2017-2021.png")),
    res=200,height=1200,width=1400)
fig_nca
dev.off()

png(here::here(pngdir, paste0("Southern CA. Betweenness and Importance of Species Groups in Fisheries Participation Networks 2017-2021.png")),
    res=200,height=1200,width=1400)
fig_sca
dev.off()
```
<br>
<br>
<br>

Summarize proportion of port groups with median values of betweenness and importance >=0.25 from 2016-2020.

Make data frame
```{r echo=FALSE, fig.width=10}
years <- seq(2016,2020,1)

# get rid of weird duplicates for WA Coast port. still don't know why this is happening in 01_create_betworks_cciea.Rmd. JS 02-13-2022
plotdat_summary <- dat_long2 %>% 
  filter(year %in% years) %>%
  select(-dataset) %>%
  distinct() %>% # this line deals with weird WA problem
  pivot_wider(
      names_from = metric,
      values_from = value
      ) %>%
  group_by(pcgroup, metier.cm) %>%
  summarise(
    median_btwn = median(Betweenness),
    median_imp = median(Importance),
    .groups = "drop" 
  ) %>%
  group_by(metier.cm) %>%
  summarise(
    prop_btwn_lg = length(which(median_btwn >= 0.25))/length(median_btwn),
    prop_imp_lg = length(which(median_imp >= 0.25))/length(median_imp)
    ) %>%
  pivot_longer(cols = c('prop_btwn_lg','prop_imp_lg'), names_to = "metric", values_to = "value") 

plotdat_summary$metric <- recode(plotdat_summary$metric, 
                          prop_btwn_lg = "Betweenness",
                          prop_imp_lg = "Importance")
  

glimpse(plotdat_summary)  
  
```
<br>
<br>
<br>

Make plot
```{r echo=FALSE, fig.width=10}
# main metric plot
fig_prop <- ggplot(plotdat_summary, aes(x=metier.cm, y=value, fill=metier.cm)) + 
  geom_col() +
  facet_grid(
    cols = vars(metric), 
    #scales="free_y",
    labeller = label_wrap_gen(12)
    ) +
  scale_fill_viridis(discrete = TRUE) +
  #scale_fill_manual(values=pal_or) +
  ylim(c(0,1)) +
  labs(title='Proportion of IOPAC Port Groups with Betweenness\nand Importance Values >=0.25 for Species Groups\nin West Coast Fisheries Participation Networks',
       subtitle = '2017-2021',
       x = "Species Group",
       y = "Proportion"#,
       #fill = 'Port Group'
       ) +
  #coord_flip() +
  theme_bw() +
  theme(legend.position = 'none') +
  theme(axis.text.x=element_text(angle=45,vjust=1,hjust=1))#,
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
fig_prop
```
<br>
<br>
<br>

Save plot as png, with filename: Proportion of IOPAC Port Groups with Betweenness and Importance Values >=0.25 for Species Groups in West Coast Fisheries Participation Networks.png
```{r echo=FALSE}
png(here::here(pngdir, paste0("Proportion of IOPAC Port Groups with Betweenness and Importance Values >=0.25 for Species Groups in West Coast Fisheries Participation Networks.png")),
    res=200,height=1400,width=1700)
fig_prop
dev.off()

```
<br>
<br>
<br>

Make plot for Importance only
```{r echo=FALSE, fig.width=10}
# main metric plot
fig_prop_imp <- ggplot(plotdat_summary %>% filter(metric == "Importance"), aes(x=metier.cm, y=value, fill=metier.cm)) + 
  geom_col() +
  scale_fill_viridis(discrete = TRUE) +
  #scale_fill_manual(values=pal_or) +
  ylim(c(0,1)) +
  labs(title='Proportion of IOPAC Port Groups with Betweenness Values >=0.25\nfor Species Groups in West Coast Fisheries Participation Networks',
       subtitle = '2017-2021',
       x = "Species Group",
       y = "Proportion"#,
       #fill = 'Port Group'
       ) +
  #coord_flip() +
  theme_bw() +
  theme(legend.position = 'none') +
  theme(axis.text.x=element_text(angle=45,vjust=1,hjust=1))#,
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
fig_prop_imp
```
<br>
<br>
<br>

Save plot as png, with filename: Proportion of IOPAC Port Groups with Importance Values >=0.25 for Species Groups in West Coast Fisheries Participation Networks.png
```{r echo=FALSE}
png(here::here(pngdir, paste0("Proportion of IOPAC Port Groups with Importance Values >=0.25 for Species Groups in West Coast Fisheries Participation Networks.png")),
    res=200,height=1400,width=1700)
fig_prop_imp
dev.off()

```
<br>
<br>
<br>



