---
title: "Compare Network Stats Across Ports"
author: "M. Fisher, J. Samhouri"
date: "Written Dec 30, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

This document graphs network metrics by port group for the US West Coast.


The two core network metrics are: 

1. Node Strength

2. Eigen Vector Centrality.


**Note that many network metrics are sensitive to network size, which will vary over the time series.** Network size has been incorporated into each graph as ___. Use the graphs for visual comparison of different network forms, but not for drawing major conclusions around network change / sensitivity. 



# Setup

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("foreign")) {install.packages("foreign")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("facetscales")) {devtools::install_github("zeehio/facetscales"); library(facetscales)}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
if(!require("ggpubr")) {install.packages("ggpubr")}
if(!require("cowplot")) {install.packages("cowplot")}
if(!require("PNWColors")) {devtools::install_github("jakelawler/PNWColors"); library(PNWColors)}
```
<br>

Select your directories.
```{r}
## directory with network statistics 
indir <- 'results/statistics'

## directory with port group lookup key
indir2 <- 'data/input'

## directory to save figures
pngdir <- 'results/figures'
```
<br>

What are the names of the `.csv` files with the network statistics that you want to compare? Would not suggest listing more than 4. Files should all be in the input directory (`indir`) specified above.
```{r}
filenames <- c("VesselLevelNodeStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv") 

port_filename <- 'iopac_conversion_table.csv'

#"VesselLevelNetworkStats_2004_2020_21pcgroups_totalrev5000_indivrev500_connectivity_rmCrab_10pContribution.csv")

```
<br>

Provide descriptive names to differentiate the datasets drawn from the files listed above. 
```{r}
dataset_names <- c("NodeStats")

#"NetworkStats")
```
<br>

Specify which network metrics to retain from the full data set.
```{r}
node_metrics <- c('node_strength','eigen')
```
<br>

Specify which fisheries to retain from the full data set.
```{r}
myfisheries <- c('Non-DTS groundfish')
```
<br>

Specify which port groups to retain from the full data set. Focus on all CA and OR port groups
```{r}

# 01-06-22: note jameal updated OSD, SD, and SDA ports to have the same lat long as OCN. Amanda P has made the same fix. Also reassigned MCR to the Eureka IOPAC port group,  Amanda P agreed with this change
iopac_lookup <- read_csv(here::here(indir2, port_filename))

ca_or_ports <- iopac_lookup %>% filter(`Agency Name` != "WDFW") %>% select(IOPAC) %>% distinct() %>% pull() 

or_ports <- iopac_lookup %>% filter(`Agency Name` == "ODFW") %>% select(IOPAC) %>% distinct() %>% pull() 

morrobaymaxlat <- iopac_lookup %>% filter(IOPAC == "Morro Bay") %>% summarise(maxlat = max(PACFIN_LAT)) %>% pull()

nca_ports <- iopac_lookup %>% filter(`Agency Name` == "CDFW" & PACFIN_LAT > morrobaymaxlat) %>% select(IOPAC) %>% distinct() %>% pull() 

sca_ports <- iopac_lookup %>% filter(`Agency Name` == "CDFW" & PACFIN_LAT <= morrobaymaxlat) %>% select(IOPAC) %>% distinct() %>% pull()


```
<br>



# Read in Data


Network metrics from different datasets. 
```{r}
for(i in seq(1,length(filenames))){
  tmpdat <- read.csv(here::here(indir, filenames[i])) %>%
    dplyr::select(pcgroup, year, metier.cm, all_of(node_metrics)) %>%
    filter(pcgroup %in% ca_or_ports) %>%
    filter(metier.cm %in% myfisheries) %>%
    mutate(dataset=dataset_names[i])
  if(i==1){
    dat <- tmpdat
  } else{ dat <- bind_rows(dat,tmpdat)}
}

head(dat)
```
<br>

Check to make sure that all of the datasets were read in. 
```{r echo=FALSE}
with(dat,table(dataset,pcgroup))
```
<br>

Pivot longer to graph multiple metrics at once.
```{r}
dat_long <- dat %>%
  pivot_longer(cols=c(all_of(node_metrics)), names_to = "metric", values_to = "value")
```
<br>

Clean up metric names.
```{r}
dat_long$metric <- recode(dat_long$metric, 
                          node_strength = "Node Strength",
                          eigen = "Importance")
#dat_long$metric <- factor(dat_long$metric, levels=c("Size", "Edge Density","Mean Degree", "Central.", "Central. (UnW)", "Mod.", "Mod. (UnW)"))

nonDTSports_sorted <- dat_long %>% left_join(iopac_lookup %>% select(IOPAC, PACFIN_LAT), by = c('pcgroup' = 'IOPAC')) %>% select(pcgroup, PACFIN_LAT) %>% group_by(pcgroup) %>% summarise(lat = mean(PACFIN_LAT)) %>% arrange(lat) %>% select(pcgroup) %>% pull() # use arrange(-lat) if coord_flip() is NOT used

dat_long$pcgroup <- factor(dat_long$pcgroup, levels=nonDTSports_sorted)

```
<br>



# 2. Graph Box Plots

## Figure x. Box plots of Node Strength and Importance

... at Oregon and California ports from 2016-2020. Network metrics were derived from vessel-level networks summarized at the port level. Nodes (fisheries) are included in each network with a minimum total fisheries revenue cutoff for vessels of `$5,000` and a minimum revenue cutoff for individual fisheries of `$500`.

Make OR, NCA, SCA data frames and color palettes.
```{r echo=FALSE, fig.width=10}
# filter and re-arrange data for plotting
years <- seq(2016,2020,1)
plotdat_or <- dat_long %>% filter(year %in% years & pcgroup %in% or_ports)
plotdat_nca <- dat_long %>% filter(year %in% years & pcgroup %in% nca_ports)
plotdat_sca <- dat_long %>% filter(year %in% years & pcgroup %in% sca_ports)

# color palette
# or. 5 ports
pal_or=pnw_palette("Shuksan2",length(unique(plotdat_or$pcgroup)), type = "discrete") # or choose a different PNWColors palette
# nca. 6 ports
pal_nca=pnw_palette("Starfish",length(unique(plotdat_nca$pcgroup)), type = "discrete") # or choose a different PNWColors palette
# sca. 4 ports
pal_sca=pnw_palette("Sunset",length(unique(plotdat_sca$pcgroup)), type = "discrete") # or choose a different PNWColors palette
```

Make OR plot
```{r echo=FALSE, fig.width=10}
# main metric plot
or_fig <- ggplot(plotdat_or, aes(x=pcgroup, y=value, fill=pcgroup)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  facet_grid(rows=vars(metric), scales="free_y") +
  scale_fill_manual(values=pal_or) +
  #ylim(c(0,1)) +
  labs(title='Importance and Node Strength of Non-DTS Groundfish in Fisheries Participation Networks: Oregon',
       subtitle = '2017-2021',
       x = "Port Group",
       y = "Value",
       fill = 'Port Group'
       ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = 'none')
  #+ theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    strip.text=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
or_fig
```
<br>
<br>
<br>

Make NCA plot
```{r echo=FALSE, fig.width=10}
# main metric plot
nca_fig <- ggplot(plotdat_nca, aes(x=pcgroup, y=value, fill=pcgroup)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  facet_grid(rows=vars(metric), scales="free_y") +
  scale_fill_manual(values=pal_nca) +
  #ylim(c(0,1)) +
  labs(title='Importance and Node Strength of Non-DTS Groundfish in Fisheries Participation Networks: Northern and Central California',
       subtitle = '2017-2021',
       x = "Port Group",
       y = "Value",
       fill = 'Port Group'
       ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = 'none')
  #+ theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    strip.text=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
nca_fig
```
<br>
<br>
<br>

Make SCA plot
```{r echo=FALSE, fig.width=10}
# main metric plot
sca_fig <- ggplot(plotdat_sca, aes(x=pcgroup, y=value, fill=pcgroup)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  facet_grid(rows=vars(metric), scales="free_y") +
  scale_fill_manual(values=pal_sca) +
  #ylim(c(0,1)) +
  labs(title='Importance and Node Strength of Non-DTS Groundfish in Fisheries Participation Networks: Southern California',
       subtitle = '2017-2021',
       x = "Port Group",
       y = "Value",
       fill = 'Port Group'
       ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = 'none')
  #+ theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    strip.text=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
sca_fig
```
<br>
<br>
<br>

Save plots as pngs, with filename: State. Importance and Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2017-2021.png
```{r echo=FALSE}
png(here::here(pngdir, paste0("Oregon. Importance and Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2017-2021.png")),
    res=200,height=800,width=1700)
or_fig
dev.off()

png(here::here(pngdir, paste0("Northern and Central CA. Importance and Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2017-2021.png")),
    res=200,height=800,width=1700)
nca_fig
dev.off()

png(here::here(pngdir, paste0("Southern CA. Importance and Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2017-2021.png")),
    res=200,height=800,width=1700)
sca_fig
dev.off()
```
<br>


Focus on node strength only, compile into a single figure. Save plot as png, with filename: Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2017-2021, Oregon and California.png

```{r echo=FALSE}

or_ns_df <- or_fig$data %>% filter(metric == "Node Strength")
or_ns_fig <- ggplot(or_ns_df, aes(x=pcgroup, y=value, fill=pcgroup)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  #facet_grid(rows=vars(metric), scales="free_y") +
  scale_fill_manual(values=pal_or) +
  ylim(c(0,1)) +
  labs(title='Node Strength of Non-DTS Groundfish in\nFisheries Participation Networks: 2017-2021',
       subtitle = 'Oregon',
       x = "Port Group",
       y = "Node Strength",
       fill = 'Port Group'
       ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = 'none')
  #+ theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    strip.text=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
#or_ns_fig

nca_ns_df <- nca_fig$data %>% filter(metric == "Node Strength")
nca_ns_fig <- ggplot(nca_ns_df, aes(x=pcgroup, y=value, fill=pcgroup)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  #facet_grid(rows=vars(metric), scales="free_y") +
  scale_fill_manual(values=pal_nca) +
  ylim(c(0,1)) +
  labs(subtitle = 'Northern and Central California',
       x = "Port Group",
       y = "Node Strength",
       fill = 'Port Group'
       ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = 'none')
  #+ theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    strip.text=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
#nca_ns_fig

sca_ns_df <- sca_fig$data %>% filter(metric == "Node Strength")
sca_ns_fig <- ggplot(sca_ns_df, aes(x=pcgroup, y=value, fill=pcgroup)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, alpha = 0.3,  position = position_jitter(seed = 1, width = .2)) +
  #facet_grid(rows=vars(metric), scales="free_y") +
  scale_fill_manual(values=pal_sca) +
  ylim(c(0,1)) +
  labs(subtitle = 'Southern California',
       x = "Port Group",
       y = "Node Strength",
       fill = 'Port Group'
       ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = 'none')
  #+ theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
  #                    axis.title.x=element_text(size=13),
  #                    axis.title.y=element_text(size=13),
  #                    strip.text=element_text(size=13),
  #                    legend.title=element_text(size=13),legend.text=element_text(size=13),
  #                    panel.grid.major=element_blank(),
  #                    panel.grid.minor=element_blank())
#sca_ns_fig

plot_grid(plotlist=list(or_ns_fig, nca_ns_fig, sca_ns_fig), ncol=1, align = "v", axis = "l", labels = "AUTO", rel_heights=c(1,1,1))

png(here::here(pngdir, paste0("Node Strength of Non-DTS Groundfish in Fisheries Participation Networks 2017-2021, Oregon and California.png")),
    res=200,height=1200,width=1500)

plot_grid(plotlist=list(or_ns_fig, nca_ns_fig, sca_ns_fig), ncol=1, align = "v", axis = "l", labels = "AUTO", rel_heights=c(1,1,1))

dev.off()


```








