---
title: "Compare Network Stats Timeseries"
author: "M. Fisher"
date: "Written Aug 17, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: For SSC-ES 
fontsize: 11pt
---

# Description

Graph network metrics over time, and compare between different network forms.


The three core network metrics are: 

1. Mean Degree (alt: Edge Density)

2. Weighted Network Centralization (alt: unweighted)

3. Weighted Modularity (alt:unweighted)


**Note that many network metrics are sensitive to network size, which will vary over the time series.** Use the timeseries graphs for visual comparison of different network forms, but not for drawing major conclusions around network change / sensitivity. 


<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("foreign")) {install.packages("foreign")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("facetscales")) {devtools::install_github("zeehio/facetscales"); library(facetscales)}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("ggord")) {devtools::install_github("fawda123/ggord"); library(facetscales)}
if(!require("gridExtra")) {install.packages("gridExtra")}
if(!require("cowplot")) {install.packages("cowplot")}
if(!require("PNWColors")) {devtools::install_github("jakelawler/PNWColors"); library(PNWColors)}
```
<br>

# User Inputs

Select your directories.
```{r}
## directory with network statistics 
indir <- 'results/statistics'

## directory to save figures
pngdir <- 'results/figures'
```
<br>

What are the names of the `.csv` files with the network statistics that you want to compare? Would not suggest listing more than 4. Files should all be in the input directory (`indir`) specified above.
```{r}
filenames <- c("NetworkStats_2004_2019_21pcgroups_rmCrab_5pContribution_bulk.csv", 
               "NetworkStats_2004_2019_21pcgroups_rmCrab_10pContribution_bulk.csv")
```
<br>

Provide descriptive names to differentiate the datasets drawn from the files listed above. 
```{r}
dataset_names <- c("5pContribution",
                   "10pContribution")
```
<br>

Specify which network metrics to retain from the full data set.
```{r}
metrics <- c("mean_deg","nc_weighted","m_weighted")
```
<br>

Specify which port groups to retain from the full data set.
```{r}
myports <- c("Newport","Crescent City", "San Francisco")
```
<br>



# 1: Data


Network metrics from different datasets. 
```{r}
for(i in seq(1,length(filenames))){
  tmpdat <- read.csv(here::here(indir, filenames[i])) %>%
    dplyr::select(pcgroup, y, N, all_of(metrics)) %>%
    filter(pcgroup %in% myports) %>%
    mutate(dataset=dataset_names[i])
  if(i==1){
    dat <- tmpdat
  } else{ dat <- bind_rows(dat,tmpdat)}
}

head(dat)
```
<br>

Check to make sure that all of the datasets were read in. 
```{r echo=FALSE}
with(dat,table(dataset,pcgroup))
```
<br>



# 2. Graph Timeseries

Pivot longer to graph multiple metrics at once.
```{r}
dat_long <- dat %>%
  pivot_longer(cols=c(all_of(metrics)), names_to = "metric", values_to = "value")
```
<br>

Clean up metric names.
```{r}
dat_long$metric <- recode(dat_long$metric, 
                          N = "Size",
                          ed = "Edge Density",
                          mean_deg = "Mean Degree",
                          nc_weighted = "Central.",
                          nc = "Central. (UnW)",
                          m_weighted = "Mod.",
                          m = "Mod. (UnW)")
dat_long$metric <- factor(dat_long$metric, levels=c("Size", "Edge Density","Mean Degree", "Central.", "Central. (UnW)", "Mod.", "Mod. (UnW)"))
```
<br>

With default 'y' axis limits
```{r echo=FALSE}
pal=pnw_palette("Sailboat",length(unique(dat$dataset)), type = "discrete") # or choose a different PNWColors palette
ggplot(dat_long, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid(rows=vars(metric), cols=vars(pcgroup), scales="free_y") +
  scale_color_manual(values=pal) +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_blank(),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank())
```
<br>

Set y axis limits and breaks
```{r echo=FALSE}
scales_y <- list(
  `Edge Density` = scale_y_continuous(limits = c(0,1)),
  `Mean Degree` = scale_y_continuous(limits = c(0,max(dat$mean_deg) + 0.5)),
  `Central.` = scale_y_continuous(limits = c(0,1)),
  `Central. (UnW)` = scale_y_continuous(limits=c(0,1)),
  `Mod.` = scale_y_continuous(limits=c(-1,1)),
  `Mod. (UnW)` = scale_y_continuous(limits=c(-1,1))
)
```


```{r echo=FALSE}
pal=pnw_palette("Sailboat",length(unique(dat$dataset)), type = "discrete") # or choose a different PNWColors palette
ggplot(dat_long, aes(x=y, y=value, color=dataset)) + 
  geom_point() + geom_line() +
  facet_grid_sc(rows = vars(metric), cols = vars(pcgroup), scales = list(y = scales_y),as.table=FALSE) +
  scale_color_manual(values=pal) +
  xlab("Fishing Year") + ylab("") +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5),
                     axis.title.x=element_text(size=13),
                     axis.title.y=element_blank(),
                     strip.text=element_text(size=13),
                     legend.title=element_text(size=13),legend.text=element_text(size=13),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank())
```
<br>

















