---
title: "Plot a Batch of Annual Participation Networks"
author: "M. Fisher, J.F. Samhouri"
date: "Written Jan. 6, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

Plot a batch of annual fisheries participation networks (from a list of igraph objects) by port group and year. This script uses a circular layout, with network nodes colored according to the gear group used in each fishery. **The nodes are consistently located across plots,** so that plots can be more easily compared through time or across port groups.

 

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("igraph")) {install.packages("igraph")}
if(!require("RColorBrewer")) {install.packages("RColorBrewer")}
if(!require("stringr")) {install.packages("stringr")}
```
<br>

And calls the following functions:
```{r}
source(here::here('R/color_network_nodes.R'))
source(here::here('R/plot_network_bulk_comparable.R'))
```
<br>


# User Inputs 

Select your directories.
```{r}
## location of igraph objects
indir = 'data/networks/participation_bulk'

## output directory for network viz
pngdir = 'data/networks/participation_bulk/plots/comparable'
```
<br>

Identify the crab years, port groups, and % contributions to port-level bulk revenue that you would like to produce networks for. 
```{r}
## (crab) years
years <- seq(2004,2019)
# years <- 2008 # testing script, 1/6
# years <- 2018 # testing script, 1/8
# years <- 2014 # testing script, 1/27

## port groups
myports <- c("Puget Sound","North WA Coast","WA Coast","Other Coastal Washington","Astoria","Tillamook","Columbia River","Newport","Coos Bay","Brookings","Crescent City","Eureka","Morro Bay","Fort Bragg","Bodega Bay","San Francisco","San Diego","Monterey","Santa Barbara","Los Angeles","Unknown Ports" )
#myports <- c("BDG") # testing script, 1/6
#myports <- c("Columbia River") # testing script for single node networks, 1/7
# myports <- c("Monterey") # testing script, 1/8
# myports <- c("Bodega Bay") # testing script, 1/27/21

## To avoid inclusion of rare or minimal fishing activity, a given fishery must contribute to at least contr_cutoff% of a port's seasonal revenue to be included in the network data. This cutoff value can be increased if more stringent fishery inclusion criteria are preferred, and decreased (i.e. for summaries over shorter temporal / smaller spatial scales).
contr_cutoff <- 0.1

```
<br>


# 1: Read in data

Read the igraph objects into a list. This for loop will also apply the functions to recolor and rename vertices
```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

graphs_list <- list()
graph_names <- c()
i=1
for(p in myports){
  for(y in years){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
    if(file.exists(here::here(indir, paste0("igraph_", p, "_", y ,"_rmCrab_bulk_",contr_cutoff*100,".rds")))){
      tmpgraph <- readRDS(here::here(indir, paste0("igraph_", p, "_", y ,"_rmCrab_bulk_",contr_cutoff*100,".rds")))
      V(tmpgraph)$p <- p
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      graphs_list[[i]] <- tmpgraph
      graph_names[i] <- paste0(p,"_",y)
      i = i + 1
      } else{ y = y +1 }
  }
}
names(graphs_list) <- graph_names
```
<br>

# 2: Plot Comparable Networks

This function plots networks using the circular layout, with consistent node placement across plots to allow for visual comparison. 

An example of the circular layout:

```{r echo=FALSE}
l <-layout.circle(graphs_list[[1]])
l <- cbind(l, 1:vcount(graphs_list[[1]]))
rownames(l) <- V(graphs_list[[1]])$name
tmpnames <- list(V(graphs_list[[1]])$name)
plot(graphs_list[[1]], vertex.size = V(graphs_list[[1]])$importance/(max(V(graphs_list[[1]])$importance)*0.02), 
     layout = l, #where to put the nodes on the plot
     edge.width = sqrt(E(graphs_list[[1]])$weight)/(0.037*max(sqrt(E(graphs_list[[1]])$weight))),
     edge.curved = F, 
     axes = F,
     edge.color = 'gray',
     vertex.label = V(graphs_list[[1]])$common_name,
     vertex.color =adjustcolor(V(graphs_list[[1]])$colors, alpha.f=0.90),
     vertex.label.family = 'sans', 
     # vertex.label.color = V(g)$colors, # to have same color as vertices
     vertex.label.color = "gray25",
     vertex.label.cex= 1.2, 
     vertex.frame.color=NA
     # vertex.label.dist = c(-9,-8,-9,-10),               # these can be adjusted manually to make it look nicer, which is super annoying
     # vertex.label.degree = c(pi^(0.9),pi/5,pi/5,pi*1.1) # these can be adjusted manually to make it look nicer, which is super annoying
)
```



Plots will be written out to .png files in the function. This is because the igraph `plot` function won't save into an object that can be returned.


```{r}
glist <- graphs_list[1:25]
plot_comparable_networks(glist, outdir=pngdir)
```



