---
title: "Plot a Batch of Annual Participation Networks"
author: "M. Fisher, J.F. Samhouri"
date: "Written Jan. 6, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Summarize networks for SSC-ES presentation

Notes JFS has not yet started to code this. 08-30-2021

*** I will need to update plot_annual_networks_bulk.Rmd and plot_annual_networks.Rmd, which will require updating dependencies like plot_network_bulk.R. It may be better to take this whole section and make it a separate Rmd

*** I will want to add multipanel figures that show the network graphs from 2004-2020 for 4 example ports using plot_network_comparable.R which does not yet exist

# Description

This document generates fisheries participation networks for the US West Coast using a variety of approaches. 

It visually summarizes a subset of networks for presentation to the SSC-ES in September 2021.

Plot a batch of annual fisheries participation networks (from a list of igraph objects) by port group or state, and year. This script uses a circular layout, with network nodes colored according to the gear group used in each fishery. **The nodes are consistently located across plots,** so that plots can be more easily compared through time or across port groups.

It depends on the code "06_create_networks_4ways.Rmd", which makes network graphs and calculates network statistics.
 

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("igraph")) {install.packages("igraph")}
if(!require("RColorBrewer")) {install.packages("RColorBrewer")}
if(!require("stringr")) {install.packages("stringr")}
```
<br>

And calls the following functions:
```{r}
source(here::here('R/color_network_nodes.R'))
source(here::here('R/plot_network_bulk_comparable.R'))
```
<br>


# User Inputs 

Select your directories.
```{r}
## location of igraph objects
indir = 'data/networks/participation_bulk'

## output directory for network viz
pngdir = 'data/networks/participation_bulk/plots/comparable'
```
<br>

Identify the crab years, port groups, and % contributions to port-level bulk revenue that you would like to produce networks for. 
```{r}
## (crab) years
years <- seq(2004,2019)
# years <- 2008 # testing script, 1/6
# years <- 2018 # testing script, 1/8
# years <- 2014 # testing script, 1/27

## port groups
myports <- c("Puget Sound","North WA Coast","WA Coast","Other Coastal Washington","Astoria","Tillamook","Columbia River","Newport","Coos Bay","Brookings","Crescent City","Eureka","Morro Bay","Fort Bragg","Bodega Bay","San Francisco","San Diego","Monterey","Santa Barbara","Los Angeles","Unknown Ports" )
#myports <- c("BDG") # testing script, 1/6
#myports <- c("Columbia River") # testing script for single node networks, 1/7
# myports <- c("Monterey") # testing script, 1/8
# myports <- c("Bodega Bay") # testing script, 1/27/21

## To avoid inclusion of rare or minimal fishing activity, a given fishery must contribute to at least contr_cutoff% of a port's seasonal revenue to be included in the network data. This cutoff value can be increased if more stringent fishery inclusion criteria are preferred, and decreased (i.e. for summaries over shorter temporal / smaller spatial scales).
contr_cutoff <- 0.1

```
<br>


# 1: Read in data

Read the igraph objects into a list. This for loop will also apply the functions to recolor and rename vertices
```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

graphs_list <- list()
graph_names <- c()
i=1
for(p in myports){
  for(y in years){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
    if(file.exists(here::here(indir, paste0("igraph_", p, "_", y ,"_rmCrab_bulk_",contr_cutoff*100,".rds")))){
      tmpgraph <- readRDS(here::here(indir, paste0("igraph_", p, "_", y ,"_rmCrab_bulk_",contr_cutoff*100,".rds")))
      V(tmpgraph)$p <- p
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      graphs_list[[i]] <- tmpgraph
      graph_names[i] <- paste0(p,"_",y)
      i = i + 1
      } else{ y = y +1 }
  }
}
names(graphs_list) <- graph_names
```
<br>

# 2: Plot Comparable Networks

This function plots networks using the circular layout, with consistent node placement across plots to allow for visual comparison. 

An example of the circular layout:

```{r echo=FALSE}
l <-layout.circle(graphs_list[[1]])
l <- cbind(l, 1:vcount(graphs_list[[1]]))
rownames(l) <- V(graphs_list[[1]])$name
tmpnames <- list(V(graphs_list[[1]])$name)
plot(graphs_list[[1]], vertex.size = V(graphs_list[[1]])$importance/(max(V(graphs_list[[1]])$importance)*0.02), 
     layout = l, #where to put the nodes on the plot
     edge.width = sqrt(E(graphs_list[[1]])$weight)/(0.037*max(sqrt(E(graphs_list[[1]])$weight))),
     edge.curved = F, 
     axes = F,
     edge.color = 'gray',
     vertex.label = V(graphs_list[[1]])$common_name,
     vertex.color =adjustcolor(V(graphs_list[[1]])$colors, alpha.f=0.90),
     vertex.label.family = 'sans', 
     # vertex.label.color = V(g)$colors, # to have same color as vertices
     vertex.label.color = "gray25",
     vertex.label.cex= 1.2, 
     vertex.frame.color=NA
     # vertex.label.dist = c(-9,-8,-9,-10),               # these can be adjusted manually to make it look nicer, which is super annoying
     # vertex.label.degree = c(pi^(0.9),pi/5,pi/5,pi*1.1) # these can be adjusted manually to make it look nicer, which is super annoying
)
```



Plots will be written out to .png files in the function. This is because the igraph `plot` function won't save into an object that can be returned.


```{r}
glist <- graphs_list[1:25]
plot_comparable_networks(glist, outdir=pngdir)
```




# 3: Grid Comparable Network Plots

The previous section saves each plotted network into its own file - in some cases, it may be useful to have 2-6 network plots saved as a single image. 

```{r}
glist <- graphs_list[1:4]
```
<br>


The multi-network .png will be written out to a file in this function. Note that the file suffix should be more distinguishable / verbose; the function only adds "MultiNetwork_" in front of it, with no port group / year details like in the individual graphs. 
```{r}
plot_comparable_networks(glist, outdir=pngdir, individual=FALSE, grid=TRUE,
                         file_suffix=paste0("PugetSound_2004-07_cciea_bulk_",contr_cutoff*100),
                         grid_layout=c(2,2))
```
<br>




