---
title: "Plot Annual Participation Networks"
author: "M. Fisher, J.F. Samhouri"
date: "Written Jan. 6, 2021. Last Run `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
subtitle: Preparation for network analysis in CCIEA ESR
fontsize: 11pt
---

# Description

Plot annual fisheries participation networks (from igraph object) by port group, crab year.

Section 2 will generate network graphs that are colored according to the gear group used in each fishery. You have the option to use either the force-directed layout algorithm developed by [Fruchterman & Reingold (1991)](http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.13.8444&rep=rep1&type=pdf) or a simple circular layout. The Fruchterman & Reingold layout is informative but can result in overlap / odd clustering, especially if there is a fishery that has no / weakly connecting edges to other fisheries in the network.

Section 3 will generate simplified network graphs using a circular layout that highlight the Dungeness crab node (the visualizations used for Figure 1 in Fisher et al. 2021, PNAS). The plotting function in this section of the script could be adjusted to more flexibly highlight other fisheries.
 



<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("igraph")) {install.packages("igraph")}
if(!require("RColorBrewer")) {install.packages("RColorBrewer")}
if(!require("stringr")) {install.packages("stringr")}
```
<br>

And calls the following functions:
```{r}
source(here::here('R/color_network_nodes.R'))
source(here::here('R/plot_network.R'))
source(here::here('R/plot_simple_network.R'))
```
<br>


# User Inputs 

Select your directories.
```{r}
## location of igraph objects
indir = 'data/networks/participation'

## output directory for network viz
pngdir = 'data/networks/participation/plots'
```
<br>

Identify the crab years and port groups that you would like to produce networks for. 
```{r}
## (crab) years
years <- seq(2004,2019)
# years <- 2008 # testing script, 1/6
# years <- 2018 # testing script, 1/8
# years <- 2014 # testing script, 1/27

## port groups
myports <- c("Puget Sound","North WA Coast","WA Coast","Other Coastal Washington","Astoria","Tillamook","Columbia River","Newport","Coos Bay","Brookings","Crescent City","Eureka","Morro Bay","Fort Bragg","Bodega Bay","San Francisco","San Diego","Monterey","Santa Barbara","Los Angeles","Unknown Ports" )
#myports <- c("BDG") # testing script, 1/6
#myports <- c("Columbia River") # testing script for single node networks, 1/7
# myports <- c("Monterey") # testing script, 1/8
# myports <- c("Bodega Bay") # testing script, 1/27/21
```
<br>

Do you want to plot using the Fruchterman & Reingold layout ["fr"], a circular layout ["c"], or both ["both"]?
```{r}
my_layout <- "both"
```
<br>


# 1: Read in data

Read the igraph objects into a list. This for loop will also apply the functions to recolor and rename vertices
```{r}
#palette(brewer.pal(n = 10, name = "Set3"))
#pal3 <- brewer.pal(n = 10, name = "Set3")

graphs_list <- list()
graph_names <- c()
i=1
for(p in myports){
  for(y in years){
  #y=2013 # testing script for single node networks, 1/7
  # y=2014 # testing script for single node networks, 1/27
    if(file.exists(here::here(indir, paste0("igraph_", p, "_", y ,"_rmCrab.rds")))){
      tmpgraph <- readRDS(here::here(indir, paste0("igraph_", p, "_", y ,"_rmCrab.rds")))
      V(tmpgraph)$p <- p
      V(tmpgraph)$year <- y
      V(tmpgraph)$colors <- vertex_color(tmpgraph)
      graphs_list[[i]] <- tmpgraph
      graph_names[i] <- paste0(p,"_",y)
      i = i + 1
      } else{ y = y +1 }
  }
}
names(graphs_list) <- graph_names
```
<br>


# 2: Plot with Color-Adjusted Nodes

Plots will be written out to .png files in the function. This is because the igraph `plot` function won't save into an object that can be returned.

```{r}
lapply(graphs_list,plot_network, my_layout, outdir=pngdir)
```
<br>

An example of the F-R layout:

```{r echo=FALSE}
l <-  layout.fruchterman.reingold(graphs_list[[1]])
# l <- layout_with_fr(g) #this is not the same as above??
l <- cbind(l, 1:vcount(graphs_list[[1]]))           # switched from a solid number to a function to get number of vertices -- M.F. 11/19/2018
rownames(l) <- V(graphs_list[[1]])$name
plot(graphs_list[[1]], vertex.size = V(graphs_list[[1]])$importance/(max(V(graphs_list[[1]])$importance)*0.02), 
     layout = l, #where to put the nodes on the plot
     edge.width = sqrt(E(graphs_list[[1]])$weight)/(0.037*max(sqrt(E(graphs_list[[1]])$weight))),
     edge.curved = F, 
     axes = F,
     edge.color = 'gray',
     vertex.label = V(graphs_list[[1]])$common_name,
     vertex.color =adjustcolor(V(graphs_list[[1]])$colors, alpha.f=0.90),
     vertex.label.family = 'sans', 
     vertex.label.color = "gray25",
     vertex.label.cex= 1.2, 
     vertex.frame.color=NA
     # vertex.label.dist = c(-9,-8,-9,-10),               # these can be adjusted manually to make it look nicer, which is super annoying
     # vertex.label.degree = c(pi^(0.9),pi/5,pi/5,pi*1.1) # these can be adjusted manually to make it look nicer, which is super annoying
)
```

An example of the circular layout:

```{r echo=FALSE}
l <-layout.circle(graphs_list[[1]])
l <- cbind(l, 1:vcount(graphs_list[[1]]))
rownames(l) <- V(graphs_list[[1]])$name
tmpnames <- list(V(graphs_list[[1]])$name)
plot(graphs_list[[1]], vertex.size = V(graphs_list[[1]])$importance/(max(V(graphs_list[[1]])$importance)*0.02), 
     layout = l, #where to put the nodes on the plot
     edge.width = sqrt(E(graphs_list[[1]])$weight)/(0.037*max(sqrt(E(graphs_list[[1]])$weight))),
     edge.curved = F, 
     axes = F,
     edge.color = 'gray',
     vertex.label = V(graphs_list[[1]])$common_name,
     vertex.color =adjustcolor(V(graphs_list[[1]])$colors, alpha.f=0.90),
     vertex.label.family = 'sans', 
     # vertex.label.color = V(g)$colors, # to have same color as vertices
     vertex.label.color = "gray25",
     vertex.label.cex= 1.2, 
     vertex.frame.color=NA
     # vertex.label.dist = c(-9,-8,-9,-10),               # these can be adjusted manually to make it look nicer, which is super annoying
     # vertex.label.degree = c(pi^(0.9),pi/5,pi/5,pi*1.1) # these can be adjusted manually to make it look nicer, which is super annoying
)
```


# 3: Simple Plot

Plots will be written out to .png files in the function. This is because the igraph `plot` function won't save into an object that can be returned.

```{r}
lapply(graphs_list,plot_simple, outdir=pngdir)
```
<br>

An example of the simplified circular graph, highlighting the `WOC_CRAB` node:
```{r echo=FALSE}
highlight="WOC_CRAB"
l <-layout.circle(graphs_list[[1]])
l <- cbind(l, 1:vcount(graphs_list[[1]]))
rownames(l) <- V(graphs_list[[1]])$name
tmpnames <- list(V(graphs_list[[1]])$common_name)
V(graphs_list[[1]])$colors <- unlist(lapply(tmpnames, function (x) {ifelse(x==highlight,"darkorange1","gray25")}))
plot(graphs_list[[1]], vertex.size = V(graphs_list[[1]])$importance/(max(V(graphs_list[[1]])$importance)*0.025), 
     layout = l, #where to put the nodes on the plot
     edge.width = sqrt(E(graphs_list[[1]])$weight)/(max(sqrt(E(graphs_list[[1]])$weight))*0.10),
     edge.curved=F,
     axes = F,
     edge.color = 'gray68',
     vertex.color = V(graphs_list[[1]])$colors, 
     vertex.label = NA, 
     vertex.frame.color=NA) #vertex.label.color = '#cb4b16'
```
<br>


Put the 2019 and all Newport circular graphs in their own directories
```{r}

list.of.files_2019 <- list.files(here::here(pngdir), "2019_circular_cciea")
file.copy(here::here(pngdir,list.of.files_2019), here::here(pngdir, "2019 only"), overwrite = TRUE)

list.of.files_newport <- intersect(list.files(here::here(pngdir), pattern = "Newport"), list.files(here::here(pngdir), pattern = "circular")) 
file.copy(here::here(pngdir,list.of.files_newport), here::here(pngdir, "newport"), overwrite = TRUE)


```








